---
title: "Bipartite networks of health care workers and residents in long-term care facilities"
output:
  pdf_document:
    extra_dependencies:
      - pdflscape
      - rotating
      - tabularx
      - booktabs
      - geometry
      - graphicx
editor_options: 
  markdown: 
    wrap: 80
---

8/8/2023

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
# a few personal functions to make things easy
naton=function(x){y=x;y[is.na(x)]=0; return(y)} # convert NA to 0
nuni=function(x){length(unique(x))} # count unique values
findmiss=function(dat){M=apply(dat,2,function(x){sum(is.na(x))}); M[M>0]} # count missings in all columns
#setwd("C:/Users/u0942431/Box/Epicenter LTCF Data Analysis/Chong/Analysis")
```

## Overview
This report provides a summary of the models fitted previously, then added new models with geometrically weighted dyad-wise shared partners and healthcare activities. 

```{r packages, message=FALSE, warning=FALSE}
library(network)
library(tidyverse)
library(data.table)
library(ergm)
library(ergm.multi)
library(netplot)
library(ggplot2)
library(gridExtra)
library(parallel)
library(Rglpk)
```

Package versions:
```{r }
packageVersion("network")
packageVersion("ergm")
packageVersion("ergm.multi")
```
Load analysis ready network data - updated last on 4/5/2023.

```{r read-clean}
load('network99_f2.RData') # a list of 99 networks to fit ERGMs
load('big_net_mat_f2.RData') # a large network (all 99 networks combined) for summaries.
```

We have a total of 99 networks, and here are 4 of them:
```{r viz}
# Example of a graph
set.seed(1231)
net_figs <- vector("list", 4L)
for (i in 1:4)
  net_figs[[i]] <- nplot(
    networks[[i]],
    edge.curvature=0,
    vertex.color = c(   hcw = "blue", residents = "green"
      )[networks[[i]] %v% "group"]
    )

# Putting them together using the gridExtra package
gridExtra::grid.arrange(grobs=net_figs, nrow = 2, ncol = 2) 
```


## Descriptive summary of ergm terms in our network

```{r}
log_n <- log(summary(bignet_mat ~ degrange(0, by = "net_id")))
# Summaries model parameters () 
termsum=summary(bignet_mat ~
    edges + 
    Sum("sum" ~ Sum(log_n ~ b1factor("net_id", levels=TRUE), "a"), "log_n" ) +
    b2star(2)+
    b2concurrent+
    b1star(2)+ #not in the model but.
    b2starmix(2, "type1")+
    b1dsp(1:9)+ # HCWs sharing 1~9 residents
    b2dsp(1:9)+ # residents sharing 1~9 HCWs
    b1starmix(2, "diabetes", diff = TRUE)+
    b1starmix(2, "wound_care", diff = TRUE)+ 
    b1starmix(2, "dialysis", diff = TRUE) +
    b1starmix(2, "wheelchair", diff = TRUE) +
    b1starmix(2, "bedridden", diff = TRUE)+
    b1starmix(2, "ventilator", diff = TRUE) +
    b1starmix(2, "mdro", diff = TRUE)
)
```

```{r, echo=FALSE}
dfrt=data.table(Ergm.Terms=names(termsum), Count=as.integer(termsum))
print(dfrt)
```

# Fitting ERGMs

Common controls for ergm models: we use Stochastic approximation since it is fast. 
```{r}
default_control.1 <- list(
  seed            = 44485,
  main.method     = "Stochastic-Approximation", # Robbins-Monro Stochastic Approx.
  SA.nsubphases   = 4 * 1, # 1 the default
  SA.burnin       = 1024 * 4, # Twice the default
  SA.interval     = 1024 * 4 # four times... This is being passed to SA instead
)
```

Model 0: edges only.
Since we do not have isolates in our network by design, we restrict the network by adding offset terms b1degree(0) and b2degree(0) with coefficients -Inf.
```{r, cache = TRUE, message=FALSE, warning=FALSE, results='hide'}
res0 <- ergm(
  Networks(networks) ~
  N(~edges, ~log(n))+
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", c(default_control.1))
  )
```

Add b2 2-star (second level 2-stars) term:

```{r , cache = TRUE, message=FALSE, warning=FALSE, results='hide'}
res1 <- ergm(
  Networks(networks) ~
  N(~edges, ~log(n))+
  N(~b2star(2), ~log(n))+
  +offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", c(default_control.1))
)
```

Try replacing the b2start(2) with b2concurrent (which counts the number of HCW with two or more residents).

```{r, cache = TRUE, message=FALSE, warning=FALSE, results='hide'}
res1v2 <- ergm(
  Networks(networks) ~
  N(~edges, ~log(n))+
  N(~b2concurrent, ~log(n))+
  +offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", c(default_control.1))
)
#mcmc.diagnostics(res1v2, which='plots')
```

```{r , echo = FALSE, results='asis'}
texreg::knitreg( list('edges only'=res0, '+ b2star(2)'=res1, '+ b2concurrent (instead of b2start(2))'=res1v2),
  single.row = TRUE,
  scalebox   = .8)
```

Model fits look similar with either b2star(2) or b2concurrent, the effect of b2concurrent is in the opposite direction. 
Since the interaction between b2start(2)/b2concurrent with log(n) is not significant, we remove it from the model.

```{r , cache = TRUE, message=FALSE, warning=FALSE, results='hide'}
res1v1 <- ergm(
  Networks(networks) ~
  N(~edges, ~log(n))+
  N(~b2star(2))+
  +offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", c(default_control.1))
)

res1v2 <- ergm(
  Networks(networks) ~
    N(~edges, ~log(n))+
  N(~b2concurrent)+
  +offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", c(default_control.1))
)
#mcmc.diagnostics(res1v2, which='plots')
```

```{r , echo = FALSE, results='asis'}
texreg::knitreg( list('edges only'=res0, '+ b2star(2)'=res1v1, '+ b2concurrent'=res1v2),
  single.row = TRUE,
  scalebox   = .8)
```

Now add b2startmix (HCW type) -- but need to use the dummy version now.
```{r, message=FALSE, warning=FALSE}
# which networks has Nurse, CNA, 
# 1) distribution of network sizes (by node type) and 
# 2) correlation matrix of resident status (e.g., # of individuals with MDRO and under wound care)

# Count the number of HCWs and residents in each network
counts=lapply(networks, function(x){data.table(hcw=table(x %v% 'group')[1], 
residents=table(x %v% 'group')[2] )}) %>% rbindlist
head(counts)
counts$id=1:93

#  plot 
dplot=counts[,.N, keyby = .(hcw, residents)]

# join dplot and count on hcw and residents
dplot1=merge(dplot, counts, by=c('hcw', 'residents')) %>% arrange(hcw, residents, id)
dp2=dplot1[N==2]
dp1=dplot1[N==1]
dpa=dp2[!duplicated(dp2[,c('hcw', 'residents')]),]
dpb=dp2[duplicated(dp2[,c('hcw', 'residents')]),]
dpa$id=paste(dpa$id, dpb$id, sep='/')
dplot=rbind(dp1, dpa)
dplot$N=as.factor(dplot$N)

p1=ggplot(dplot, aes(hcw, residents, fill=N))+geom_tile()+scale_fill_manual(values=c('#a7e7a7', '#78aa78'))+theme_bw()+xlab('#HCWs')+ylab('#Residents')#+ylim(0, 18)
p1+geom_text(aes(label=id), size=3.5, color='blue')
ggsave('network_size2.pdf', width=10, height=15)
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

#view nodal attributes of bignet_mat
# see what attributes are available
#get all the nodal attributes of net and combine them into a data frame
# make a function to do this for all networks
dd=NULL
for (i in 1:93)
{net=networks[[i]] 
# create a data.frame of nodal attributes
ddi=as.data.frame(cbind(net %v% "group", 
                        net %v% "Nurse", net %v% 'CNA', net %v% 'Other', net %v% 'PTOTRT',   
                        net %v% "diabetes", net %v% "wound_care", net %v% "dialysis",net %v% "wheelchair", 
                        net %v% "bedridden", net %v% "ventilator", net %v% "mdro", 
                        id=i))
names(ddi)=c("group", "Nurse", "CNA","Other","PTOTRT", "diabetes", "wound_care", "dialysis", "wheelchair", "bedridden", "ventilator", "mdro", "NetID")
dd=rbind(dd, ddi)}
save(dd, file='dd_nodal_attr.RData')

dim(dd)
head(dd)
setDT(dd)
dd[,NetID:=as.numeric(NetID)]
dd[,size:=.N, by=NetID]
dd[size<15]

diab=dd[diabetes=='Yes', unique(NetID)]
wound_care=dd[wound_care=='Yes', unique(NetID)]
dialysis=dd[dialysis=='Yes', unique(NetID)]
wheelchair=dd[wheelchair=='Yes', unique(NetID)]
bedridden=dd[bedridden=='Yes', unique(NetID)]
ventilator=dd[ventilator=='Yes', unique(NetID)]
mdro=dd[mdro=='Yes', unique(NetID)]

Nurse=dd[Nurse==1, unique(NetID)]
CNA=dd[CNA==1, unique(NetID)]
Other=dd[Other==1, unique(NetID)]
PTOTRT=dd[PTOTRT==1, unique(NetID)]
#length of the above vectors
c(length(diab), length(wound_care), length(dialysis), length(wheelchair), length(bedridden), length(ventilator), length(mdro), 
length(Nurse), length(CNA), length(Other), length(PTOTRT))
# 93 92 52 91 79 26 50 90 90 39 57
```


```{r, cache = TRUE, message=FALSE, warning=FALSE, results='hide'}

res3 <- ergm(
  Networks(networks) ~
    N(~edges, ~log(n))+
    N(~b2star(2))+
    N(~b2starmix(2, "Nurse"), subset=Nurse)+
    N(~b2starmix(2, "PTOTRT"), subset=PTOTRT)+
    N(~b2starmix(2, "CNA"), subset=CNA)+
    N(~b2starmix(2, "Other"), subset=Other)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", c(default_control.1))
)

texreg::knitreg( list('HCW'=res2, '+ HCW nurse diff'=res3),
  single.row = TRUE, scalebox   = .8)
mcmc.diagnostics(res3)

```

```{r, include=FALSE, eval=FALSE}
#mcmc.diagnostics(res2, vars.per.page = 4, which=c('plots'))
coef(res3)
```

Add sqrt(age)
```{r , cache = TRUE, message=FALSE, warning=FALSE, results='hide'}
#res4 <- ergm(
#  Networks(networks) ~
 #   N(~edges, ~log(n))+
  #  N(~b2star(2))+
   # N(~b2starmix(2, "Nurse"), subset=Nurse)+
  #  N(~b2starmix(2, "PTOTRT"), subset=PTOTRT)+
   # N(~b2starmix(2, "CNA"), subset=CNA)+
  #  N(~b2starmix(2, "Other"), subset=Other)+
    #N(~gwb1dsp(0.25, fixed=TRUE)+gwb2dsp(0.75, fixed=TRUE))+ 
   # N(~b2cov("sqrt_age"))+
  #  offset(b1degree(0))+offset(b2degree(0)),
  #offset.coef = c(-Inf, -Inf) ,
 #control = do.call("control.ergm", c(default_control.1))
#)

# remove age and add gwb1dsp terms

#starting values use coef(res3) 
 #                N(1)~edges             N(log(n))~edges                N(1)~b2star2  N(1)~b2starmix.2.Nurse.0.0 
ini=  c(    1.590407e+00      ,         -6.627237e-01      ,       2.395023e-01      ,    -2.675583e-01 ,
 #N(1)~b2starmix.2.Nurse.0.1 N(1)~b2starmix.2.PTOTRT.0.0 N(1)~b2starmix.2.PTOTRT.0.1    N(1)~b2starmix.2.CNA.0.0 
              -6.022397e-01       ,         9.444308e-02          ,     -7.057782e-01         ,      -3.539598e-01 ,
 #   N(1)~b2starmix.2.Other.0.1      *** new variables here        offset(b1degree0) 
                    -1.620547e-01   ,        -1,    0,         -4.503600e+15 ,
        #  offset(b2degree0) 
              -4.503600e+15 )

default_control<- list(
  seed            = 44485,
  init=ini) 

res4 <- ergm(
  Networks(networks) ~
    N(~edges, ~log(n))+
    N(~b2star(2))+
    N(~b2starmix(2, "Nurse"), subset=Nurse)+
    N(~b2starmix(2, "PTOTRT"), subset=PTOTRT)+
    N(~b2starmix(2, "CNA"), subset=CNA)+
    #N(~b2starmix(2, "Other"), subset=Other)+
    N(~gwb1dsp(0.25, fixed=TRUE)+gwb2dsp(0.75, fixed=TRUE))+ 
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf) ,
 control = do.call("control.ergm", default_control)
)
mcmc.diagnostics(res4)  
save(res4, file='newres4.RData')
g=gofN(res4)
plot(g)


```

```{r , echo = FALSE, results='asis'}
texreg::knitreg( list(res3, res4),
  single.row = TRUE, scalebox   = .8)
```

```{r, cache = TRUE, message=FALSE }
      #  N(1)~edges          N(log(n))~edges              N(1)~b2star2   N(1)~b2starmix.2.Nurse.0.0  N(1)~b2starmix.2.Nurse.0.1 
ini=c(    2.94594539      ,          -1.03156069     ,        0.22193994     ,   -0.09499432       ,      -0.31145355 ,
#N(1)~b2starmix.2.PTOTRT.0.0 N(1)~b2starmix.2.PTOTRT.0.1    N(1)~b2starmix.2.CNA.0.0    N(1)~b2starmix.2.CNA.0.1     N(1)~gwb1dsp.fixed.0.25 
              0.06716365          ,       -0.21269879       ,          -0.19256182     ,         -0.31732729     ,     -1.19893075 ,
 #   N(1)~gwb2dsp.fixed.0.75           offset(b1degree0)           offset(b2degree0) 
                 0.02775008        ,   0,0,  0,0,   0,0,    0,0,             -Inf         ,               -Inf )
default_control<- list(
  seed            = 44485,
  init=ini) 

res5 <- ergm(
  Networks(networks) ~
    N(~edges, ~log(n))+
    N(~b2star(2))+
    N(~b2starmix(2, "Nurse"), subset=Nurse)+
    N(~b2starmix(2, "PTOTRT"), subset=PTOTRT)+
    N(~b2starmix(2, "CNA"), subset=CNA)+
    #N(~b2starmix(2, "Other"), subset=Other)+
    N(~gwb1dsp(0.25, fixed=TRUE)+gwb2dsp(0.75, fixed=TRUE))+ 
    #N(~ b1starmix(2, "diabetes", subset=diab))+
    N(~b1starmix(2, "wound_care"), subset=wound_care)+ 
    N(~b1starmix(2, "dialysis"), subset=dialysis) +
    N(~b1starmix(2, "wheelchair"), subset=wheelchair) +
    N(~b1starmix(2, "mdro"), subset=mdro)+
    # b1starmix(2, "bedridden", diff = FALSE)+
    #b1starmix(2, "ventilator", diff = FALSE) +   
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf) ,
 control = do.call("control.ergm", default_control)
)
#mcmc.diagnostics(res5)  
save(res5, file='newres5.RData')
mcmc.diagnostics(res5, which='plots') # 

```
so far so good.


```{r, cache = TRUE, message=FALSE }
coef(res5) 
 #                        N(1)~edges                     N(log(n))~edges                        N(1)~b2star2          N(1)~b2starmix.2.Nurse.0.0 
 ini=c(  3.325624764       ,     -1.091847998              ,           0.152487180      ,                  -0.116528818 ,
  #       N(1)~b2starmix.2.Nurse.0.1         N(1)~b2starmix.2.PTOTRT.0.0         N(1)~b2starmix.2.PTOTRT.0.1            N(1)~b2starmix.2.CNA.0.0 
                       -0.486284504    ,                     0.084032677          ,              -0.263221637     ,                   -0.147340955 
   #        N(1)~b2starmix.2.CNA.0.1             N(1)~gwb1dsp.fixed.0.25           N(1)~b1starmix.2.wound_care.HCW.No 
                   ,    -0.311708480              ,          -1.163490567          ,   0.011171241 ,
#N(1)~b1starmix.2.wound_care.HCW.Yes    N(1)~b1starmix.2.dialysis.HCW.No   N(1)~b1starmix.2.dialysis.HCW.Yes  N(1)~b1starmix.2.wheelchair.HCW.No 
                        0.042597496        ,                -0.006372741          ,               0.053473594          ,               0.036141721 ,
#N(1)~b1starmix.2.wheelchair.HCW.Yes        N(1)~b1starmix.2.mdro.HCW.No       N(1)~b1starmix.2.mdro.HCW.Yes          ,         offset(b1degree0) 
                        0.022773141     ,                   -0.003542324        ,                -0.010612774           ,                     -Inf ,
 #                 offset(b2degree0) 
                               -Inf )

default_control<- list(
  seed            = 44485,
  init=ini) 

res5.reduced <- ergm(
  Networks(networks) ~
    N(~edges, ~log(n))+
    N(~b2star(2))+
    N(~b2starmix(2, "Nurse"), subset=Nurse)+
    N(~b2starmix(2, "PTOTRT"), subset=PTOTRT)+
    N(~b2starmix(2, "CNA"), subset=CNA)+
    #N(~b2starmix(2, "Other"), subset=Other)+
    N(~gwb1dsp(0.25, fixed=TRUE))+ 
    N(~b1starmix(2, "wound_care"), subset=wound_care)+ 
    N(~b1starmix(2, "dialysis"), subset=dialysis) +
    N(~b1starmix(2, "wheelchair"), subset=wheelchair) +
    N(~b1starmix(2, "mdro"), subset=mdro)+
    # b1starmix(2, "bedridden", diff = FALSE)+
    #b1starmix(2, "ventilator", diff = FALSE) +   
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf) ,
 control = do.call("control.ergm", default_control)
)
#mcmc.diagnostics(res5)  
save(res5.reduced, file='newres5_reduced.RData')
mcmc.diagnostics(res5.reduced, which='plots') # 

```










```{r , echo = FALSE, results='asis'} 
texreg::knitreg( list(res3, res4, res5),
single.row = TRUE, scalebox   = .8)
```

The above models converged fine using the stochastic approximation, however to fit the following models, we have to use MCMLE. 
```{r }
default_control.2 <- list(
  seed            = 44485,
  main.method = "MCMLE" )
```

Replace dsp1(1:2) with geometrically weighted dyad-wise shared partners - with decay parameter fixed.

```{r , eval=FALSE}
res3 <- ergm( 
  Networks(networks) ~ 
    N(~edges, ~log(n))+
    N(~b2concurrent)+
    b2starmix(2, "type")+
    gwb1dsp(0.25, fixed=TRUE)+gwb2dsp(0.75, fixed=TRUE)+ 
  +offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.2)) 
```

This is the only model that could converge with both gwb1 and gwb2 terms. 
```{r , echo=FALSE}
load('resMCMLE_gwdspfixed_updated.RData' ) # this is done using newer version of ergm.
#mcmc.diagnostics(res3, which='plots')
```

```{r , echo = FALSE, results='asis'}
texreg::knitreg( list('fix decay MCMLE'=res3),
  single.row = TRUE, scalebox   = .8)
```

```{r, eval=FALSE, iclude=FALSE, echo=FALSE}  
res4 <- ergm( # This is the previous model with healthcare activities diff = FALSE, however we now set diff=TRUE. 
  Networks(networks) ~ 
    N(~edges, ~log(n))+
    N(~b2concurrent)+
    b2starmix(2, "type")+
    b1starmix(2, "diabetes", diff = FALSE)+
    b1starmix(2, "wound_care", diff = FALSE)+ 
    b1starmix(2, "dialysis", diff = FALSE) +
    b1starmix(2, "wheelchair", diff = FALSE) +
    b1starmix(2, "bedridden", diff = FALSE)+
    b1starmix(2, "ventilator", diff = FALSE) +
    b1starmix(2, "mdro", diff = FALSE)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.2))

load('mod4MCMLE.RData')
mcmc.diagnostics(res4, which='plots')
```

## Adding healthcare activities.
To add healthcare activities, I had to remove gwb2dsp, and add the healthcare activities one at a time.

```{r, echo=FALSE}
load('res5.RData')
load('res6.RData')
load('res7a.RData')
load('res8.RData')
load('res9.RData')
load('res10.RData')
load('res11.RData')
```

```{r, eval=FALSE, echo=FALSE}
ini=c(3.47721682,-1.07529636,-0.47033109,-0.23445749,-0.31067993 ,-0.21617495 ,-Inf ,-Inf, 
      -0.70827275 ,-0.53181691,-1.10704273,0,0,-Inf,-Inf)
default_control <- list( 
  init=ini,
  seed    = 44485,
  main.method = "MCMLE" )

res5 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~log(n))+
    N(~b2concurrent)+
    b2starmix(2, "type")+
    gwb1dsp(0.25, fixed=TRUE)+ #gwb2dsp(0.75, fixed=TRUE)+ 
    b1starmix(2, "ventilator", diff = TRUE) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control))


ini= c(3.34484429 ,-1.05566170, -0.44519306, -0.22558890 ,-0.30309431, -0.20403348, -Inf,-Inf,
       -0.65068886 ,-0.76050064, -1.09022003  ,0.03140801,  0.08758120, 0, 0,-Inf ,-Inf)

default_control <- list( 
  init=ini,
  seed    = 44485,
  main.method = "MCMLE" )
  
res6 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~log(n))+
    N(~b2concurrent)+
    b2starmix(2, "type")+
    gwb1dsp(0.25, fixed=TRUE)+ #gwb2dsp(0.75, fixed=TRUE)+ 
    b1starmix(2, "ventilator", diff = TRUE) +
    b1starmix(2, "wound_care", diff = TRUE) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control))
```

```{r , echo = FALSE, results='asis'}
texreg::knitreg( list('gwdsp1 fixed decay'=res3, '+ventilator'=res5, '+wound care'= res6, '+bedridden'=res7a, '+diabetes'= res8), 
                 single.row = TRUE, scalebox   = .8)
```

```{r, eval=FALSE, echo=FALSE}
ini= c(3.37326280, -1.07218107, -0.42527004, -0.21391778, -0.31036547 ,-0.19382235 ,-Inf ,  -Inf,
       -0.64199114,-0.82399763,-1.08407403,0.02141473,0.06471287,0.01229770,0.07969637, 0,0, -Inf , -Inf)
default_control <- list( 
  init=ini,
  seed    = 44485,
  main.method = "MCMLE" )

res7 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~log(n))+
    N(~b2concurrent)+
    b2starmix(2, "type")+
    gwb1dsp(0.25, fixed=TRUE)+ #gwb2dsp(0.75, fixed=TRUE)+ 
    b1starmix(2, "ventilator", diff = TRUE) +
    b1starmix(2, "wound_care", diff = TRUE) +
    b1starmix(2, "bedridden", diff = TRUE) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control))



ini=c(3.375912147, -1.078292657, -0.422125408, -0.210430579, -0.310929129, -0.190675862 ,-Inf , -Inf ,
      -0.636094978 ,-0.843199056, -1.072204626 , 0.018675938,
 0.056218721 , 0.010765569 , 0.076089791 , 0.005229504 , 0.036783659  ,0,0  , -Inf ,  -Inf)

default_control <- list( 
  init=ini,
  seed    = 44485)

res8 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~log(n))+
    N(~b2concurrent)+
    b2starmix(2, "type")+
    gwb1dsp(0.25, fixed=TRUE)+ #gwb2dsp(0.75, fixed=TRUE)+ 
    b1starmix(2, "ventilator", diff = TRUE) +
    b1starmix(2, "wound_care", diff = TRUE) +
    b1starmix(2, "bedridden", diff = TRUE) +
    b1starmix(2, "diabetes", diff=TRUE) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control))


ini=c( 3.4220895361, -1.0895774188 ,-0.4260651540 ,-0.2134303808, -0.3181922234 ,-0.1740911307  ,  -Inf  ,    -Inf,
       -0.6253580979 ,-0.8856941938 ,-1.0761154917 , 0.0094049780 , 0.0387256888,
  0.0093037338,  0.0707953441 , 0.0004334862 , 0.0250877874,  0.0237880914 , 0.0336166593, 0, 0.01, -Inf  ,-Inf)

default_control <- list( 
  init=ini,
  seed    = 44485)

res9 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~log(n))+
    N(~b2concurrent)+
    b2starmix(2, "type")+
    gwb1dsp(0.25, fixed=TRUE)+ #gwb2dsp(0.75, fixed=TRUE)+ 
    b1starmix(2, "ventilator", diff = TRUE) +
    b1starmix(2, "wound_care", diff = TRUE) +
    b1starmix(2, "bedridden", diff = TRUE) +
    b1starmix(2, "diabetes", diff=TRUE) +
    b1starmix(2, "mdro", diff=TRUE) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control))



ini=c( 3.4375035568, -1.0918000923, -0.4256032660, -0.2157707568, -0.3205728296 ,-0.1791093346, -Inf  , -Inf,
       -0.6274064325, -0.9046699707, -1.0780538275 , 0.0041520358 , 0.0337860322,  0.0089902734 , 0.0706534435 ,0,0,
       0.0230620999 , 0.0326673457 ,  -Inf , -Inf)

default_control <- list( 
  init=ini,
  seed    = 44485)

res10 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~log(n))+
    N(~b2concurrent)+
    b2starmix(2, "type")+
    gwb1dsp(0.25, fixed=TRUE)+ #gwb2dsp(0.75, fixed=TRUE)+ 
    b1starmix(2, "ventilator", diff = TRUE) +
    b1starmix(2, "wound_care", diff = TRUE) +
    b1starmix(2, "wheelchair", diff = TRUE) +
    b1starmix(2, "diabetes", diff=TRUE) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control))


ini=c(3.431193033, -1.075791164, -0.445590184 ,-0.224438208 ,-0.342860801, -0.172101013,    -Inf  ,   -Inf,
      -0.655549717, -0.893647216 ,-1.111844295 , 0.006933418,
  0.043404094 , 0.003345014 , 0.035416426,  0.024568491 , 0.014448774 , 0.014466141 , 0.024784239 ,0,0,        -Inf   ,      -Inf)

default_control <- list( 
  init=ini,
  seed    = 44485)

res11 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~log(n))+
    N(~b2concurrent)+
    b2starmix(2, "type")+
    gwb1dsp(0.25, fixed=TRUE)+ #gwb2dsp(0.75, fixed=TRUE)+ 
    b1starmix(2, "ventilator", diff = TRUE) +
    b1starmix(2, "wound_care", diff = TRUE) +
    b1starmix(2, "wheelchair", diff = TRUE) +
    b1starmix(2, "diabetes", diff=TRUE) +
    b1starmix(2, "dialysis", diff = TRUE) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control))
```


Some more models we fit. Non-significant terms are dropped.
```{r , echo = FALSE, results='asis'}
texreg::knitreg( list('model 9: +mdro'=res9,'model 10: +wheelchair'= res10,'model 11: +dialysis'=res11), single.row = TRUE, scalebox   = .8)
```


Just check convergence of the last model. 

```{r }
mcmc.diagnostics(res11, which='plots')
```
Log likelihood of the models

```{r}
c( res9$loglikelihood,
res10$loglikelihood,
res11$loglikelihood)
```


Check goodness of fit of the last two models (model 10 and 11) via simulation (function gof() always crashes..)

```{r}
log_n <- log(summary(bignet_mat ~ degrange(0, by = "net_id")))

# Summaries model parameters () 
sumnet =summary(bignet_mat ~
    edges + 
    Sum("sum" ~ Sum(log_n ~ b1factor("net_id", levels=TRUE), "a"), "log_n" ) +
    b2concurrent+
    b2starmix(2, "type")+
    gwb1dsp(0.25, fixed=TRUE)+ 
    b1starmix(2, "ventilator", diff = TRUE) +
    b1starmix(2, "wound_care", diff = TRUE) +
    b1starmix(2, "wheelchair", diff = TRUE) +
    b1starmix(2, "diabetes", diff=TRUE) +
    b1starmix(2, "dialysis", diff = TRUE)
)
sim11=simulate(res11, nsim=50)
sumsim=colMeans(attr(sim11,'stats'))

sim10=simulate(res10, nsim=50)
sumsim2=colMeans(attr(sim10,'stats'))
```

```{r, echo=FALSE}
options(width = 300)
dt1=data.frame(Terms=names(sumnet), observed=as.integer(sumnet), simulated=as.integer(sumsim[1:length(sumnet)]))
dt2=data.frame(Terms=names(sumsim2), simulated.mod2=as.integer(sumsim2))
dt2$Terms[1:3]=names(sumnet)[1:3]

dt12=left_join(dt1, dt2, by='Terms')
names(dt12)=c('Ergm terms','Observed','Simulated (mod 11)', 'Simulated (mod 10)')
dt12
```









