---
title: "Bipartite networks of health care workers and residents in long-term care facilities"
output:
  pdf_document:
    extra_dependencies:
      - pdflscape
      - rotating
      - tabularx
      - booktabs
      - geometry
      - graphicx
editor_options: 
  markdown: 
    wrap: 80
---

8/17/2023

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
# a few personal functions to make things easy
naton=function(x){y=x;y[is.na(x)]=0; return(y)} # convert NA to 0
nuni=function(x){length(unique(x))} # count unique values
findmiss=function(dat){M=apply(dat,2,function(x){sum(is.na(x))}); M[M>0]} # count missings in all columns
#setwd("C:/Users/u0942431/Box/Epicenter LTCF Data Analysis/Chong/Analysis_New")
```

## Overview
Recent updates: 
1. We removed 6 networks with only 1 HCW - given the bipartite and no-isolate constraints they are fixed and no variation.
2. For ergm terms present in only a subset of networks, we now restricted their evaluation in only the subset of networks where they are present.
3. There are only 3 physicians and 1 PA/NP in all the networks, therefore it is not very meaning for to evaluate the hemophily in these two types of HCWs. We grouped physician with 'Other', and PA/NP with 'Nurse'.

```{r packages, message=FALSE, warning=FALSE}
library(network)
library(tidyverse)
library(data.table)
library(ergm)
library(ergm.multi)
library(netplot)
library(ggplot2)
library(gridExtra)
library(parallel)
library(Rglpk)
```

Package versions:
```{r }
packageVersion("network")
packageVersion("ergm")
packageVersion("ergm.multi")
```
Load network data 
```{r read-clean}
load('network99_f2.RData') # a list of 93 networks to fit ERGMs
load('big_net_mat_f2.RData') # a large network (all 99 networks combined) for summaries.
```

We now have a total of 93 networks, and here are 4 of them:
```{r viz, echo=FALSE, fig.width=10, fig.height=10, fig.align='center'}
# Example of a graph
set.seed(1231)
net_figs <- vector("list", 4L)
outliers <- c(4, 54, 57, 62 ) # networks with only 1 HCW
nplot(
    networks[[4]],
    edge.curvature=0,
    vertex.color = c(   hcw = "blue", residents = "green"
      )[networks[[4]] %v% "group"]
    )
# Putting them together using the gridExtra package
pdf('outliers.pdf')
gridExtra::grid.arrange(grobs=net_figs, nrow = 2, ncol = 2) 
dev.off()

#dd[NetID==54,]

```

## Descriptive summary of ergm terms in our network
```{r, echo=FALSE}
log_n <- log(summary(bignet_mat ~ degrange(0, by = "net_id")))
# Summaries model parameters () 
termsum=summary(bignet_mat ~
    edges + 
    Sum("sum" ~ Sum(log_n ~ b1factor("net_id", levels=TRUE), "a"), "log_n" ) +
    b2star(2)+
    b2concurrent+
    b1star(2)+ #not in the model but.
    b2starmix(2, "type1")+ 
    b1dsp(1:9)+ # HCWs sharing 1~9 residents
    b2dsp(1:9)+ # residents sharing 1~9 HCWs
    b1starmix(2, "diabetes", diff = TRUE)+
    b1starmix(2, "wound_care", diff = TRUE)+ 
    b1starmix(2, "dialysis", diff = TRUE) +
    b1starmix(2, "wheelchair", diff = TRUE) +
    b1starmix(2, "bedridden", diff = TRUE)+
    b1starmix(2, "ventilator", diff = TRUE) +
    b1starmix(2, "mdro", diff = TRUE)
)
```

```{r, echo=FALSE}
dfrt=data.table(Ergm.Terms=names(termsum), Count=as.integer(termsum))
print(dfrt)
```

# Fitting ERGMs

Common controls for ergm models: we use Stochastic approximation since it is fast. 
```{r}
default_control.1 <- list(
  seed            = 44485,
  main.method     = "Stochastic-Approximation", # Robbins-Monro Stochastic Approx.
  SA.nsubphases   = 4 * 1, # 1 the default
  SA.burnin       = 1024 * 4, # Twice the default
  SA.interval     = 1024 * 4 # four times... This is being passed to SA instead
)
```

Model 0: edges only.
Since we do not have isolates in our network by design, we restrict the network by adding offset terms b1degree(0) and b2degree(0) with coefficients -Inf.
```{r, cache = TRUE, message=FALSE, warning=FALSE, results='hide'}
res0 <- ergm(
  Networks(networks) ~
  N(~edges, ~log(n))+
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", c(default_control.1))
  )
```

Add b2 2-star (second level 2-stars) term:

```{r , cache = TRUE, message=FALSE, warning=FALSE, results='hide'}
res1 <- ergm(
  Networks(networks) ~
  N(~edges, ~log(n))+
  N(~b2star(2))+
  +offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", c(default_control.1))
)
```

Try replacing the b2start(2) with b2concurrent (which counts the number of HCW with two or more residents).

```{r, cache = TRUE, message=FALSE, warning=FALSE, results='hide'}
res1v2 <- ergm(
  Networks(networks) ~
  N(~edges, ~log(n))+
  N(~b2concurrent)+
  +offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", c(default_control.1))
)
#mcmc.diagnostics(res1v2, which='plots')
```

```{r , echo = FALSE, results='asis'}
texreg::knitreg( list('edges only'=res0, '+ b2star(2)'=res1, '+ b2concurrent (instead of b2start(2))'=res1v2),
  single.row = TRUE,
  scalebox   = .8)
```

Model fits look similar with either b2star(2) or b2concurrent, the effect of b2concurrent is in the opposite direction. 
Checking interaction between b2start2 and network size log(n)
```{r , cache = TRUE, message=FALSE, warning=FALSE, results='hide'}
res1v3 <- ergm(
  Networks(networks) ~
  N(~edges, ~log(n))+
  N(~b2star(2), ~log(n))+
  +offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", c(default_control.1))
)
```

```{r , echo = FALSE, results='asis'}
texreg::knitreg( list('Model 0: edges only'=res0, 'Model 1: +b2star(2)'=res1, 'Model 1.v2+log(n)'=res1v3),
  single.row = TRUE,
  scalebox   = .8)
```

The earlier model 1 is better - remove log(n).


Before adding HCW type and resident care activities, we check the number of networks with each attribute, because most HCW types and healthcare activities are present only in some networks.
```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
# Count the number of HCWs and residents in each network
counts=lapply(networks, function(x){data.table(hcw=table(x %v% 'group')[1], 
residents=table(x %v% 'group')[2] )}) %>% rbindlist
head(counts)
counts$id=1:93

#  plot 
dplot=counts[,.N, keyby = .(hcw, residents)]
# join dplot and count on hcw and residents
dplot1=merge(dplot, counts, by=c('hcw', 'residents')) %>% arrange(hcw, residents, id)
dp2=dplot1[N==2]
dp1=dplot1[N==1]
dpa=dp2[!duplicated(dp2[,c('hcw', 'residents')]),]
dpb=dp2[duplicated(dp2[,c('hcw', 'residents')]),]
dpa$id=paste(dpa$id, dpb$id, sep='/')
dplot=rbind(dp1, dpa)
dplot$N=as.factor(dplot$N)

p1=ggplot(dplot, aes(hcw, residents, fill=N))+geom_tile()+scale_fill_manual(values=c('#a7e7a7', '#78aa78'))+theme_bw()+xlab('#HCWs')+ylab('#Residents')#+ylim(0, 18)
#p1+geom_text(aes(label=id), size=3.5, color='blue')
#ggsave('network_size2.pdf', width=10, height=15)
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@


#get all the nodal attributes of net and combine them into a data frame
dd=NULL
for (i in 1:93)
{net=networks[[i]] 
# create a data.frame of nodal attributes
ddi=as.data.frame(cbind(net %v% "group", 
                        net %v% "Nurse", net %v% 'CNA', net %v% 'Other', net %v% 'PTOTRT',   
                        net %v% "diabetes", net %v% "wound_care", net %v% "dialysis",net %v% "wheelchair", 
                        net %v% "bedridden", net %v% "ventilator", net %v% "mdro", 
                        id=i))
names(ddi)=c("group", "Nurse", "CNA","Other","PTOTRT", "diabetes", "wound_care", "dialysis", "wheelchair", "bedridden", "ventilator", "mdro", "NetID")
dd=rbind(dd, ddi)}

setDT(dd)
dd[,NetID:=as.numeric(NetID)]
dd[,size:=.N, by=NetID]
dd[size<15]

diab=dd[, n:=sum(diabetes=='Yes'), NetID][n>1, unique(NetID)]
wound_care=dd[, n:=sum(wound_care=='Yes'), NetID][n>1, unique(NetID)]
dialysis=dd[, n:=sum(dialysis=='Yes'), NetID][n>1, unique(NetID)]
wheelchair=dd[, n:=sum(wheelchair=='Yes'), NetID][n>1, unique(NetID)]
bedridden=dd[, n:=sum(bedridden=='Yes'), NetID][n>1, unique(NetID)]
ventilator=dd[, n:=sum(ventilator=='Yes'), NetID][n>1, unique(NetID)]
mdro=dd[, n:=sum(mdro=='Yes'), NetID][n>1, unique(NetID)]


# summarize number of networks with each attribute
Var=c("diabetes", "wound_care", "dialysis", "wheelchair", "bedridden", "ventilator", "mdro")
N.nets=c(length(diab), length(wound_care), length(dialysis), length(wheelchair), length(bedridden),  length(ventilator), length(mdro))
Nnets.res=data.frame(Var, N.nets)

Nurse=dd[, n:=sum(Nurse==1), NetID][n>1, unique(NetID)]
CNA=dd[, n:=sum(CNA==1), NetID][n>1, unique(NetID)]
PTOTRT=dd[, n:=sum(PTOTRT==1), NetID][n>1, unique(NetID)]
Other=dd[, n:=sum(Other==1), NetID][n>1, unique(NetID)]

# summarize number of networks with each attribute
Var=c("Nurse", "CNA",  "PTOTRT","Other")
N.nets=c(length(Nurse), length(CNA), length(PTOTRT), length(Other))
Nnets.hcw=data.frame(Var, N.nets)
```

Number of networks with each resident attributes:
```{r, echo=FALSE}
Nnets.res
```
Number of networks with each HCW attributes:
```{r, echo=FALSE}  
Nnets.hcw
```


```{r, cache = TRUE, message=FALSE, warning=FALSE, results='hide'}
res3 <- ergm(
  Networks(networks) ~
    N(~edges, ~log(n))+
    N(~b2star(2))+
    N(~b2starmix(2, "Nurse"), subset=Nurse)+
    N(~b2starmix(2, "PTOTRT"), subset=PTOTRT)+
    N(~b2starmix(2, "CNA"), subset=CNA)+
   #N(~b2starmix(2, "Other"), subset=Other)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", c(default_control.1))
)
mcmc.diagnostics(res3)
texreg::knitreg( res3, single.row = TRUE, scalebox   = .8)
g=gofN(res3)
coef(res3)

```

```{r, cache = TRUE, message=FALSE, warning=FALSE, results='hide'}

              #   N(1)~edges             N(log(n))~edges                N(1)~b2star2 
ini=c(           1.981403e+00     ,      -7.634431e-01         ,   1.675727e-01 ,
# N(1)~b2starmix.2.Nurse.0.0  N(1)~b2starmix.2.Nurse.0.1 N(1)~b2starmix.2.PTOTRT.0.0 
              -1.914323e-01  ,         -3.872203e-01      ,       1.606438e-01 ,
#N(1)~b2starmix.2.PTOTRT.0.1    N(1)~b2starmix.2.CNA.0.0    N(1)~b2starmix.2.CNA.0.1 
              -6.161217e-01     ,     -4.201434e-01    ,       -5.085695e-01 ,
 #    gwb1     offset(b1degree0)           offset(b2degree0) 
        0,      -4.503600e+15   ,       -4.503600e+15 )


default_control<- list(
  seed            = 44485,
  init=ini) 

res4 <- ergm(
  Networks(networks) ~
    N(~edges, ~log(n))+
    N(~b2star(2))+
    N(~b2starmix(2, "Nurse"), subset=Nurse)+
    N(~b2starmix(2, "PTOTRT"), subset=PTOTRT)+
    N(~b2starmix(2, "CNA"), subset=CNA)+
    N(~gwb1dsp(0.25, fixed=TRUE))+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control)
)

save(res4, file='newres4a.RData')
mcmc.diagnostics(res4)
texreg::knitreg( res4, single.row = TRUE, scalebox   = .8)g=gofN(res3)


```
```{r, cache = TRUE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

texreg::knitreg( list('HCW'=res2, '+ HCW nurse diff'=res3),
  single.row = TRUE, scalebox   = .8)
mcmc.diagnostics(res3)

```



