---
title: "modeldata"
output: html_document
date: "2022-08-19"
---

---
title: "Bipartite networks of health care workers and residents in long-term care facilities"
output:
  pdf_document:
    extra_dependencies:
      - pdflscape
      - tabularx
      - booktabs
      - geometry
      - graphicx
editor_options: 
  markdown: 
    wrap: 72
author: George G. Vega Yon
date: May 3, 2022
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Data Preparation

For this analysis we will use the following R packages:

- network: For manipulation of network data.

- data.table: For data cleaning and processing.

- ergm: For estimation and simulation of Exponential Random Graph Models.

- netplot: For visualization of networks.

- ggplot2: For general visualization.

- gridExtra: Utilities for grid graphics (base of ggplot2 and netplot.)

```{r packages, message=FALSE, warning=FALSE}
library(network)
library(data.table)
library(ergm)
library(netplot)
library(ggplot2)
library(gridExtra)
```

We start with reading the data.

```{r read-clean}
# Reading the data in
dat <- fread("data/Matched Interaction and Resident.csv")
```

Creating a few covariates to be used in the ERGM. We also need to deal with yes/no
variables to make sure these are represented as 0/1 integers in the data.

```{r}
dat[!is.na(race), white := as.integer(race == "White")]
# Wound Care
# Diabetes
# Dialysis
# Wheelchair
# Bedridden
# Ventilator
# MDRO
dat[!is.na(wound_care), wound_care := as.integer(wound_care == "Yes")]
dat[!is.na(diabetes), diabetes := as.integer(diabetes == "Yes")]
dat[!is.na(dialysis), dialysis := as.integer(dialysis == "Yes")]
dat[!is.na(wheelchair), wheelchair := as.integer(wheelchair == "Yes")]
dat[!is.na(bedridden), bedridden := as.integer(bedridden == "Yes")]
dat[!is.na(ventilator), ventilator := as.integer(ventilator == "Yes")]
dat[!is.na(mdro), mdro := as.integer(mdro == "Yes")]
```

The following function will be used to create unique identifiers for
each network. The function replaces non-alphanumeric characters from the data
with underscore and combines multiple columns into a single string.

```{r}
# Function to collapse multiple fields into a single string
# keeping only alphanumeric characters.
namemilize <- function(...) {
  do.call(
    paste,
    c(lapply(list(...), \(x) gsub("[^[:alnum:]]", "_", trimws(x))),
      list(sep = "-"))
  ) |>
    tolower() |>
    gsub(pattern = "_+", replacement = "_")
  
}

# Generating the network ID
dat[, net_id := namemilize(state, date, Fac_Name, Fac_Unit)]
```

We noticed that for some records (7) the resident ID coincides with the
room id. For now, we will remove those records:

```{r}
dat[room_num == res_id, .(room_num, res_id)]
dat <- dat[room_num != res_id]
```

Some of the health variables have missing data or are inconsistent,
showing 0/1/NA values in the same day. We will take the most frequent
value, if no NAs are registered.

```{r}
f <- \(d) {
  tmp <- table(d, useNA = "no")
  rep(as.integer(names(tmp[which.max(tmp)])), length(d))
}

dat[, wound_care := f(wound_care), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, diabetes := f(diabetes), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, dialysis := f(dialysis), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, wheelchair := f(wheelchair), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, bedridden := f(bedridden), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, ventilator := f(ventilator), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, mdro := f(mdro), by = .(state, date, Fac_Name, Fac_Unit, res_id)]

# There are some other missing, so we will replace them by the individual (regardless)
# of the day
dat[, wound_care := f(wound_care), by = .(state, Fac_Name, Fac_Unit, res_id)]
dat[, mdro := f(mdro), by = .(state, Fac_Name, Fac_Unit, res_id)]

# # Still a couple of missings... just 0 for now
# dat[, wound_care := fcoalesce(wound_care, 0L)]
# dat[, mdro := fcoalesce(mdro, 0L)]
```

Now that we have idividual ids for the unit/location, we can start
creating the network data:

```{r}
# Edgelist
edges <- unique(dat[, .(net_id, hcw_id, res_id, unit_type)])

# Nodes dataset
nodes <- rbind(
  dat[, .(
    id = hcw_id,
    net_id, state, Fac_Name, Fac_Unit,
    white = 0L,
    race = "", res_age = 0,
    type = hcw_type, group = "hcw",
    is_actor = TRUE,
    # These are set to  be zero for the case of hwc
    wound_care = 0L,
    diabetes   = 0L,
    dialysis   = 0L,
    wheelchair = 0L,
    bedridden  = 0L,
    ventilator = 0L,
    mdro       = 0L #,
    # facility_t = unit_type
    )],
  dat[, .(
    id = res_id,
    net_id, state, Fac_Name, Fac_Unit,
    white,
    race, res_age,
    type = "resident", group = "residents",
    is_actor = FALSE,
    # But no for the residents
    wound_care, diabetes, dialysis, wheelchair, bedridden, ventilator, mdro #,
    # facility_t = unit_type
    )]
) |> unique()
```

The next steps will be to check for duplicates in the data, in particular, at the
node level.

## Duplicated records (hcw)

```{r check-dpls}
setorder(nodes, net_id, id)
nodes[, n := .N, by = .(net_id, id)][n > 1 & group == "hcw", .(id, type)]
# nodes[, n := .N, by = .(net_id, id)][n > 1 & group == "hcw", ] |> View()
```

We can see in the data that the duplicated records here (currently 12)
are in its majority if type "Nurse" or "CNA" (certified nurse
assistant.) We will replace those with "Nurse." We also see "RT" (respiratory
therapist) and "other".

```{r check-dpls-replace}
nodes[
  n > 1 & group == "hcw" & any(type %in% c("Nurse", "CNA")),
  type := "Nurse",
  by = .(net_id, id)
  ]
```

## Duplicated records (residents)

```{r check-dpls-residents}
nodes[n > 1 & group == "residents"][1:10 ,.(id, res_age)]
```

Most of the duplicates are in age. For now we will take the 50th quantile and use
that as the observed age. Notice that, while most cases the duplicates
have ages that show little differences, e.g., one or two years, there
are some cases with large age gaps, e.g., 20 years.

```{r check-dpls-residents-fix}
nodes[
  n > 1 & group == "residents",
  res_age := floor(quantile(res_age, na.rm = TRUE, probs = .5)),
  by = .(net_id, id)
  ]
```

## Checking no duplicates exists

After the data pre-processing, let's check whether the duplicates are
now gone

```{r}
nodes <- unique(nodes) # No change
nodes[, n := .N, by = .(net_id, id)]
table(nodes[, n])
```

Five observations that are duplicated twice each:

```{r}
nodes[n > 1, .(id, race, n)]
```

In this case, the observations have differences in race, e.g., the first
of them registers both "White" and "Black or African American." Since we
cannot correctly identify which is the correct, we will set them as "NK -
Not Known."

```{r}
nodes[n > 1, c("race", "white") := list("NK - Not Known", 2L)]

# Set the NK - Not Known to be zero
nodes[race == "NK - Not Known", white := NA_integer_]
nodes <- unique(nodes)
nodes[, n := NULL]

```

Observations with missing race or age will be dropped.

```{r}
nodes <- nodes[!is.na(white) & !is.na(res_age)]
```

Finally, we need to make sure that edges match the nodes

```{r}
edges <- merge(
  edges, nodes[, .(id, net_id, here_hcw = TRUE)],
  by.x = c("hcw_id", "net_id"), 
  by.y = c("id", "net_id"), all.x = TRUE, all.y = FALSE
)

edges <- merge(
  edges, nodes[, .(id, net_id, here_res = TRUE)],
  by.x = c("res_id", "net_id"), 
  by.y = c("id", "net_id"), all.x = TRUE, all.y = FALSE
)

edges[, here_hcw := fcoalesce(here_hcw, FALSE)]
edges[, here_res := fcoalesce(here_res, FALSE)]
edges <- edges[here_hcw & here_res]

edges[, c("here_hcw", "here_res") := c(NULL, NULL)]
edges <- unique(edges)
```

A couple of extra variables, for now, the square-root of age, which we will use
in the ERMG model.

```{r}
nodes[, sqrt_age := sqrt(res_age)]
```

# Building the networks

The following code of block will iterate through all network ids and create one
network per case. Right now we are counting 100 networks:

```{r build-nets}
# Making room to store the networks, we will keep two types of
# networks: HCW - Patient
networks <- vector("list", length(unique(nodes$net_id)))
names(networks) <- unique(nodes$net_id)
networks_proj <- networks

nerrors <- 0L
for (n in names(networks)) {
  
  # Selecting the edges and nodes relevant for the network
  edges_n <- edges[net_id == n, .(hcw_id, res_id, unit_type)]
  
  nodes_hcw_n <- nodes[net_id == n & group == "hcw"]
  nodes_res_n <- nodes[net_id == n & group == "residents"]
  
  # Trying to read the data
  networks[[n]] <- tryCatch({
    suppressWarnings(as.network(
      x         = as.data.frame(edges_n),
      vertices  = rbind(nodes_hcw_n, nodes_res_n),
      bipartite = TRUE,
      bipartite_col = "is_actor"
      ))
  }, error = function(e) e)
  
  # Notification!
  if (inherits(networks[[n]], what = "error")) {
    message("Network ", n, " failed to be built")
    nerrors <- nerrors + 1L
    stop()
    next
  }
  

}

# Reporting how many went OK
message("Done! ", nerrors, "/", length(networks), " failed.")
```

Here are four networks:

```{r viz}
# Example of a graph
set.seed(1231)

net_figs <- vector("list", 4L)
for (i in 1:4)
  net_figs[[i]] <- nplot(
    networks[[i]],
    edge.curvature=0,
    vertex.color = c(
      hcw = "blue", residents = "darkgray"
      )[networks[[i]] %v% "group"] |> adjustcolor(alpha.f = .5),
    vertex.nsides = c(hcw = 4, residents = 20)[networks[[i]] %v% "group"],
    vertex.rot    = 2 * pi / 8, vertex.size.range = c(.025, .05)
    )

# Putting them together using the gridExtra package
gridExtra::grid.arrange(grobs=net_figs, nrow = 2, ncol = 2) 
```

```{r}
set.seed(1331);nplot(
    networks[[1]], 
    edge.curvature=0,
    bg.col = "transparent",
    vertex.color = c(
      hcw = "blue", residents = "darkgray"
      )[networks[[i]] %v% "group"] |> adjustcolor(alpha.f = .9),
    vertex.nsides = c(hcw = 4, residents = 20)[networks[[i]] %v% "group"],
    vertex.rot    = 2 * pi / 8,
    vertex.size.range = c(.025, .05),
    vertex.label = "",
    edge.color = ~ego + alter,
    vertex.frame.color = "darkgray"
    ) |>
  netplot::nplot_legend(
    labels = c("healthcare worker", "Resident"),
    pch    = c(15, 19),
    packgrob.args = list(side="bottom"),
    ncol = 2,
    gp = gpar(
      fill = c("blue", "gray"),
      col = c("blue", "gray")
      )
    
  )
```


# Checking the networks

```{r checking-networks}
wrong_length <- sapply(networks, \(n) {
  length(unique(n %e% "unit_type")) != 1L
})

which(wrong_length)
```

# Final step: Saving the data

```{r save-data}
saveRDS(networks, "data/model_data_networks.rds")
```

