---
title: "Descriptive Analysis"
format: html
author: George G Vega Yon, Ph.D.
date: 2026-06-12
embed-resources: true
---

```{r}
source("models/helpers.R")
load('data/network93_f2.RData')
```

## Assessing the distribution of the resident features

```{r}
features <- lapply(seq_along(networks), function(i) {
  net <- networks[[i]]

  data.table(
    is_actor = net %v% "is_actor",
    net_id = i,
    diabetes = net %v% "diabetes",
    dialysis = net %v% "dialysis",
    wound_care = net %v% "wound_care",
    bedridden = net %v% "bedridden",
    ventilator = net %v% "ventilator",
    mdro = net %v% "mdro",
    wheelchair = net %v% "wheelchair",
    degree = sna::degree(net),
    state = net %v% "state"
  )
}) |> rbindlist()

# Excluding healtcare providers
features_actor <- features[is_actor == FALSE]

# Checking if individuals will have more than one condition
features_actor[, n_conditions := 
  (diabetes == "Yes") +
  (dialysis == "Yes") +
  (wound_care == "Yes") +
  (bedridden == "Yes") +
  (ventilator == "Yes") +
  (mdro == "Yes") +
  (wheelchair == "Yes")
  ]

features_actor[, table(n_conditions)] |>
  knitr::kable()
```

So individuals may have up to 6 different conditions. We can also visualize this in a single figure showing the distribution of patients with each condition using ggridges. Looking into the degree distribution by type of node.

```{r}
#| label: degree-distribution-by-node-type
#| fig-height: 10
#| fig-width: 6
features[, n := .N, by = net_id]
features[, net_id := factor(net_id, levels = unique(net_id))]
setorder(features, state, -n)
features |>
  ggplot(aes(x = degree, fill = is_actor, y = net_id)) +
  geom_density_ridges() +
  facet_wrap(~state, scales = "free_y") +
  labs(
    x = "Degree",
    y = "Network ID"
  )
  # Removing the y-axis ticks
  theme(axis.ticks.y = element_blank())
```

And overall

```{r}
#| label: degree-distribution-overall
features |>
  ggplot(aes(x = degree, y = state)) +
  geom_density_ridges(
    jittered_points = TRUE, position = "raincloud",
    alpha = 0.9, scale = 0.9
  ) +
  theme_minimal()
```

Checking if the degree distribution changes across different states.

```{r}
#| label: degree-distribution-by-state
features[, n_hcp := sum(is_actor == TRUE), by = net_id]
features[, n_patients := sum(is_actor == FALSE), by = net_id]
lm(
  degree ~
    n_hcp + n_patients*state + is_actor +
    I(n_patients ^ 2),
  data = features
) |>
  texreg::screenreg(
    single.row = TRUE, 
    custom.model.names = "Degree"
    )
```

```{r}
library(ggridges)

features_summary <- features[,
  .(
    Diabetes = sum(diabetes == "Yes"),
    Dialysis = sum(dialysis == "Yes"),
    `Wound Care` = sum(wound_care == "Yes"),
    Bedridden = sum(bedridden == "Yes"),
    Ventilator = sum(ventilator == "Yes"),
    MDRO = sum(mdro == "Yes"),
    Wheelchair = sum(wheelchair == "Yes")
  ),
  by = net_id
]

# Reshaping the data to long (from wide)
features_long <- data.table::melt(features_summary, id.vars = "net_id", measure.vars = c("Diabetes", "Dialysis", "Wound Care", "Bedridden", "Ventilator", "MDRO", "Wheelchair"), variable.name = "Condition", value.name = "Count")

features_long |>
  ggplot(aes(x = Count, y = Condition)) +
  geom_density_ridges(
    alpha = 0.7,
    jittered_points = TRUE, position = "raincloud"
    ) +
  labs(
    x = "Number of Patients",
    y = "Condition"
    ) +
  theme_minimal()
```

Repeating the same analysis for each type of healthcare professional


```{r}
features <- lapply(seq_along(networks), function(i) {
  net <- set_ergm(networks[i])$nets[[1]]

  data.table(
    is_actor = net %v% "is_actor",
    net_id = i,
    type = net %v% "type"
  )
}) |> rbindlist()

# Excluding healtcare providers
features <- features[is_actor == TRUE]

features[, table(type)] |>
  knitr::kable()
```

Same type of visual to show the distribution of each type of healthcare professional across the networks.

```{r}
features[, .(Count = .N), by = .(net_id, type)] |>
  ggplot(aes(x = Count, y = type)) +
  geom_density_ridges(
    alpha = 0.7,
    jittered_points = TRUE, position = "raincloud"
    ) +
  labs(
    x = "Number of HCPs",
    y = "Type of HCP"
  ) +
  theme_minimal()
```

