---
title: "Individual ERGMs"
format: html
---


```{r}
#| label: setup
source("models/helpers.R")
load('data/network93_f2.RData')

# When this flag is TRUE, the document will
# avoid trying to fit the models that are not available
# in the "models/02-individual_ergms_fits" folder.
skip_missing <- TRUE
```


```{r}
#| label: setup-cluster
# Setting up the cluster object
library(parallel)
cl <- makeForkCluster(40)
invisible(clusterEvalQ(cl, {
  load('data/network93_f2.RData')
  source("models/helpers.R")
  NULL
}))
```

## Partial model

```{r}
#| label: run-partial-models
set.seed(as.integer(Sys.time()))
indices <- sample(seq_along(networks))

results_partial <- parLapplyLB(cl, indices, \(i, skip_missing) {
  
  # Building the file path and figuring out
  # if the model has been fitted already
  fn <- sprintf(
    "models/02-individual_ergms_fits/%02d_partial.rds",
    i
  )

  if (file.exists(fn))
    return(readRDS(fn))

  if (skip_missing)
    return(NULL)

  # Preparing the data for the model
  default_control <- control.ergm(seed = 1233)

  dat <- set_ergm(nets = networks[i])

  model <- with(dat, {
    model <- Networks(nets) ~ edges + offset(edges)
    update(model, . ~ . + offset(b1degree(0)) + offset(b2degree(0)))
  })

  ans <- tryCatch({
    with(dat, ergm(model,
      offset.coef = c(-log(n_hcp), -Inf, -Inf),
      control = default_control
    ))
  }, error = \(e) e)

  # Returns invisibly
  saveRDS(ans, fn)

}, skip_missing = skip_missing)
```

We now assess the individual model fits, looking at the variability of the parameter estimates

```{r}
#| label: check-partial-results
res_partial <- lapply(indices, \(i) {
  fn <- sprintf(
    "models/02-individual_ergms_fits/%02d_partial.rds",
    i
  )
  if (!file.exists(fn))
    return(NULL)

  x <- readRDS(fn)

  # Extracting information about the state
  state <- unique(x$network %v% "state")

  data.frame(
    id    = i,
    coef  = names(coef(x)),
    value = coef(x),
    sd    = sqrt(diag(vcov(x))),
    state = state,
    n_hcp = sum(x$network %v% "is_actor"),
    n_pat = sum(!x$network %v% "is_actor"),
    n     = length(x$network %v% "is_actor"),
    facility = paste(
      unique(x$network %v% "Fac_Name"), "-",
      unique(x$network %v% "Fac_Unit")
    )
  )
}) |> rbindlist()

# Removing the offset
res_partial <- res_partial[!grepl("offset", coef)]
```

We now visualize the distribution of the parameter estimates across the models

```{r}
#| label: visualize-partial-results
library(ggplot2)
res_partial <- setorder(res_partial, state, facility, n, value)
res_partial[, id := as.factor(.I)]

set.seed(12)
res_partial[, col := sample(colors())[as.integer(as.factor(facility))]]

ggplot(res_partial, aes(x = id, y = value)) +
  # Adding a convex hull by `facility`
  geom_polygon(
    aes(group = facility, colour = col),
    linewidth = 2,
    alpha = .5,
    fill = "transparent",
    # colour = "steelblue"
    show.legend = FALSE
  ) +
  geom_point(aes(size = n, colour = col)) +
  # 95% CI bars based on the sd variable
  geom_errorbar(aes(ymin = value - 1.96 * sd, ymax = value + 1.96 * sd), width = 0.2) +
  facet_wrap(~state, scales = "free_x") +
  # Dropping the x-axis since it is uninformative
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    # Also the x-axis label
    axis.title.x = element_blank()
    ) +
  scale_colour_identity() +
  # Dropping the colour legend
  guides(colour = "none")  +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", size = 1) +
  labs(
    y = "Edges Parameter Estimate"
  )
```

We can assess the stabililty of the parameter estimate by facility looking at the confidence intervals. For any given facility, we can test whether the confidence intervals of the parameter estimates overlap each other:

```{r}
#| label: check-overlap
overlap <- NULL
for (fac in unique(res_partial$facility)) {

  # Subsetting facility
  res_fac <- res_partial[facility == fac]

  # Checking if there are at least 2 rows
  if (nrow(res_fac) < 2)
    next

  for (i in 2:nrow(res_fac)) {

    # Generating the 95% confidence intervals for i
    ci_i <- c(res_fac[i, "value"] - 1.96 * res_fac[i, "sd"], res_fac[i, "value"] + 1.96 * res_fac[i, "sd"]) |> unlist()

    for (j in 1:(i - 1)) {

      # Generating the 95% confidence intervals for j
      ci_j <- c(res_fac[j, "value"] - 1.96 * res_fac[j, "sd"], res_fac[j, "value"] + 1.96 * res_fac[j, "sd"]) |> unlist()

      # Checking if the confidence intervals overlap
      if (ci_i[1] < ci_j[2] && ci_i[2] > ci_j[1]) {
        overlap <- rbind(overlap, data.frame(
          state = res_fac[i, "state"],
          facility = fac, i = i, j = j, overlap = TRUE)
          )
      }
      else {
        overlap <- rbind(overlap, data.frame(
          state = res_fac[i, "state"],
          facility = fac, i = i, j = j, overlap = FALSE)
          )
      }

    }

  }
}

overlap <- data.table(overlap)
# Tabulating, by state, how CI overlap
overlap_summary <- overlap[, .(
  n_overlap = sum(overlap),
  prop = sum(overlap) / .N,
  n_total = .N
  ), by = .(state)]

setorder(overlap_summary, prop)

knitr::kable(overlap_summary)
```



## Full model

```{r}
#| label: run-models
# Randomly fit the models in parallel
# Flushing things out
if (exists("cl")) {
  parallel::stopCluster(cl)
}
cl <- parallel::makeForkCluster(40)
invisible(parallel::clusterEvalQ(cl, {
  load('data/network93_f2.RData')
  source("models/helpers.R")
  NULL
}))

set.seed(as.integer(Sys.time()))
indices <- sample(seq_along(networks))

results <- parallel::parLapplyLB(cl, indices, \(i, skip_missing) {
  
  # Building the file path and figuring out
  # if the model has been fitted already
  fn <- sprintf(
    "models/02-individual_ergms_fits/%02d_full.rds",
    i
  )

  if (file.exists(fn))
    return(readRDS(fn))

  if (skip_missing)
    return(NULL)

  # Preparing the data for the model
  default_control <- control.ergm(seed = 1233)

  dat <- set_ergm(nets = networks[i])

  types <- summary(Networks(dat$nets) ~ b1factor("type", levels = 1:10)) |>
    names() |>
    gsub(pattern = ".+\\.type\\.", replacement = "")

  types <- types[types != "NA"]

  model <- with(dat, {
    model <- Networks(nets) ~ edges + offset(edges)

    # Checking which will be the base. By priority
    if ("Other" %in% types) {
      id_base <-which(types == "Other")
    } else if ("RT" %in% types) {
      id_base <-which(types == "RT")
    } else if ("PT/OT" %in% types) {
      id_base <-which(types == "PT/OT")
    } else if ("CNA" %in% types) {
      id_base <-which(types == "CNA")
    } else {
      id_base <- 1
    }
    model <- update(model, . ~ . + N(~b1factor("type", levels = -id_base)) )

    print(id_base)

    
    # if (length(dat$Nurse))
    #   model <- update(model, . ~ . + N(~b1factor("Nurse")))
    
    # if (length(dat$CNA))
    #   model <- update(model, . ~ . + N(~b1factor("CNA")))

    # if (length(dat$PTOTRT))
    #   model <- update(model, . ~ . + N(~b1factor("PTOTRT")))

    if (length(dat$dialysis))
      model <- update(model, . ~ . + N(~b2factor('dialysis')))

    if (length(dat$wound_care))
      model <- update(model, . ~ . + N(~b2factor('wound_care')))

    if (length(dat$bedridden))
      model <- update(model, . ~ . + N(~b2factor('bedridden')))

    if (length(dat$ventilator))
      model <- update(model, . ~ . + N(~b2factor('ventilator')))

    if (length(dat$mdro))
      model <- update(model, . ~ . + N(~b2factor('mdro')))

    if (length(dat$wheelchair))
      model <- update(model, . ~ . + N(~b2factor('wheelchair')))

    update(model, . ~ . + offset(b1degree(0)) + offset(b2degree(0)))
  })

  ans <- tryCatch({
    with(dat, ergm(model,
      offset.coef = c(-log(n_hcp), -Inf, -Inf),
      control = default_control
    ))
  }, error = \(e) e)

  # Returns invisibly
  saveRDS(ans, fn)

}, skip_missing = skip_missing)

parallel::stopCluster(cl)
```

## Fitting models with issues

Verifying what proportion has been fitted, and which models failed to fit

```{r}
#| label: check-results
saved_models <- list.files(
  "models/02-individual_ergms_fits/",
  full.names = TRUE,
  pattern = ".*_full\\.rds"
) 

# Getting the indices to see which ones are missing
saved_indices <- as.integer(gsub(".*\\/([0-9]+)_full\\.rds", "\\1", saved_models))

# Identifying the missing indices
missing_indices <- setdiff(seq_along(networks), saved_indices)
```

Visualizing the results of the networks we couldn't fit

```{r}
library(netplot)
grobs <- lapply(missing_indices, function(i) {
  netplot::nplot(networks[[i]], vertex.col = ~type)
})

gridExtra::grid.arrange(grobs = grobs, nrow = 5, ncol=3)
```

We also need to identify to which state each of the missing networks belongs.

```{r}
missing_states <- sapply(missing_indices, \(i) {
  unique(networks[[i]] %v% "state")
})

data.frame(
  State = missing_states,
  Index = missing_indices
) |> knitr::kable()
```

### Model 2

In the case of model 2, there is a single nurse in the data, so, it is hard to capture the ERGM struggles converging the model (apparently), furthermore, there's also only CNA and a single nurse, so, ultimately struggles with this unevenly distributed roles. Refitting the model with at least 2 of each kind helps.

```{r}
#| label: refit-model-2
if (2 %in% missing_indices) {
  dat <- set_ergm(nets = networks[2])
  ans_2 <- with(dat,
    ergm(
      Networks(nets) ~
        edges +
        N(~b2factor("dialysis")) +
        N(~b2factor("wound_care")) + N(~b2factor("bedridden")) + 
        offset(edges) + offset(b1degree(0)) + offset(b2degree(0))
      ,
      control = control.ergm(seed = 1233),
      offset.coef = c(-log(n_hcp), -Inf, -Inf)
    )
  )

  saveRDS(ans_2, "models/02-individual_ergms_fits/02_full.rds")
}
```

### Model 12

In the case of model 12, there is colinearity between edges and the type of b1factor terms

```{r}
#| label: refit-model-12
if (12 %in% missing_indices) {
  dat <- set_ergm(nets = networks[12])

  ans_12 <- with(dat,
    ergm(
      Networks(nets) ~
        edges + 
        N(~b1factor("type", levels = -3)) +
        # N(~b2factor("dialysis")) +
        N(~b2factor("wound_care")) + 
        N(~b2factor("bedridden")) +
        offset(edges) + offset(b1degree(0)) + offset(b2degree(0)),
      control = control.ergm(seed = 1233),
      offset.coef = c(-log(n_hcp), -Inf, -Inf)
    )
  )

  saveRDS(ans_12, "models/02-individual_ergms_fits/12_full.rds")
}
```

### Model 13

For model 13, dialisis is too close to the boundaries, with b2factor.dialysis.Yes = 1

```{r}
#| label: refit-model-13
if (13 %in% missing_indices) {
  dat <- set_ergm(nets = networks[13])

  ans_13 <- with(dat,
    ergm(
      Networks(nets) ~
        edges + N(~b1factor("type", levels = -3)) + 
        N(~b2factor("wound_care")) + N(~b2factor("bedridden")) + 
        offset(edges) + offset(b1degree(0)) + offset(b2degree(0)),
      control = control.ergm(seed = 1233),
      offset.coef = c(-log(n_hcp), -Inf, -Inf)
    )
  )
  saveRDS(ans_13, "models/02-individual_ergms_fits/13_full.rds")
}
```

### Model 17


```{r}
#| label: refit-model-17
if (17 %in% missing_indices) {
  dat <- set_ergm(nets = networks[17])

  ans_17 <- with(dat,
    ergm(
      Networks(nets) ~ edges + N(~b1factor("type", levels = -3)) + 
    N(~b2factor("wound_care")) + N(~b2factor("bedridden")) + 
    offset(edges) + offset(b1degree(0)) + offset(b2degree(0)),
      control = control.ergm(seed = 1233),
      offset.coef = c(-log(n_hcp), -Inf, -Inf)
    )
  )
  saveRDS(ans_17, "models/02-individual_ergms_fits/17_full.rds")
}
```

### Model 33

Dialisis was also the issue here

```{r}
#| label: refit-model-33
if (33 %in% missing_indices) {
  dat <- set_ergm(nets = networks[33])

  ans_33 <- with(dat,
    ergm(
      Networks(nets) ~ edges + N(~b1factor("type", levels = -3)) + 
      N(~b2factor("wound_care")) + N(~b2factor("bedridden")) + 
      offset(edges) + offset(b1degree(0)) + offset(b2degree(0)),
      control = control.ergm(seed = 1233),
      offset.coef = c(-log(n_hcp), -Inf, -Inf)
    )
  )
  saveRDS(ans_33, "models/02-individual_ergms_fits/33_full.rds")
}
```


### Model 35

Dialysis was equal to 2, so the model was failing (again)

```{r}
#| label: refit-model-35
if (35 %in% missing_indices) {
  dat <- set_ergm(nets = networks[35])

  ans_35 <- with(dat,
    ergm(
      Networks(nets) ~ edges + N(~b1factor("type", levels = -4)) + # N(~b2factor("dialysis")) + 
      N(~b2factor("wound_care")) + N(~b2factor("bedridden")) + 
      offset(edges) + offset(b1degree(0)) + offset(b2degree(0)),
      control = control.ergm(seed = 1233),
      offset.coef = c(-log(n_hcp), -Inf, -Inf)
    )
  )
  saveRDS(ans_35, "models/02-individual_ergms_fits/35_full.rds")
}
```

### Model 36

Here the issue was b1factor.type.PT/OT (1 only) and bedridden (1 only).

```{r}
#| label: refit-model-36
if (36 %in% missing_indices) {
  dat <- set_ergm(nets = networks[36])

  ans_36 <- with(dat,
    ergm(
      Networks(nets) ~ edges + b1factor("type", levels = -c(3, 4)) +
      N(~b2factor("dialysis")) + 
      N(~b2factor("wound_care")) + # N(~b2factor("bedridden")) + 
      offset(edges) + offset(b1degree(0)) + offset(b2degree(0)),
      control = control.ergm(seed = 1233),
      offset.coef = c(-log(n_hcp), -Inf, -Inf)
    )
  )
  saveRDS(ans_36, "models/02-individual_ergms_fits/36_full.rds")
}
```

### Model 38

We just kept CNA and Nurse, the other two levels were too small

```{r}
#| label: refit-model-38
if (38 %in% missing_indices) {
  dat <- set_ergm(nets = networks[38])

  ans_38 <- with(dat,
    ergm(
      Networks(nets) ~ edges +
      b1factor("type", levels = c(1, 2)) +
      N(~b2factor("wound_care")) + 
      N(~b2factor("bedridden")) + offset(edges) + offset(b1degree(0)) + 
      offset(b2degree(0)),
      control = control.ergm(seed = 1233),
      offset.coef = c(-log(n_hcp), -Inf, -Inf)
    )
  )
  saveRDS(ans_38, "models/02-individual_ergms_fits/38_full.rds")
}
```

### Model 83

Dialisis has a single case. Also, we are using Nurse as the reference category.

```{r}
#| label: refit-model-83
if (83 %in% missing_indices) {
  dat <- set_ergm(nets = networks[83])

  ans_83 <- with(dat,
    ergm(
      Networks(nets) ~ edges + N(~b1factor("type", levels = 2)) + #  N(~b2factor("dialysis")) + 
      N(~b2factor("wound_care")) + offset(edges) + offset(b1degree(0)) + 
      offset(b2degree(0)),
      control = control.ergm(seed = 1233),
      offset.coef = c(-log(n_hcp), -Inf, -Inf)
    )
  )
  saveRDS(ans_83, "models/02-individual_ergms_fits/83_full.rds")
}
```

## Taking a look at potential errors

```{r}
error_fits <- NULL
for (s in saved_indices) {
  fn <- sprintf("models/02-individual_ergms_fits/%02i_full.rds", s)
  if (inherits(readRDS(fn), "error"))
    error_fits <- c(error_fits, s)
}
```

Only error is 39

### Model 39

```{r}
#| label: refit-model-39
if (39 %in% error_fits) {
  dat <- set_ergm(nets = networks[39])

  ans_39 <- with(dat,
    ergm(
      Networks(nets) ~ edges + N(~b1factor("type", levels = -c(3, 5))) + 
      N(~b2factor("wound_care")) +
      N(~b2factor("bedridden")) + 
    offset(edges) + offset(b1degree(0)) + offset(b2degree(0)),
      control = control.ergm(seed = 1233),
      offset.coef = c(-log(n_hcp), -Inf, -Inf)
    )
  )
  saveRDS(ans_39, "models/02-individual_ergms_fits/39_full.rds")
}
```


## Analyzing all fits


```{r}
fits_coefs <- lapply(seq_along(networks), \(i) {

  out <- if (i %in% saved_indices) {
    readRDS(sprintf("models/02-individual_ergms_fits/%02i_full.rds", i))
  } else if (exists(paste0("ans_", i))) {
    get(paste0("ans_", i))
  } else {
    return(NULL)
  }

  if (inherits(out, "error")) {
    return(NULL)
  }

  ans <- summary(out)$coefficients

  cbind(data.frame(
    id = i,
    coef = rownames(ans)
  ), as.data.frame(ans))

}) |> rbindlist()

# Excluding the offset coefficients
fits_coefs <- fits_coefs[!grepl("offset", fits_coefs$coef), ]

# Correcting names (removing the N(1)~ prefix)
fits_coefs$coef <- gsub("N\\(1\\)~", "", fits_coefs$coef)
```

We now can visualize using ggridges the distribution of the coefficients

```{r}
library(ggridges)

fits_coefs[
  abs(Estimate) <= 5 &
  !grepl("Other", coef)
  ] |>
ggplot(aes(x = Estimate, y = coef)) +
  geom_density_ridges(
    alpha = 0.5,
    jittered_points = TRUE, position = "raincloud"
    ) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(
    # Adding a note regarding the ones excluded (x out of y)
    caption = sprintf("Excluded %i out of %i coefficients with |Estimate| > 5", sum(abs(fits_coefs$Estimate) > 5), nrow(fits_coefs))
  )
```

Instead of looking at the distribution, we can assess how often the effects were positive or negative (significant and not):


```{r}
fits_coefs[, .(
  `Positive` = mean(Estimate > 0, na.rm = TRUE),
  `Negative` = mean(Estimate < 0, na.rm = TRUE),
  `Positive (*)` = mean(Estimate > 0 & `Pr(>|z|)` < 0.05, na.rm = TRUE),
  `Negative (*)` = mean(Estimate < 0 & `Pr(>|z|)` < 0.05, na.rm = TRUE),
  count    = .N
), by = coef] |>
  setorder(-`Positive (*)`) |>
  knitr::kable(digits = 2)
```

Compared to other categories, CNAs and Nurses both tend to have a significantly higher number of ties compared to other care professionals, including therapists, physicians, and other healthcare providers. The oposite is true for PT and OTs who have a significantly lower tendency to form ties.

On the other hand, bedridden and wound care patients tend to have a significantly higher number of ties compared to other patient populations, but the significance is less consistent (fewer than 6 and 5 % of the time).

This is the most sophisticated analysis we have conducted so far. Adding mixing patterns at the individual network level is cumbersome. Instead, we will focus on using pooled models to analyze the data.

