---
title: "Bipartite networks of health care workers and residents in long-term care facilities"
format: html
editor_options: 
  markdown: 
    wrap: 80
date: 2023-08-26
author:
  - Chong Zhang
  - George G. Vega Yon
date-modified: 2026-06-16
embed-resources: true
---
## Overview

Recent updates:

1. We removed 6 networks with only 1 HCW - given the bipartite and no-isolate constraints they are fixed and no variation.

2. For ergm terms present in only a subset of networks, we now restricted their evaluation in only the subset of networks. For starmix(2),  at least two such nodes should be present for the network to be included.

3. There are only 3 physicians and 1 PA/NP in all the networks, therefore it is not very meaningful to evaluate the hemophily in these two types of HCWs. We grouped physician with 'Other', and PA/NP with 'Nurse'.

4. We detected one outlier, and removed it when fitting models to speed up the convergence, and for the final model selected, we added it back in.   

# Statistical methods

We have interaction data between HCWs and residents from a sample of 100 networks. ERGMs were used to understand network structure and its associations with nodal attribues. Given the fact that only interactions can be recorded in the data, we used constriction to forbid isolates. As a result, we excluded 7 networks with only 1 HCWS, because by the definitions of bipartite + no isolates, there are no variations in these networks given the set of nodes (saturated?), leaving 93 networks. Because our data is a sample of multiple networks, the N() ergm term in package ergm.mulit{} was used to evaluate each term jointly across multiple networks. ERGM terms we examined included edges and its interaction with network size, b2start(2), b2concurrent (residents having more than 1 HCWs), gwb1dsp (geometrically weighted shared partners between HCWs), HCW type (Nurse, CNA, PT/OT/RT, Other), and resident care needs (dialysis, wound_care, bedridden, ventilator, diabetes, wheelchair, mdro). Terms are added one at a time to ensure convergence (assessed using trace plot) and reasonable goodness of fit (assessed using standardized pearson's residuals using function gofN()). For highly correlated terms (b2star, b2condurrent, gwb1dsp), we kept the one that gives the smallest AIC/BIC and dropped the rest. As the model got more complex and the algorithms were slow to converge, we removed one outlier to speed up the convergence. The largest model with all terms being significant was selected as the best model. We then used the estimated coefficients from this model as initial values to fit the model on all 93 networks, and compared results before and after adding back the outlier.



```{r}
#| label: packages
#| message: false
#| warning: false
source("models/helpers.R")
load('data/network93_f2.RData') # a list of 93 networks to fit ERGMs
```


# Fitting ERGMs

## Model 0: edges only, with scaling by log(n) and state level effect

```{r}
#| label: prepare-data
model_data <- set_ergm(nets = networks)
```

We will also be applying scaling based on the number of actors/entities in the network. We calculate the number of each using the following code:

Common controls for ergm models: we use Stochastic approximation since it is fast. 
```{r}
#| label: controls
default_control_mcmle <- control.ergm(seed = 55551)

default_control_sa <- control.ergm(
  seed            = 44485,
  main.method     = "Stochastic-Approximation", # Robbins-Monro Stochastic Approx.
  SA.nsubphases   = 4 * 1, # 1 the default
  SA.burnin       = 1024 * 4, # Twice the default
  SA.interval     = 1024 * 4 # four times... This is being passed to SA instead
)
```

Model 0: edges only.
Since we do not have isolates in our network by design, we restrict the network by adding offset terms b1degree(0) and b2degree(0) with coefficients -Inf.
```{r}
#| label: model0
#| message: false
#| warning: false
#| results: hide
with_restore("res0", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~log(n_tot) + I(log(n_tot)^2))+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res0_state", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~log(n_tot) + facility_state)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res0_state2", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~log(n_tot) + I(log(n_tot)^2) + facility_state)+
    offset(b1degree(0))+offset(b2degree(0)),
    offset.coef = c(-Inf, -Inf),
    control = default_control_mcmle
))

with_restore("res0_state_centered", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~I(log(n_tot) - mean(log(n_tot))) + facility_state)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

texreg::screenreg(
  list(
    res0,
    res0_state,
    res0_state2,
    res0_state_centered
  ),
  single.row = TRUE
)
```

Model 4, the mean-centered, is the one with best fit overall. Adding the square term for log(n) did not improve the fit, but adding the state level effect did. Let's investiage the convergence:

```{r}
#| label: model0-convergence
#| eval: false
mcmc.diagnostics(res0_state_centered)
```

The model does a good job in convergence: the MCMC trace shows good mixing, the omnibus test is not significant, and the autocorrelation reduces decently fater the third lag. We will be using this model as the baseline for the subsequent models.

```{r}
best_model <- res0_state_centered

with_restore("res0_hcp", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~I(log(n_hcp) - mean(log(n_hcp))) + facility_state)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res0_patients", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~I(log(n_patients) - mean(log(n_patients))) + facility_state) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res0_patients_no_state", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~I(log(n_hcp) - mean(log(n_hcp)))) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res0_patients_no_state2", model_data, ergm(
  Networks(nets) ~
    N(
      ~edges,
      ~I(log(n_hcp) - mean(log(n_hcp))) + I(log(n_hcp)^2 - mean(log(n_hcp)^2))
      ) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

texreg::screenreg(
  list(
    best_model, res0_hcp, res0_patients, res0_patients_no_state,
    res0_patients_no_state2
    ),
  single.row = TRUE
)
```

The Model 2 (correcting for the log(n_hcp)) has the best fit overall. This is inline with what we expect in theory. The degree distribution of HCPs is more likely to be affected by the network size than that of patients, because as the number of patients increases, the degree of HCPs increases, but the degree of patients does not necessarily increase as much. Thus, adjusting for log(n_hcp) is more appropriate in this application. We will be using log(n_hcp) for the rest of the models. Let's investiagete convergence for this model:

```{r}
#| label: model0_hcp-convergence
#| eval: false
mcmc.diagnostics(res0_hcp)
```

The omnibus test is not significant, the MCMC trace shows good mixing, and the autocorrelation reduces decently after the third lag. We will be using this model as the baseline for the subsequent models. **Nevertheless, the individual fits are worse off, particularly for the OR facilities**; let's see if we can adjust for both simultaneously:

```{r}
with_restore("res0_both", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_hcp) - mean(log(n_hcp))) +
      I(log(n_tot) - mean(log(n_tot))) +
      + facility_state
      )+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

texreg::screenreg(
  list(best_model, res0_hcp, res0_patients, res0_both),
  single.row = TRUE
)
```

At first look, it seems this approach performs significantly better. Let's look at the convergence diagnostics:

```{r}
#| label: model0_both-convergence
#| eval: false
mcmc.diagnostics(res0_both)
```

Looking at the diagnostics, this model did not converge well. The omnibus test is significant, so we are worse off than the previous model. Although scaling by the number of healthcare providers seemed like the right theoretical approach, the model is overall performing worse than just scaling by the number of nodes. We will be using the model with log(n) as the baseline for the subsequent models.

## Controling for type of HCP

```{r}
#| label: model-hcp-type
best_model <- res0_hcp

with_restore("res1_hcp_type", model_data, ergm(
  Networks(nets) ~
    N(
      ~edges,
      ~ 
        I(log(n) - mean(log(n))) +
        I(log(n_hcp) - mean(log(n_hcp))) +
        facility_state)+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res1_hcp_type2", model_data, ergm(
  Networks(nets) ~
    N(
      ~edges,
      ~ 
        I(log(n_hcp) - mean(log(n_hcp))) +
        facility_state)+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

texreg::screenreg(
  list(
    best_model,
    res0_both,
    res1_hcp_type,
    res1_hcp_type2
  ),
  single.row = TRUE
)
```

Let's check the convergence diagnostics:

```{r}
#| label: model-hcp-type-convergence
#| eval: false
mcmc.diagnostics(res1_hcp_type)
```

Since the overall fit is good but the omnibus test is failing, let's see if adding more structure changes things. Comparing the model in which we had both log(n) and log(n_hcp) with the model that only has log(n_hcp): The model featuring both variables has a better fit with better omnibus test results. So we will use that for the future.

## Investigating the effect of gwb1dsp

Since we expect the healthcare providers been somewhat evenly spread across residents, we expect to see a tendency of negative GWDSP effect for the HCP. Let's start with a basic model. Since the process generally is slow to converge, we will instead use the Stochastic Approximation method for now on:

Add b2 2-star (second level 2-stars) term:
```{r}
#| label: model1
#| message: false
#| warning: false
#| results: hide
with_restore("res1_gwb1dsp", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~ I(log(n_hcp) - mean(log(n_hcp))) + facility_state)+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 40) + 
    +offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res1_gwb1dsp2", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n) - mean(log(n))) +
      I(log(n_hcp) - mean(log(n_hcp))) +
      facility_state
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 40) + 
    +offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

tabulator(
  list(
    best_model,
    res1_hcp_type,
    res1_gwb1dsp,
    res1_gwb1dsp2
  ),
  single.row = TRUE
)
```

Although the AIC and BIC suggest that the new model is better, we need to check the convergence and goodness of fit before making a conclusion.

```{r}
#| label: model1_mcmle_convergence
#| eval: false
mcmc.diagnostics(res1_gwb1dsp)
```

The omnibus test is not significant, and the same applies to the MCMC joint P-value. The new model behaves significantly better than the previous, even improving in the omnibus test. The model that only includes the log(n_hcp) term performs better than the model with both log(n) and log(n_hcp) (omnibus test).

Using Stewart, Schweinberger, Bojanowski, and Morris (2019) (as suggested by the reviewer), I estimated the decay parameter directly. I was not able to directly estimate it with the interaction effect as ERGM failed with an error `objfun returned gradient not having all elements finite`.

## Looking into other terms

Now add b2factor and b2starmix (just Nurse for now)
```{r}
#| label: model5
#| eval: true
#| message: false
#| warning: false
#| results: hide
best_model <- res1_gwb1dsp

with_restore("res2_b2factor_terms", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~ I(log(n_hcp) - mean(log(n_hcp))) + facility_state)+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 40) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2factor('diabetes') +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res2_b2factor_terms2", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n) - mean(log(n))) +
      I(log(n_hcp) - mean(log(n_hcp))) +
      facility_state
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 40) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2factor('diabetes') +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res2_b2factor_terms2_logpat", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_patients) - mean(log(n_patients))) +
      I(log(n_hcp) - mean(log(n_hcp))) +
      facility_state
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 40) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2factor('diabetes') +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res2_b2factor_terms2_logpat_uncetered", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_patients)) +
      I(log(n_hcp)) +
      facility_state
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 40) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2factor('diabetes') +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))


with_restore("res2_b2factor_terms_nogwdsp", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~ I(log(n_hcp) - mean(log(n_hcp))) + facility_state)+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2factor('diabetes') +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))



tabulator(
  list(
    best_model,
    res2_b2factor_terms_nogwdsp,
    res2_b2factor_terms,
    res2_b2factor_terms2,
    res2_b2factor_terms2_logpat,
    res2_b2factor_terms2_logpat_uncetered
  ),
  single.row = TRUE
)
```

Comparing the centered with the uncentered models, the centered model converged quicker (120,553,472 vs 207,552,512) and had better convergence properties (Omni test of 1e-8 and Geweke test joint 0.54 vs Omni test of 2e-35 Geweke test joint 0.22). It is important to note that both models yield roughly the same results in terms of estimated coefficients. This is mostly an MCMC artifact used to improve the mixing of the Markov chain.

The model with both log(n) and log(n_hcp) terms performs worse if we look at the omnibus test. The overall fit drops, but, perhaps, we can simply run it for a bit longer (both) and assess if we can improve the fit:

```{r}
with_restore("res2_b2factor_terms_bis", model_data, ergm(
  formula(res2_b2factor_terms),
  control = control.ergm(
    seed = 3313,
    init = coef(res2_b2factor_terms),
    MCMLE.confidence = .995
  )
))

with_restore("res2_b2factor_terms2_bis", model_data, ergm(
  formula(res2_b2factor_terms2),
  control = control.ergm(
    seed = 3313,
    init = coef(res2_b2factor_terms2),
    MCMLE.confidence = .995
  )
))

texreg::screenreg(
  list(
    res2_b2factor_terms,
    res2_b2factor_terms_bis,
    res2_b2factor_terms2,
    res2_b2factor_terms2_bis
  ),
  single.row = TRUE
)
```

In this case, `res2_b2factor_terms_bis` does a better job in the omnibus test but it still fails to reject the null hypothesis. At the same time, `res2_b2factor_terms2_bis` does a worse job. The overall coefficients do not change much between the two runs.

```{r}
with_restore("gofN_res2_b2factor_terms", list(), {
  gofN(res2_b2factor_terms)
})
if (FALSE && interactive()) {
  pdf("gofN_res2_b2factor_terms.pdf")
  op <- par(mfrow = c(3, 2))
  plot(gofN_res2_b2factor_terms)
  par(op)
  dev.off()
}
```

## Fitting the model using the b2starmix terms

```{r}
best_model<- res2_b2factor_terms2_logpat

with_restore("res3_b2starmix_terms", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_patients) - mean(log(n_patients))) +
      I(log(n_hcp) - mean(log(n_hcp))) +
      facility_state
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    # gwb1dsp(fixed = FALSE, cutoff = 40) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b1starmix(2, 'diabetes') +
    b2starmix(2, 'type', base = c(3)) +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res3_b2starmix_terms_gwdsp", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_patients) - mean(log(n_patients))) +
      I(log(n_hcp) - mean(log(n_hcp))) +
      facility_state
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 100) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b1starmix(2, 'diabetes') +
    b2starmix(2, 'type', base = c(3)) +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res3_bnodematch_terms", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_patients) - mean(log(n_patients))) +
      I(log(n_hcp) - mean(log(n_hcp))) +
      facility_state
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    # gwb1dsp(fixed = FALSE, cutoff = 40) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2nodematch('diabetes', alpha = .25, diff = TRUE) +
    b1nodematch('type', alpha = .25, diff = TRUE) +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res3_bnodematch_terms0", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_patients) - mean(log(n_patients))) +
      I(log(n_hcp) - mean(log(n_hcp))) +
      facility_state
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    # gwb1dsp(fixed = FALSE, cutoff = 40) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2nodematch('diabetes', alpha = 0, diff = TRUE) +
    b1nodematch('type', alpha = 0, diff = TRUE) +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res3_bnodematch_terms_no_state", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_patients) - mean(log(n_patients))) +
      I(log(n_hcp) - mean(log(n_hcp)))
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    # gwb1dsp(fixed = FALSE, cutoff = 40) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2nodematch('diabetes', alpha = .25, diff = TRUE, levels = 2) +
    b1nodematch('type', alpha = .25, diff = TRUE) +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

tabulator(
  list(
    best_model,
    res3_b2starmix_terms,
    res3_b2starmix_terms_gwdsp,
    res3_bnodematch_terms,
    res3_bnodematch_terms0,
    res3_bnodematch_terms_no_state
  ),
  single.row = TRUE
)
```

The model that excludes the state term (facility) is significantly better than the model with it. This is also true if we compare the models using the b\*starmix vs the b\*nodematch terms. For the next batch, we will try to re-estimate model using the GWDSP term. Before we do such, let's first evaluate the individual counts of each variable to assess their correlatedness

```{r}
gwdsp_and_nodematch <- lapply(model_data$nets, \(net) {
  ans <- summary(net ~ gwb1dsp(0.7, fixed = TRUE) +
    b2nodematch('diabetes', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('mdro', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('ventilator', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('wound_care', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('bedridden', alpha = .25, diff = TRUE, levels = 2)
    )
}) |> do.call(what=rbind)

cor(gwdsp_and_nodematch)

```

## Estimating the model with the GWDSP term + bnodematch

```{r}
with_restore("res4_gwdsp_bnodematch", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_patients) - mean(log(n_patients))) +
      I(log(n_hcp) - mean(log(n_hcp)))
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(0.7, fixed = TRUE) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2nodematch('diabetes', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('mdro', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('ventilator', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('wound_care', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('bedridden', alpha = .25, diff = TRUE, levels = 2) +
    b1nodematch('type', alpha = .25, diff = TRUE) +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res4_gwdsp_bnodematch_no_logres", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_hcp) - mean(log(n_hcp)))
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(0.7, fixed = TRUE) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2nodematch('diabetes', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('mdro', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('ventilator', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('wound_care', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('bedridden', alpha = .25, diff = TRUE, levels = 2) +
    b1nodematch('type', alpha = .25, diff = TRUE) +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

init2 <- coef(res4_gwdsp_bnodematch_no_logres)
names(init2)[names(init2) == "gwb1dsp.fixed.0.7"] <- "gwb1dsp"

with_restore("res4_gwdsp_var_bnodematch_no_logres", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_hcp) - mean(log(n_hcp)))
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 100) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2nodematch('diabetes', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('mdro', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('ventilator', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('wound_care', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('bedridden', alpha = .25, diff = TRUE, levels = 2) +
    b1nodematch('type', alpha = .25, diff = TRUE) +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = control.ergm(
    init = init2,
    seed = 112
  )
))

tabulator(
  list(
    best_model,
    res3_bnodematch_terms_no_state,
    res4_gwdsp_bnodematch,
    res4_gwdsp_bnodematch_no_logres,
    res4_gwdsp_var_bnodematch_no_logres
  ),
  single.row = TRUE
)
```

Let's look at the MCMC diagnostics.

```{r}
#| eval: false
mcmc.diagnostics(res4_gwdsp_var_bnodematch_no_logres)
```

Let's do another run of the last model, but including the log(n_patients) term.

```{r}
with_restore("res4_gwdsp_var_bnodematch", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_patients) - mean(log(n_patients))) +
      I(log(n_hcp) - mean(log(n_hcp)))
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 100) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2nodematch('diabetes', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('mdro', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('ventilator', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('wound_care', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('bedridden', alpha = .25, diff = TRUE, levels = 2) +
    b1nodematch('type', alpha = .25, diff = TRUE) +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = control.ergm(
    seed = 112
  )
))

tabulator(
  list(
    best_model,
    res4_gwdsp_var_bnodematch_no_logres,
    res4_gwdsp_var_bnodematch
  ),
  single.row = TRUE
)
```

Let's look into the MCMC diagnostics.

```{r}
#| eval: false
mcmc.diagnostics(res4_gwdsp_var_bnodematch)
```

## Reproducing the original paper model

```{r}
with_restore("res5_original", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~ I(log(n)))+
    # Using "Other" as the reference level
    gwb1dsp(0.75, fixed = TRUE) +
    # Nodal effects
    b2factor('wound_care', levels = 2) +
    b2factor('bedridden', levels = 2) +
    # HCW centered 2-star mix
    b1starmix(2, 'ventilator') +
    b1starmix(2, 'diabetes') +
    # Resident-centered 2-star mix
    b2starmix(2, 'Nurse') +
    b2starmix(2, 'PTOTRT') +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = control.ergm(
    seed = 112,
    # Original estimates
    init = c(
      # edges, log(n) x edgs, gwdsp
      3.58, -1.11, -1.02,
      # Resident under wound, bedridden
      0.16, 0.2,
      # HCW centered 2-star mix
      0.01, 0.02, 0.04, 0.05,
      # Resident-centered 2-star mix
      -.11, -.37, .11, -.25,
      # offset
      -Inf, -Inf
      )
  )
))

tabulator(
  list(
    best_model,
    res4_gwdsp_var_bnodematch_no_logres,
    res4_gwdsp_var_bnodematch,
    res5_original
  ),
  single.row = TRUE
)
```

# Extra

We will try to fit the model using the state intercept


```{r}
with_restore("res6_full_and_facility", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_patients) - mean(log(n_patients))) +
      I(log(n_hcp) - mean(log(n_hcp))) +
      facility_state
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 100) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2nodematch('diabetes', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('mdro', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('ventilator', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('wound_care', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('bedridden', alpha = .25, diff = TRUE, levels = 2) +
    b1nodematch('type', alpha = .25, diff = TRUE) +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = control.ergm(
    main.method = "Stochastic-Approximation",
    seed = 112,
    SA.samplesize = 1024 * 8,
    SA.burnin = 1024 * 64, # Double
    SA.interval = 1024 * 8,
    SA.phase3_n = 1024 * 4
  )
))

with_restore("res6_full_and_facility_mcmle", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_patients) - mean(log(n_patients))) +
      I(log(n_hcp) - mean(log(n_hcp))) +
      facility_state
      )+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 100) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
    b2nodematch('diabetes', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('mdro', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('ventilator', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('wound_care', alpha = .25, diff = TRUE, levels = 2) +
    b2nodematch('bedridden', alpha = .25, diff = TRUE, levels = 2) +
    b1nodematch('type', alpha = .25, diff = TRUE) +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = control.ergm(
    init = coef(res6_full_and_facility),
    seed = 112
  )
))

tabulator(
  list(
    res4_gwdsp_var_bnodematch,
    res6_full_and_facility,
    res6_full_and_facility_mcmle
  ),
  single.row=TRUE
)
```


Saving the results

```{r}  
#| label: gof
#| eval: true
#| cache: true
#| message: false
#| warning: false
saveRDS(
  list(
    bfactors = best_model,
    gwdsp_no_log_patients = res4_gwdsp_var_bnodematch_no_logres,
    gwdsp_with_log_patients = res4_gwdsp_var_bnodematch,
    original = res5_original,
    full_and_facility = res6_full_and_facility,
    full_and_facility_mcmle = res6_full_and_facility_mcmle
  ),
  file = "models/03-pooled-ergms-final-results.rds"
)
```
