---
title: "Bipartite networks of health care workers and residents in long-term care facilities"
format: html
  # pdf:
  #   extra_dependencies:
  #     - pdflscape
  #     - rotating
  #     - tabularx
  #     - booktabs
  #     - geometry
  #     - graphicx
editor_options: 
  markdown: 
    wrap: 80
date: 2023-08-26
author:
  - Chong Zhang
  - George G. Vega Yon
date-modified: 2025-01-03
embed-resources: true
---

8/28/2023

```{r setup, include=FALSE}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
# a few personal functions to make things easy


```

## Overview

Recent updates:

1. We removed 6 networks with only 1 HCW - given the bipartite and no-isolate constraints they are fixed and no variation.

2. For ergm terms present in only a subset of networks, we now restricted their evaluation in only the subset of networks. For starmix(2),  at least two such nodes should be present for the network to be included.

3. There are only 3 physicians and 1 PA/NP in all the networks, therefore it is not very meaningful to evaluate the hemophily in these two types of HCWs. We grouped physician with 'Other', and PA/NP with 'Nurse'.

4. We detected one outlier, and removed it when fitting models to speed up the convergence, and for the final model selected, we added it back in.   

# Statistical methods

We have interaction data between HCWs and residents from a sample of 100 networks. ERGMs were used to understand network structure and its associations with nodal attribues. Given the fact that only interactions can be recorded in the data, we used constriction to forbid isolates. As a result, we excluded 7 networks with only 1 HCWS, because by the definitions of bipartite + no isolates, there are no variations in these networks given the set of nodes (saturated?), leaving 93 networks. Because our data is a sample of multiple networks, the N() ergm term in package ergm.mulit{} was used to evaluate each term jointly across multiple networks. ERGM terms we examined included edges and its interaction with network size, b2start(2), b2concurrent (residents having more than 1 HCWs), gwb1dsp (geometrically weighted shared partners between HCWs), HCW type (Nurse, CNA, PT/OT/RT, Other), and resident care needs (dialysis, wound_care, bedridden, ventilator, diabetes, wheelchair, mdro). Terms are added one at a time to ensure convergence (assessed using trace plot) and reasonable goodness of fit (assessed using standardized pearson's residuals using function gofN()). For highly correlated terms (b2star, b2condurrent, gwb1dsp), we kept the one that gives the smallest AIC/BIC and dropped the rest. As the model got more complex and the algorithms were slow to converge, we removed one outlier to speed up the convergence. The largest model with all terms being significant was selected as the best model. We then used the estimated coefficients from this model as initial values to fit the model on all 93 networks, and compared results before and after adding back the outlier.



```{r}
#| label: packages
#| message: false
#| warning: false
source("models/helpers.R")

sessionInfo()
```

Load network data

```{r}
#| label: read-clean
load('data/network93_f2.RData') # a list of 93 networks to fit ERGMs
```


# Fitting ERGMs

## Model 0: edges only, with scaling by log(n) and state level effect

```{r}
#| label: prepare-data
model_data <- set_ergm(nets = networks)
```

We will also be applying scaling based on the number of actors/entities in the network. We calculate the number of each using the following code:

Common controls for ergm models: we use Stochastic approximation since it is fast. 
```{r}
#| label: controls
default_control_mcmle <- control.ergm(seed = 55551)

default_control_sa <- control.ergm(
  seed            = 44485,
  main.method     = "Stochastic-Approximation", # Robbins-Monro Stochastic Approx.
  SA.nsubphases   = 4 * 1, # 1 the default
  SA.burnin       = 1024 * 4, # Twice the default
  SA.interval     = 1024 * 4 # four times... This is being passed to SA instead
)
```

Model 0: edges only.
Since we do not have isolates in our network by design, we restrict the network by adding offset terms b1degree(0) and b2degree(0) with coefficients -Inf.
```{r}
#| label: model0
#| message: false
#| warning: false
#| results: hide
with_restore("res0", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~log(n_tot) + I(log(n_tot)^2))+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res0_state", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~log(n_tot) + facility_state)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res0_state2", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~log(n_tot) + I(log(n_tot)^2) + facility_state)+
    offset(b1degree(0))+offset(b2degree(0)),
    offset.coef = c(-Inf, -Inf),
    control = default_control_mcmle
))

with_restore("res0_state_centered", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~I(log(n_tot) - mean(log(n_tot))) + facility_state)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

tabulator(
  list(
    res0,
    res0_state,
    res0_state2,
    res0_state_centered
  ),
  single.row = TRUE
)
```

Model 4, the mean-centered, is the one with best fit overall. Adding the square term for log(n) did not improve the fit, but adding the state level effect did. Let's investiage the convergence:

```{r}
#| label: model0-convergence
mcmc.diagnostics(res0_state_centered)
```

The model does a good job in convergence: the MCMC trace shows good mixing, the omnibus test is not significant, and the autocorrelation reduces decently fater the third lag. We will be using this model as the baseline for the subsequent models.

```{r}
best_model <- res0_state_centered

with_restore("res0_hcp", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~I(log(n_hcp) - mean(log(n_hcp))) + facility_state)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res0_patients", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~I(log(n_patients) - mean(log(n_patients))) + facility_state) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

tabulator(
  list(best_model, res0_hcp, res0_patients),
  single.row = TRUE
)
```

The Model 2 (correcting for the log(n_hcp)) has the best fit overall. This is inline with what we expect in theory. The degree distribution of HCPs is more likely to be affected by the network size than that of patients, because as the number of patients increases, the degree of HCPs increases, but the degree of patients does not necessarily increase as much. Thus, adjusting for log(n_hcp) is more appropriate in this application. We will be using log(n_hcp) for the rest of the models. Let's investiagete convergence for this model:

```{r}
#| label: model0_hcp-convergence
mcmc.diagnostics(res0_hcp)
```

The omnibus test is not significant, the MCMC trace shows good mixing, and the autocorrelation reduces decently after the third lag. We will be using this model as the baseline for the subsequent models. **Nevertheless, the individual fits are worse off, particularly for the OR facilities**; let's see if we can adjust for both simultaneously:

```{r}
with_restore("res0_both", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~
      I(log(n_hcp) - mean(log(n_hcp))) +
      I(log(n_tot) - mean(log(n_tot))) +
      + facility_state
      )+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

tabulator(
  list(best_model, res0_hcp, res0_patients, res0_both),
  single.row = TRUE
)
```

At first look, it seems this approach performs significantly better. Let's look at the convergence diagnostics:

```{r}
#| label: model0_both-convergence
mcmc.diagnostics(res0_both)
```

Looking at the diagnostics, this model did not converge well. The omnibus test is significant, so we are worse off than the previous model. Although scaling by the number of healthcare providers seemed like the right theoretical approach, the model is overall performing worse than just scaling by the number of nodes. We will be using the model with log(n) as the baseline for the subsequent models.

## Controling for type of HCP

```{r}
#| label: model-hcp-type
best_model <- res0_hcp

with_restore("res1_hcp_type", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~ I(log(n_hcp) - mean(log(n_hcp))) + facility_state)+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

tabulator(
  list(
    best_model,
    res1_hcp_type
  ),
  single.row = TRUE
)
```

Let's check the convergence diagnostics:

```{r}
#| label: model-hcp-type-convergence
mcmc.diagnostics(res1_hcp_type)
```

Since the overall fit is good but the omnibus test is failing, let's see if adding more structure changes things.

## Investigating the effect of gwb1dsp

Since we expect the healthcare providers been somewhat evenly spread across residents, we expect to see a tendency of negative GWDSP effect for the HCP. Let's start with a basic model. Since the process generally is slow to converge, we will instead use the Stochastic Approximation method for now on:

Add b2 2-star (second level 2-stars) term:
```{r}
#| label: model1
#| message: false
#| warning: false
#| results: hide
with_restore("res1_gwb1dsp", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~ I(log(n_hcp) - mean(log(n_hcp))) + facility_state)+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 40) + 
    +offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

tabulator(
  list(
    best_model,
    res1_hcp_type,
    res1_gwb1dsp
  ),
  single.row = TRUE
)
```

Although the AIC and BIC suggest that the new model is better, we need to check the convergence and goodness of fit before making a conclusion.

```{r}
#| label: model1_mcmle_convergence
mcmc.diagnostics(res1_gwb1dsp)
```

The omnibus test is not significant, and the same applies to the MCMC joint P-value. The new model behaves significantly better than the previous, even improving in the omnibus test.

Using Stewart, Schweinberger, Bojanowski, and Morris (2019) (as suggested by the reviewer), I estimated the decay parameter directly. I was not able to directly estimate it with the interaction effect as ERGM failed with an error `objfun returned gradient not having all elements finite`.

## Looking into other terms

Now add b2factor and b2starmix (just Nurse for now)
```{r}
#| label: model5
#| eval: true
#| message: false
#| warning: false
#| results: hide
best_model <- res1_gwb1dsp

with_restore("res2_b2factor_terms", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~ I(log(n_hcp) - mean(log(n_hcp))) + facility_state)+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    gwb1dsp(fixed = FALSE, cutoff = 40) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

with_restore("res2_b2factor_terms_nogwdsp", model_data, ergm(
  Networks(nets) ~
    N(~edges, ~ I(log(n_hcp) - mean(log(n_hcp))) + facility_state)+
    # Using "Other" as the reference level
    b1factor("type", levels = -3) +
    b2factor('dialysis') +
    b2factor('wound_care') +
    b2factor('bedridden') +
    b2factor('ventilator') +
    b2factor('mdro') +
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))


tabulator(
  list(
    best_model,
    res2_b2factor_terms,
    res2_b2factor_terms_nogwdsp
  ),
  single.row = TRUE
)
```

```{r}
with_restore("gofN_res2_b2factor_terms", list(), {
  gofN(res2_b2factor_terms)
})
if (interactive()) {
  pdf("gofN_res2_b2factor_terms.pdf")
  op <- par(mfrow = c(3, 2))
  plot(gofN_res2_b2factor_terms)
  par(op)
  dev.off()
}
```

There are some states for which the edge count is not fitted well by the model, and the same happens for the number of nurses. The intercept effect for NY and TN fails the test predicting the number of edges in those states, and the same applies to the number of nurses.


## Subsetting the network to exclude MD

Maryland has 6 networks only, and the degree distribution of these networks is quite different from the rest of the networks. 

```{r}
#| label: model3-without-md
nets2 <- networks[model_data$facility_state != "MD"]
nets2 <- lapply(seq_along(nets2), \(i) {
  net <- nets2[[i]]
  net %v% "net_id" <- i
  net
})
model_data_no_md <- set_ergm(nets = nets2)

with_restore("res2_b2factor_terms_no_md", model_data_no_md, ergm(
  Networks(nets) ~
    N(~edges, ~ I(log(n_tot) - mean(log(n_tot))) + facility_state)+
    N(~b1factor("Nurse"), subset = Nurse)+
    # N(~b1factor("CNA"), subset = CNA)+
    N(~b1factor("PTOTRT"), subset = PTOTRT)+
    gwb1dsp(0.2, fixed = TRUE) +
    # N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = control.ergm(
    seed = 55551,
    MCMLE.confidence = .995
  )
))

tabulator(
  list(
    best_model,
    res2_b2factor_terms_refit,
    res2_b2factor_terms_no_md
  ),
  single.row = TRUE
)
```

We now assess the MCMC convergence and GOF for the new model without MD networks:

```{r}
#| label: model3-no-md-convergence
mcmc.diagnostics(res2_b2factor_terms_no_md)
```


## Adding the HCP and patient centered 2-star mix

```{r}
#| label: model4
with_restore("res4_b1starmix", model_data_no_md, ergm(
  Networks(nets) ~
    N(~edges, ~ I(log(n_tot) - mean(log(n_tot))) + facility_state)+
    N(~b1factor("Nurse"), subset = Nurse)+
    # N(~b1factor("CNA"), subset = CNA)+
    N(~b1factor("PTOTRT"), subset = PTOTRT)+
    gwb1dsp(0.2, fixed = TRUE) +
    # N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'ventilator'), subset=ventilator2)+
    N(~b1starmix(2, 'diabetes'), subset=diabetes2)+
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

tabulator(
  list(
    best_model,
    res2_b2factor_terms_no_md,
    res4_b1starmix
  ),
  single.row = TRUE
)
```

Taking a look at the convergence diagnostics:

```{r}
#| label: model4-convergence
mcmc.diagnostics(res4_b1starmix)
```

The model is failing to converge. Ths SD of the diabetes.hwc.no has a high deviation from 0 (111.08), and the omnibus test is significant. Let's drop that statistic and use a b2factor for diabetes instead:

```{r}
#| label: model4-refit
with_restore("res4_b1starmix_diabetes", model_data_no_md, ergm(
  Networks(nets) ~
    N(~edges, ~ I(log(n_tot) - mean(log(n_tot))) + facility_state)+
    N(~b1factor("Nurse"), subset = Nurse)+
    # N(~b1factor("CNA"), subset = CNA)+
    N(~b1factor("PTOTRT"), subset = PTOTRT)+
    gwb1dsp(0.2, fixed = TRUE) +
    # N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'ventilator'), subset=ventilator2)+
    N(~b2factor('diabetes'), subset=diabetes)+
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = default_control_mcmle
))

tabulator(
  list(
    best_model,
    res2_b2factor_terms_no_md,
    res4_b1starmix,
    res4_b1starmix_diabetes
  ),
  single.row = TRUE
)
```

Verifying the convergence diagnostics for the new model:

```{r}
#| label: model4-refit-convergence
mcmc.diagnostics(res4_b1starmix_diabetes)
```

```{r}
#| label: model6
#| eval: true
#| message: false
#| warning: false
#| results: hide
best_model <- res5new_old

res6_new_old <- ergm(
  Networks(networks1) ~
    N(~edges, ~ log(n) + I(as.factor(facility_state)))+
    N(~gwb1dsp(gb1dsp_decay, fixed=TRUE), ~log(n_hcp) + log(n_patients))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
    N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'wound_care', diff=TRUE), subset=wound_care2)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1))

res6_new <- ergm(
  Networks(networks1) ~
    N(~edges, ~ log(n) + I(as.factor(facility_state)))+
    N(~gwb1dsp(gb1dsp_decay, fixed=TRUE), ~log(n_hcp) + log(n_patients))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
    N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'wound_care', diff=TRUE), ~ log(n_hcp), subset=wound_care2)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1))

res6_new_all_scaled <- ergm(
  Networks(networks1) ~
    N(~edges, ~ log(n) + I(as.factor(facility_state)))+
    N(~gwb1dsp(gb1dsp_decay, fixed=TRUE), ~log(n_hcp) + log(n_patients))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
    N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'wound_care', diff=TRUE), ~ log(n_hcp) + log(n_patients), subset=wound_care2)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1))

res6_new_n_patients <- ergm(
  Networks(networks1) ~
    N(~edges, ~ log(n) + I(as.factor(facility_state)))+
    N(~gwb1dsp(gb1dsp_decay, fixed=TRUE), ~log(n_hcp) + log(n_patients))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
    N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'wound_care', diff=TRUE), ~ log(n_patients), subset=wound_care2)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1)
)

tabulator( 
  list(
    'Current' = best_model,
    '+ wound *mix'=res6_new_old,
    '+ log(n_hcp)'=res6_new,
    '+ log(n_hcp) + log(n_patients)'=res6_new_all_scaled,
    '+ log(n_patients)'=res6_new_n_patients #,
    # '+ log(n_patients) + log(n_patients) MCMLE'=res6_new_n_patients_mcmle
  ),
  single.row = TRUE
)
```

```{r}
#| label: model8new
#| eval: true
#| message: false
#| warning: false
#| results: hide
best_model <- best_model
res8_new_old <- ergm(
  Networks(networks1) ~
    N(~edges, ~ log(n) + I(as.factor(facility_state)))+
    N(~gwb1dsp(gb1dsp_decay, fixed=TRUE), ~log(n_hcp) + log(n_patients))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
    N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'ventilator', diff=TRUE), subset=ventilator2)+
    N(~b1starmix(2, 'diabetes', diff=TRUE), subset=diabetes2)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1))

res8_new <- ergm(
  Networks(networks1) ~
    N(~edges, ~ log(n) + I(as.factor(facility_state)))+
    N(~gwb1dsp(gb1dsp_decay, fixed=TRUE), ~log(n_hcp) + log(n_patients))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
    N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'ventilator', diff=TRUE), ~ log(n), subset=ventilator2)+
    N(~b1starmix(2, 'diabetes', diff=TRUE), ~ log(n), subset=diabetes2)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1))

res8_new_old_diabetes <- ergm(
  Networks(networks1) ~
    N(~edges, ~ log(n) + I(as.factor(facility_state)))+
    N(~gwb1dsp(gb1dsp_decay, fixed=TRUE), ~log(n_hcp) + log(n_patients))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
    N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'diabetes', diff=TRUE), subset=diabetes2)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1))

res8_new_diabetes <- ergm(
  Networks(networks1) ~
    N(~edges, ~ log(n) + I(as.factor(facility_state)))+
    N(~gwb1dsp(gb1dsp_decay, fixed=TRUE), ~log(n_hcp) + log(n_patients))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
    N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'diabetes', diff=TRUE), ~ log(n), subset=diabetes2)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1))

tabulator(
  list(
    'Current' = best_model,
    'b1starmix(vent|diabetes)'=res8_new_old,
    '+ log(n)'=res8_new,
    'b1starmix(diabetes) + log(n)'=res8_new_diabetes,
    'b1starmix(diabetes) only'=res8_new_old_diabetes
  ),
  single.row = TRUE
)
```


```{r}
#| label: res9new
#| cache: true
#| eval: true
#| message: false
#| warning: false
#| results: hide
best_model <- best_model

res9_new <- ergm(
  Networks(networks1) ~
    N(~edges, ~ log(n) + I(as.factor(facility_state)))+
    N(~gwb1dsp(gb1dsp_decay, fixed=TRUE), ~log(n_hcp) + log(n_patients))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
    N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b2factor('mdro'), subset=mdro)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1))

res9_new_log_n <- ergm(
  Networks(networks1) ~
    N(~edges, ~ log(n) + I(as.factor(facility_state)))+
    N(~gwb1dsp(gb1dsp_decay, fixed=TRUE), ~log(n_hcp) + log(n_patients))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
    N(~b2factor('dialysis'), subset=dialysis)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b2factor('mdro'), ~ log(n), subset=mdro)+
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1))

tabulator(
  list(
    'Current' = best_model,
    '+ mdro factor'=res9_new,
    '+ mdro factor log(n)'=res9_new_log_n
  ),
  single.row = TRUE
)
```


```{r}
res10 <- ergm(
  Networks(networks1) ~ N(~edges, ~log(n) + I(as.factor(facility_state))) + 
    N(~gwb1dsp(gb1dsp_decay, fixed = TRUE), ~log(n_hcp) + log(n_patients)) + 
    N(~b2starmix(2, "Nurse"), subset = Nurse) + N(~b2factor("dialysis"), 
    subset = dialysis) + N(~b2factor("wound_care"), subset = wound_care) + 
    N(~b2factor("bedridden"), subset = bedridden) +
    N(~b2starmix(2, 'PTOTRT'), subset=PTOTRT)+
    offset(b1degree(0)) + offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1)
)

res10_log_n <- ergm(
  Networks(networks1) ~ N(~edges, ~log(n) + I(as.factor(facility_state))) + 
    N(~gwb1dsp(gb1dsp_decay, fixed = TRUE), ~log(n_hcp) + log(n_patients)) + 
    N(~b2starmix(2, "Nurse"), subset = Nurse) + N(~b2factor("dialysis"), 
    subset = dialysis) + N(~b2factor("wound_care"), subset = wound_care) + 
    N(~b2factor("bedridden"), subset = bedridden) +
    N(~b2starmix(2, 'PTOTRT'), ~log(n_hcp),subset=PTOTRT)+
    offset(b1degree(0)) + offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1)
)

tabulator(
  list(
    'Current' = best_model,
    '+ PTOTRT starmix'=res10,
    '+ PTOTRT starmix log(n)'=res10_log_n
  ),
  single.row = TRUE
)
```

[REVIEW] Adding the PTOTRT starmix term improved the model fit, but adding the log(n) scaling did not improve the fit. So far, the new model is doing substantially better than the first version of the paper

Additional test, centering log(n) at mean(log(n)) to see if it helps with convergence.

```{r}
# Centered
log_n_centered <- log(n_tot) - mean(log(n_tot))
log_n_hcp_centered <- log(n_hcp) - mean(log(n_hcp))
log_n_patients_centered <- log(n_patients) - mean(log(n_patients))

res10_log_n_centered <- ergm(
  Networks(networks1) ~
    N(~edges, ~log_n_centered + I(as.factor(facility_state))) + 
    N(~gwb1dsp(gb1dsp_decay, fixed = TRUE), ~log_n_hcp_centered + log_n_patients_centered) + 
    N(~b2starmix(2, "Nurse"), subset = Nurse) + N(~b2factor("dialysis"), 
    subset = dialysis) + N(~b2factor("wound_care"), subset = wound_care) + 
    N(~b2factor("bedridden"), subset = bedridden) + N(~b2starmix(2, 
    "PTOTRT"), subset = PTOTRT) + offset(b1degree(0)) + offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1)
)

res10_log_n_centered_simpler <- ergm(
  Networks(networks1) ~
    N(~edges, ~log_n_centered + I(as.factor(facility_state))) + 
    N(~gwb1dsp(gb1dsp_decay, fixed = TRUE), ~log_n_hcp_centered) + 
    N(~b2starmix(2, "Nurse"), subset = Nurse) + N(~b2factor("dialysis"), 
    subset = dialysis) + N(~b2factor("wound_care"), subset = wound_care) + 
    N(~b2factor("bedridden"), subset = bedridden) + N(~b2starmix(2, 
    "PTOTRT"), subset = PTOTRT) + offset(b1degree(0)) + offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call("control.ergm", default_control.1)
)

tabulator(
  list(
    'Current' = res10,
    '+ log(n) centered'=res10_log_n_centered,
    '+ log(n) centered simpler'=res10_log_n_centered_simpler
  ),
  single.row = TRUE
)
```

```{r}
model_may2026 <- ergm(
  formula(res10),
  offset.coef = c(-Inf, -Inf),
  control = control.ergm(seed = 113, init = coef(res10))
)

model_may2026_mean_centered <- ergm(
  formula(res10_log_n_centered),
  offset.coef = c(-Inf, -Inf),
  control = control.ergm(seed = 113, init = coef(res10_log_n_centered))
)

model_may2026_log_n_simpler <- ergm(
  formula(res10_log_n_centered_simpler),
  offset.coef = c(-Inf, -Inf),
  control = control.ergm(seed = 113, init = coef(res10_log_n_centered_simpler))
)

tabulator(
  list(
    'Model May 2026'=model_may2026,
    'Model May 2026 log(n) centered'=model_may2026_mean_centered,
    'Model May 2026 log(n) centered simpler'=model_may2026_log_n_simpler
  ),
  single.row = TRUE
)

saveRDS(model_may2026, file = 'data/model_may2026.rds')
saveRDS(model_may2026_mean_centered, file = 'data/model_may2026_mean_centered.rds')
saveRDS(model_may2026_log_n_simpler, file = 'data/model_may2026_log_n_simpler.rds')
```

```{r}
#| label: final-model-92-networks
#| cache: true
#| eval: true
#| message: false
#| warning: false
#| results: hide
default_control.1$main.method <- "MCMLE"

timer_res12_92 <- system.time({
  res12_92 <- ergm(
    Networks(networks1) ~
      N(~edges, ~log(n))+
      N(~gwb1dsp(0.75, fixed=TRUE))+
      N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
      #N(~b2factor('dialysis'), subset=dialysis)+
      N(~b2factor('wound_care'), subset=wound_care)+
      N(~b2factor('bedridden'), subset=bedridden)+
      N(~b1starmix(2, 'wound_care', diff=TRUE), subset=wound_care)+
      N(~b1starmix(2, 'ventilator', diff=TRUE), subset=ventilator2)+
      N(~b1starmix(2, 'diabetes', diff=TRUE), subset=diabetes2)+
      N(~b2starmix(2, 'PTOTRT'), subset=PTOTRT)+ # has to be numeric
      offset(b1degree(0))+offset(b2degree(0)),
    offset.coef = c(-Inf, -Inf),
    control = do.call("control.ergm", default_control.1))
})

# Printing time
timer_res12_92

save(
  list = c(
    "res12_92", "Nurse", "wound_care", "bedridden", "ventilator2",
    "diabetes2", "PTOTRT", "timer_res12_92"
    ),
  file='data/res12_92.rda')
```

Priting the summary

```{r}
#| label: summary-model-12-92-nets
summary(res12_92)
```

```{r}
#| label: res12_93
#| echo: false
#| cache: true
#| eval: true
#| message: false
#| warning: false
#| results: hide
# Resetting, keeping only the seed
load('data/dd_nodal_attr.RData')
dim(dd)
setDT(dd)
dd$NetID=as.numeric(dd$NetID)

Nurse  = dd[, n:=sum(Nurse==1), NetID][n>1, unique(NetID)] |> as.numeric() # requiring at least 2 nurses in a network 
CNA    = dd[, n:=sum(CNA==1), NetID][n>1, unique(NetID)] |> as.numeric()
PTOTRT = dd[, n:=sum(PTOTRT==1), NetID][n>1, unique(NetID)] |> as.numeric()
Other  = dd[, n:=sum(Other==1), NetID][n>1, unique(NetID)] 
nuni(c(Other, PTOTRT)) # 46

wh=intersect(Nurse, PTOTRT)
length(wh)
dcheck=dd[NetID %in% wh & group=='hcw', .(NetID, Nurse, PTOTRT, Other, CNA)]
dcheck[, count:=sum(CNA=="1")+sum(Other=="1"), NetID]
PTOTRT.nurse=as.numeric(dcheck[count>1, unique(NetID)])
length(PTOTRT.nurse) # all 30 networks are fine!

dialysis=dd[, n:=sum(dialysis=='Yes'), NetID][n>0, unique(NetID)]
wound_care=dd[, n:=sum(wound_care=='Yes'), NetID][n>0, unique(NetID)] 
mdro=dd[, n:=sum(mdro=='Yes'), NetID][n>0, unique(NetID)] 
bedridden=dd[, n:=sum(bedridden=='Yes'), NetID][n>0, unique(NetID)]  # requiring at least 2 nurses in a network 
bedridden

dialysis2=dd[, n:=sum(dialysis=='Yes'), NetID][n>1, unique(NetID)]
diabetes2=dd[, n:=sum(diabetes=='Yes'), NetID][n>1, unique(NetID)]
wound_care2=dd[, n:=sum(wound_care=='Yes'), NetID][n>1, unique(NetID)] 
ventilator2=dd[, n:=sum(ventilator=='Yes'), NetID][n>1, unique(NetID)] 
mdro2=dd[, n:=sum(mdro=='Yes'), NetID][n>1, unique(NetID)] 
wheelchair2=dd[, n:=sum(wheelchair=='Yes'), NetID][n>1, unique(NetID)]  # requiring at least 2 nurses in a network 
wheelchair2

# Too slow to fit this model, it's already been about 2 hours.
# default_control.1 <- list(seed = 1123)
timer_res12_93 <- system.time({
  res12_93 <- ergm(
    Networks(networks) ~
      N(~edges, ~log(n))+
      N(~gwb1dsp(0.75, fixed=TRUE))+
      N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
      #N(~b2factor('dialysis'), subset=dialysis)+
      N(~b2factor('wound_care'), subset=wound_care)+
      N(~b2factor('bedridden'), subset=bedridden)+
      N(~b1starmix(2, 'wound_care', diff=TRUE), subset=wound_care)+
      N(~b1starmix(2, 'ventilator', diff=TRUE), subset=ventilator2)+
      N(~b1starmix(2, 'diabetes', diff=TRUE), subset=diabetes2)+
      N(~b2starmix(2, 'PTOTRT'), subset=PTOTRT)+ # has to be numeric
      offset(b1degree(0))+offset(b2degree(0)),
    offset.coef = c(-Inf, -Inf),
    control = do.call(
      "control.ergm",
      c(list(init = coef(res12_92)), default_control.1)
      )
  )
})

# Printing time
timer_res12_93

save(
  list = c(
    "res12_93", "Nurse", "wound_care", "bedridden", "ventilator2",
    "diabetes2", "PTOTRT", "timer_res12_93"
    ),
  file='data/res12_93.rda'
  )
```

```{r}
#| label: res12_93_no_start
#| echo: false
#| cache: true
#| eval: false
#| message: false
#| warning: false
#| results: hide
timer_res12_93_no_start <- system.time({
  res12_93_no_start <- ergm(
    Networks(networks) ~
      N(~edges, ~log(n))+
      N(~gwb1dsp(0.75, fixed=TRUE))+
      N(~b2starmix(2, 'Nurse'), subset=Nurse)+ # has to be numeric
      #N(~b2factor('dialysis'), subset=dialysis)+
      N(~b2factor('wound_care'), subset=wound_care)+
      N(~b2factor('bedridden'), subset=bedridden)+
      N(~b1starmix(2, 'wound_care', diff=TRUE), subset=wound_care)+
      N(~b1starmix(2, 'ventilator', diff=TRUE), subset=ventilator2)+
      N(~b1starmix(2, 'diabetes', diff=TRUE), subset=diabetes2)+
      N(~b2starmix(2, 'PTOTRT'), subset=PTOTRT)+ # has to be numeric
      offset(b1degree(0))+offset(b2degree(0)),
    offset.coef = c(-Inf, -Inf),
    control = do.call(
      "control.ergm",
      default_control.1
      )
  )
})

# Printing time
timer_res12_93_no_start

save(
  list = c(
    "res12_93_no_start", "Nurse", "wound_care", "bedridden", "ventilator2",
    "diabetes2", "PTOTRT", "timer_res12_93_no_start"
    ),
  file='data/res12_93_no_start.rda'
  )

summary(res12_93_no_start)
```

Printing the ERGM results

```{r}
#| label: summary-model-12
summary(res12_93)

# Printing estimation time
timer_res12_93
```




```{r}
#| eval: true
#| results: asis
#| echo: true
texreg::htmlreg(
  list(
    'Model 12 on 92 networks'=res12_92,
    'Model 12 on 93 networks'=res12_93 #,
    # 'Model 12 on 93 networks (no start)'=res12_93_no_start
    ),
  single.row = TRUE,
  scalebox = .8,
  use.packages = FALSE
  )
```


Evaluate goodness of fit.
```{r}  
#| label: gof
#| eval: true
#| cache: true
#| message: false
#| warning: false
g=gofN(res12_93) # saved as gofN_model12_93
summary(g)
saveRDS(g, 'data/gofN_model12_93.rds')
```

