---
title: "Bootstrapping ERGMs (blocked at the facility level)"
format: html
editor_options:
  markdown:
    wrap: 80
date: 2026-08-19
author:
  - George G. Vega Yon
embed-resources: true
---

This is the blocked-bootstrap counterpart of `04d-boot-ergms.qmd`. Instead of
resampling networks independently, we resample *facilities*: whenever a facility
is drawn, all of the networks belonging to it (its units, across dates) enter the
replicate together. This accounts for the within-facility correlation that the
naive bootstrap ignores.

```{r}
source("models/helpers.R")
load('data/network93_f2.RData')
fits <- readRDS("models/03-pooled-ergms-final-results.rds")
```

## Defining the blocking variable

Each network corresponds to a `state`/`Fac_Name`/`Fac_Unit`/date combination, so
a single facility can contribute several networks. We block on `state` +
`Fac_Name` since facility names are not guaranteed to be unique across states.

```{r}
cluster_id <- sapply(networks, \(net) {
  paste(unique(net %v% "Fac_Name"), unique(net %v% "Fac_Unit"), sep = "::")
}) |> unname()

# How many networks per facility?
table(table(cluster_id))
```

```{r}
c(
  n_networks  = length(networks),
  n_clusters  = length(unique(cluster_id))
)
```

## Evaluating in the final model

```{r}
# Naive 9 cluster (for now)
cl <- parallel::makeForkCluster(29)

set.seed(333)
ans <- boot_ergm(
  networks = networks,
  ergm_formula =
    Networks(nets) ~
      N(~ edges,
        ~
          I(log(n_patients) - mean(log(n_patients))) +
          I(log(n_hcp) - mean(log(n_hcp)))
      ) + b1factor("type", levels = -3) +
    N(
      ~gwb1dsp(0.84, fixed = TRUE),
      ~I(log(n_patients) - mean(log(n_patients))) +
      I(log(n_hcp) - mean(log(n_hcp)))
      ) +
    b2factor("dialysis") +
    b2factor("wound_care") + b2factor("bedridden") + b2factor("ventilator") +
    b2factor("mdro") + b2nodematch("diabetes", alpha = 0.25,
    diff = TRUE, levels = 2) + b2nodematch("mdro", alpha = 0.25,
    diff = TRUE, levels = 2) + b2nodematch("ventilator", alpha = 0.25,
    diff = TRUE, levels = 2) + b2nodematch("wound_care", alpha = 0.25,
    diff = TRUE, levels = 2) + b2nodematch("bedridden", alpha = 0.25,
    diff = TRUE, levels = 2) + b1nodematch("type", alpha = 0.25,
    diff = TRUE) + offset(b1degree(0)) + offset(b2degree(0)),
  control = control.ergm(
    seed = 1123,
    # fits$gwdsp_with_log_patients |> coef() |> unname() |> dput()
    main.method = "Stochastic-Approximation",
    SA.samplesize = 1024 * 8,
    SA.burnin = 1024 * 64, # Double
    SA.interval = 1024 * 8,
    SA.phase3_n = 1024 * 4,
    # We don't care about these errors, or evaluating the log-likelihood
    MCMC.addto.se = FALSE,
    MCMC.return.stats = FALSE
  ),
  offset.coef = c(-Inf, -Inf),
  nboot = 200,
  cluster_id = cluster_id,
  cl = cl,
  cache = "models/04e-boot-ergms-cache"
)
parallel::stopCluster(cl)
```


## Processing the bootstrap results


### Intercept model

```{r}
#| eval: true
# Listing all the bootstrap results
boot_res <- list.files(
  "models/04e-boot-ergms-cache",
  pattern = "\\.rds$",
  full.names = TRUE
  ) |>
  lapply(readRDS)

# Replicates that failed are stored as error objects
boot_res <- boot_res[!sapply(boot_res, inherits, what = "error")]

boot_res <- do.call(what = rbind, boot_res)

# Keeping smaller than 20
n_terms <- ncol(boot_res)
boot_res <- boot_res[apply(boot_res[, c(1:(n_terms-2))], 1, \(x) all(abs(x) < 20)), ]

# Generating a regression table
naive_fit <- fits$gwdsp_intercept
boot_res_mean <- apply(boot_res, 2, mean, na.rm = TRUE)
boot_res_sd <- apply(boot_res, 2, sd, na.rm = TRUE)
pvalues <- 2 * (1 - pnorm(abs(coef(naive_fit)) / boot_res_sd))
```


```{r}

texreg::screenreg(
  list(
    'No intercept (naive)'=fits$gwdsp_intercept,
    'No intercept (block boot)'=fits$gwdsp_intercept
    ),
  override.se = list(NULL, boot_res_sd),
  override.pvalues = list(NULL, pvalues),
  single.row = TRUE
)
```


```{r}
#| eval: true
texreg::screenreg(
  list(
    'No intercept (naive)'=fits$gwdsp_intercept,
    'No intercept (block boot)'=fits$gwdsp_intercept
    ),
  override.se = list(NULL, boot_res_sd),
  override.pvalues = list(NULL, pvalues),
  single.row = TRUE,
  file = "models/04e-boot-ergms-results.txt"
)

saveRDS(
  list(
    bootstrap_results = boot_res,
    bootstrap_means = boot_res_mean,
    bootstrap_sds = boot_res_sd,
    pvalues = pvalues,
    naive_fit = naive_fit,
    cluster_id = cluster_id
  ),
  file = "models/04e-boot-ergms-results.rds"
)
```

## Comparing against the naive bootstrap

```{r}
#| eval: true
naive_boot <- readRDS("models/04d-boot-ergms-results.rds")

data.frame(
  term       = names(boot_res_sd),
  sd_naive   = naive_boot$bootstrap_sds,
  sd_blocked = boot_res_sd,
  ratio      = boot_res_sd / naive_boot$bootstrap_sds
)
```
