---
format: docx
title: "Supplemental Materials for \"Exponential-Family Random Graph Models in Resident-Healthcare Provider Networks: An Application using data from Long-term Healthcare Facilities in the United States\""
author:
  - George G. Vega Yon
  - Chong Zhang
  - Karim Khader
  - Nai-Chung Nelson Chang
  - Lindsay D. Visnovsky
  - Alun Thomas
  - Candace Haroldsen
  - Kristina Stratford
  - Matthew H. Samore
date: 2025-04-01
date-modified: 2025-05-15
embed-resources: true
---


This document presents the supplemental materials for the main manuscript, focusing on the goodness-of-fit (GOF) assessment and diagnostics for the final multi-ERGM model. The following sections provide detailed explanations and outputs to facilitate reproducibility and transparency for readers interested in the modeling process and results.

The code chunks below are annotated to clarify their purpose and the rationale behind each step, following best practices for reporting ERGM-based social network analyses.

```{r}
#| warning: false
#| message: false
library(ergm.multi)
library(texreg)
library(ergmito)
library(ggplot2)
library(ggrepel)
library(ggridges)
library(data.table)
library(xtable)
```

For terms wrapped in `ergm.multi::N()`, the subset argument is applied at the network level. For example, in `N(~b2starmix(2, "Nurse"), subset = Nurse)`, "Nurse" inside `b2starmix()` denotes the vertex-level HCW type indicator, whereas the object Nurse supplied to subset is a numeric vector of network IDs. This vector contains networks with at least two nurse HCWs, so the nurse-specific 2-star mixing term is evaluated only for networks in which that statistic is defined and informative. Similar network-index subsets were constructed for other attribute-specific terms.

## Summary stats

The following code chunk loads the relevant model objects and nodal attribute data required for the subsequent analyses. It also prepares a summary of the presence of multidrug-resistant organisms (MDROs) across the networks, which is a key covariate in the study. The summary_formula function is then used to compute the observed values for each model term, providing a basis for later comparison with model-based expectations.

```{r}
#| label: load-packages-and-data
# Loading data from the models
load("data/res12_93.rda")
load('data/network93_f2.RData')
load('data/dd_nodal_attr.RData')
setDT(dd)
dd$NetID=as.numeric(dd$NetID)

# Summary about MDROs
mdro <- dd[, n := sum(mdro == 'Yes'), NetID][n>0, unique(NetID)] 

counts <- summary_formula(
  res12_93$network ~
    N(~edges, ~log(n))+
    N(~gwb1dsp(0.75, fixed=TRUE))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'wound_care', diff=TRUE), subset = wound_care) +
    N(~b1starmix(2, 'ventilator', diff=TRUE), subset = ventilator2) +
    N(~b1starmix(2, 'diabetes', diff=TRUE), subset=diabetes2) +
    N(~b2starmix(2, 'PTOTRT'), subset=PTOTRT) +
    N(~b2starmix(2, 'mdro', diff = TRUE), subset=mdro)
  )

n <- 93 #uncombine_network(res12_93$network) |>
  #length()

mytab <- data.table(
  Statistic = names(counts),
  Value     = as.integer(counts),
  Avg       = counts/n
)

mytab[, Statistic := gsub("b.starmix.{3}", "", Statistic)]
knitr::kable(
  mytab,
  caption = "Summary statistics of the model",
  digits = 2
)
```

```{r}
#| label: latex-table
#| echo: false
#| include: false
# Creating the latex table
mytab <- data.table(
  Statistic = paste("\\hspace{.5cm}", names(counts)),
  Value     = as.integer(counts),
  Avg       = counts/n
)

mytab <- xtable::xtable(mytab)

ans <- list(
  mytab[1:5,],
  mytab[6:11,],
  mytab[11:nrow(mytab),]
  )

lapply(ans, print, 
  include.colnames = FALSE,
  include.rownames = FALSE,
  only.contents = TRUE,
  hline.after = NULL,
  sanitize.text.function	= \(x) x
  ) |> invisible()

```

## Clustering sizes (do we want to group them all?)

```{r}
#| label: cluster-sizes
sizes <- lapply(networks, \(n) {
  data.table(
    net_id = n %v% "net_id",
    size = network.size(n),
    state = n %v% "state",
    nties = summary(n ~ edges)
  )
}) |>
  rbindlist() |>
  unique()

# sizes[, net_id := factor(net_id, levels = unique(net_id))]
# sizes[, size := factor(size, levels = sort(unique(size)))]

ggplot(sizes, aes(y = state, x = size)) +
  geom_density_ridges(
    jittered_points = TRUE, position = "raincloud",
    alpha = 0.9, scale = 0.9
  ) +
  # Annotate the number of networks in each state
  geom_text_repel(
    data = sizes[, .N, by = state],
    aes(x = 0, y = state, label = paste0("N=", N)),
    nudge_x = -0.5, size = 3, color = "black",
    nudge_y = 0.2
  ) +
  labs(
    x = "Network Size",
    y = "Count",
    title = "Distribution of Network Sizes by State"
  ) +
  theme_minimal()
```

Maryland has only 6 networks and the sizes are highly irregular. Let's take a look at how network size compares with the network density

```{r}
ggplot(sizes, aes(x = size, y = nties)) +
  geom_point() +
  facet_wrap(~state, ncol = 2) +
  geom_smooth(method = "lm") +
  labs(
    x = "Network Size",
    y = "Number of Ties",
    title = "Number of Ties vs Network Size"
  ) +
  theme_minimal()
```

## Degree distribution across types of nodes and sizes

```{r}
#| label: degree-distribution
degs <- lapply(networks, \(n) {
  hcw_type <- ifelse(
    n %v% "Nurse" == 1, "Nurse",
    ifelse(
      n %v% "PTOTRT" == 1,
      "PTOTRT",
      ifelse(
        n %v% "Other" == 1, "Other",
        ifelse(
          n %v% "CNA" == 1, "CNA",
          "Unknown"
        )
      )
    )
  )

  data.table(
    net_id = n %v% "net_id",
    degree = sna::degree(n),
    is_actor = n %v% "is_actor",
    size = network.size(n),
    state = n %v% "state",
    type_hcp = hcw_type
  )
}) |> rbindlist()

degs[is_actor == FALSE, type_hcp := "Patient"]
degs[, net_id := factor(net_id, levels = unique(net_id))]
degs[, size := factor(size, levels = sort(unique(size)))]
degs[, type_hcp := factor(type_hcp)]

degs[, is_actor := ifelse(is_actor, "HCP", "Patient")]

degs |>
  ggplot(aes(y = size, x=degree, fill = is_actor)) +
  geom_density_ridges(
    jittered_points = TRUE, position = "raincloud",
    alpha = 0.9, scale = 0.9
  ) +
  facet_wrap(~state, scales = "free_y", ncol = 2) +
  labs(
    x = "Degree",
    y = "Network Size",
    fill = "Node Type",
    title = "Degree distribution across node types and network sizes"
  ) +
  theme_minimal()
```

Checking the association with degree and size through a regression model

```{r}
degs[, size_num := as.integer(size)]
lm(degree ~ 
  state +
  size_num*state + I(size_num^2) + is_actor*size_num, data = degs) |>
  summary() |>
  texreg::screenreg(single.row = TRUE, digits = 2)
```

We can also compute the statistics

```{r}
degs[is_actor == "HCP", .(
  mean_degree = mean(degree/size_num),
  median_degree = median(degree/size_num),
  sd_degree = sd(degree/size_num)
), by = .(state)][order(mean_degree)] |>
  knitr::kable(
    caption = "Summary of degree distribution for HCPs across states (scaled by size)",
    digits = 2
  )
```

```{r}
degs[is_actor == "HCP"] |>
  ggplot(aes(y = type_hcp, x=degree)) +
  geom_density_ridges(
    jittered_points = TRUE, position = "raincloud",
    alpha = 0.9, scale = 0.9
  ) +
  facet_wrap(~state, scales = "free_y", ncol = 2) +
  labs(
    x = "Degree",
    y = "Network Size",
    title = "Degree distribution across node types and network sizes"
  ) +
  theme_minimal()
```

Let's now look at the concurrent distribution across sizes and states.

```{r}
dat_b2concurrent <- lapply(networks, \(n) {

  data.table(
    net_id = n %v% "net_id",
    b2concurrent = summary(n ~ b2concurrent()),
    size = network.size(n),
    state = n %v% "state"
  )
  
}) |>
  rbindlist() |>
  unique()

dat_b2concurrent[, net_id := factor(net_id, levels = unique(net_id))]
# dat_b2concurrent[, size := factor(size, levels = sort(unique(size)))]

dat_b2concurrent |>
  ggplot(aes(y = size, x=b2concurrent, color = state)) +
  geom_jitter() +
  geom_smooth(method = "lm")
```

There is a clear linear association with size, so this variable can be, in principle, captured without scaling by the log(n hcp).

## Final model

The following code chunk displays the results of the final multi-ERGM model fit. The `screenreg` function from the `texreg` package is used to present the estimated coefficients and their standard errors in a clear, publication-ready format.

```{r}
#| label: final-model
texreg::screenreg(
  res12_93,
  digits = 2
)
```

## Loading the GOF

In this section, we load the pre-computed goodness-of-fit (GOF) statistics for the final model. These statistics were generated during the model fitting process and are essential for evaluating how well the model reproduces key features of the observed networks. The code below loads the GOF object and produces diagnostic plots for a selection of network statistics, allowing for visual assessment of model fit across multiple dimensions.

```{r}
#| label: load-gof
gof12_93 <- readRDS("data/gofN_model12_93.rds")
```

```{r}
#| label: load-and-plot-gofN
#| fig-height: 6
#| eval: false

if (interactive())
  pdf("gof12_93-new.pdf")

op <- par(
  mfrow = c(4, 2),
  # Putting the x-labels inside of the plots
  oma = c(0, 0, 0, 0),
  mar = c(4, 4, 2, 1),
  # Reducing the size of the title
  cex.main = 0.8
  )
# figs <- autoplot(gof12_93, main = "{name} - {type}", id.n = 5)
plot(
  gof12_93, main = "{name}\n{type}", id.n = 5, , which = 1,
  xlab = ""
  )
par(op)

if (interactive())
  dev.off()

```

The next section explores whether the model could be improved by comparing observed versus predicted values for each statistic, rather than focusing solely on error terms. This provides a more direct assessment of the model's predictive accuracy for each network feature.

```{r}
#| label: gof-obs-vs-fitted
#| warning: false
#| fig-height: 6
vars <- names(gof12_93) 

if (interactive())
  pdf("gof12_93-obs-vs-fitted.pdf")
op <- par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
for (v in vars) {
  tmp_var <- gof12_93[[v]]

  tryCatch(with(tmp_var, {
    plot(observed, fitted, main = v, xlab = "Observed", ylab = "Fitted")
    abline(a = 0, b=1, col = "red")
  }), error = function(e) {
    if (interactive())
      message("Error in variable:", v, "\n")
  })
}
par(op)
if (interactive())
  dev.off()

```

## Visualizing some features of the networks

This section provides descriptive visualizations of key network features across the observed networks. We summarize and plot the distribution of network sizes, the number of healthcare workers (HCWs), and the mean degree of HCWs. These plots help contextualize the structure and composition of the networks under study, offering insight into the heterogeneity present in the data and supporting interpretation of the model results.

```{r}
#| label: network-feats
# Mean degree
netsapply <- function(f) {
  sapply(networks, \(x) {
    f(x)
  })
}
network_stats <- data.frame(
  size = sapply(networks, network.size),
  n_hcw = netsapply(\(x) {sum(x %v% "is_actor")}),
  n_mdro = netsapply(\(x) {sum((x %v% "mdro") == "Yes")}),
  n_dialysis = netsapply(\(x) {sum((x %v% "dialysis") == "Yes")}),
  n_wound_care = netsapply(\(x) {sum((x %v% "wound_care") == "Yes")}),
  n_ventilator = netsapply(\(x) {sum((x %v% "ventilator") == "Yes")}),
  n_bedridden = netsapply(\(x) {sum((x %v% "bedridden") == "Yes")}),
  # Now, tabulating the network composition
  n_nurses = netsapply(\(x) {sum((x %v% "Nurse") == 1)}),
  n_PTOTRT = netsapply(\(x) {sum((x %v% "PTOTRT") == 1)}),
  hcw_mean_degree = netsapply(\(x) {mean(sna::degree(x)[x %v% "is_actor"])})
)

# Size distribution
fig_size <- ggplot(network_stats, aes(x = size)) +
  geom_histogram() +
  labs(
    x = "Network Size",
    y = "Count",
    title = "Size Distribution"
  ) +
  theme_minimal()

# Size vs number of HCWs
fig_size_hcw <- ggplot(network_stats, aes(x = size, y = n_hcw)) +
  geom_point() +
  labs(
    x = "Network Size",
    y = "Number of HCWs",
    title = "Number of HCWs"
  ) +
  theme_minimal()

# Size vs HCW's mean degree
fig_size_hcw_degree <- ggplot(network_stats, aes(x = size, y = hcw_mean_degree)) +
  geom_point() +
  labs(
    x = "Network Size",
    y = "Mean Degree of HCWs",
    title = "Mean Degree of HCWs"
  ) +
  theme_minimal()

# Combine both figures into a single ggplot object
library(patchwork)
fig_size + fig_size_hcw + fig_size_hcw_degree
ggsave("size-distribution.png", width = 9*.75, height = 3*.75)

library(gridExtra)
```

## MCMC diagnostics

To assess the adequacy of the Markov Chain Monte Carlo (MCMC) estimation procedure, we present standard MCMC diagnostic plots for the final model.

```{r}
#| label: mcmc-diagnostics
#| fig-height: 10 
op <- par(mfrow = c(3, 2))
if (interactive())
  pdf("mcmc_diagnostics.pdf", width = 8, height = 10)
mcmc.diagnostics(res12_93, which = "plots")
par(op)
if (interactive())
  dev.off()
```

We further investigate model fit by examining the mean squared error (MSE) of the Pearson residuals for each network. This approach highlights the networks that are most poorly fitted by the model, allowing for targeted assessment of potential sources of misfit and guiding future model refinement.

```{r}
#| label: regress-networks
# Retrieving the errors
errs <- lapply(gof12_93, "[[", "pearson")

# Most commonly misfitted networks looking at the MSE
errs <- data.table(
  network = as.character(1:length(errs[[1]])),
  mse = do.call(cbind, errs) |>
    apply(1, \(x) mean(abs(x), na.rm = TRUE))
)

setorder(errs, mse)
errs[, network := factor(network, levels = network)]

# Plotting
if (interactive())
  pdf("errs.pdf")

ggplot(tail(errs, 20), aes(x = network, y = mse)) +
  geom_point() +
  geom_line() +
  labs(
    x = "Network",
    y = "Mean Squared Error",
    title = "Mean Squared Error of the Networks",
    subtitle = "(top 20)"
  ) +
  theme_minimal()

if (interactive())
  dev.off()
```

## Simulating and visualizing networks

Finally, we compare simulated networks generated from the fitted model to the observed networks. Simulating networks based on this model could be a powerful tool for simulation studies. We use the `netplot` R package to visualize the networks.

```{r}
#| label: simulate-networks
sims <- simulate(res12_93, nsim = 1, seed = 1234) |>
  splitnetwork(".NetworkID")

# Original networks
networks_0 <- res12_93$network |>
  splitnetwork(".NetworkID")
```

```{r}
#| label: plot-simulated-networks
library(netplot)
if (interactive())
  svg("example1.svg", width = 8, height = 4)
ans_row0 <- NULL
ans_row1 <- NULL
ans <- NULL

mynplot <- function(net, ...) {
  nplot(
    net, edge.line.breaks = 1, vertex.label = "", vertex.size = 1,
    vertex.nsides = ifelse(
      net %v% "is_actor", 3, 4
    ),
    vertex.color = ifelse(
      net %v% "is_actor", "black", "white"
    ),
    edge.color = "darkgrey",
    vertex.size.range = c(.05, .05, 1),
    ...
  )
}

for (i in c(6, 8, 9)) {
  set.seed(331)
  res0 <- mynplot(sims[[i]])
  res1 <- mynplot(networks_0[[i]])
  ans <- c(ans, list(res0, res1))
}

# Groups them by column (column major)
marrangeGrob(
  grobs = ans, ncol = 3, nrow = 2, padding = unit(.5, "line"),
  top = ""
  )

if (interactive())
  dev.off()
```
