---
format: docx
title: "Supplemental Materials for \"Exponential-Family Random Graph Models in Resident-Healthcare Provider Networks: An Application using data from Long-term Healthcare Facilities in the United States\""
author:
  - George G. Vega Yon
  - Chong Zhang
  - Karim Khader
  - Nai-Chung Nelson Chang
  - Lindsay D. Visnovsky
  - Alun Thomas
  - Candace Haroldsen
  - Kristina Stratford
  - Matthew H. Samore
date: 2025-04-01
date-modified: 2025-05-15
embed-resources: true
---


This document presents the supplemental materials for the main manuscript, focusing on the goodness-of-fit (GOF) assessment and diagnostics for the final multi-ERGM model. The following sections provide detailed explanations and outputs to facilitate reproducibility and transparency for readers interested in the modeling process and results.

The code chunks below are annotated to clarify their purpose and the rationale behind each step, following best practices for reporting ERGM-based social network analyses.

For terms wrapped in `ergm.multi::N()`, the subset argument is applied at the network level. For example, in `N(~b2starmix(2, "Nurse"), subset = Nurse)`, "Nurse" inside `b2starmix()` denotes the vertex-level HCW type indicator, whereas the object Nurse supplied to subset is a numeric vector of network IDs. This vector contains networks with at least two nurse HCWs, so the nurse-specific 2-star mixing term is evaluated only for networks in which that statistic is defined and informative. Similar network-index subsets were constructed for other attribute-specific terms.

## Summary stats

The following code chunk loads the relevant model objects and nodal attribute data required for the subsequent analyses. It also prepares a summary of the presence of multidrug-resistant organisms (MDROs) across the networks, which is a key covariate in the study. The summary_formula function is then used to compute the observed values for each model term, providing a basis for later comparison with model-based expectations.

```{r}
#| label: load-packages-and-data
# Loading data from the models
load('data/network93_f2.RData')
fits <- readRDS("models/03-pooled-ergms-final-results.rds")
fits_boot <- readRDS("models/04c-boot-ergms-results.rds")
source('models/helpers.R')
```

## Clustering sizes (do we want to group them all?)

```{r}
#| label: cluster-sizes
sizes <- lapply(networks, \(n) {
  data.table(
    net_id = n %v% "net_id",
    size = network.size(n),
    state = n %v% "state",
    nties = summary(n ~ edges)
  )
}) |>
  rbindlist() |>
  unique()

# sizes[, net_id := factor(net_id, levels = unique(net_id))]
# sizes[, size := factor(size, levels = sort(unique(size)))]

ggplot(sizes, aes(y = state, x = size)) +
  geom_density_ridges(
    jittered_points = TRUE, position = "raincloud",
    alpha = 0.9, scale = 0.9
  ) +
  # Annotate the number of networks in each state
  geom_text_repel(
    data = sizes[, .N, by = state],
    aes(x = 0, y = state, label = paste0("N=", N)),
    nudge_x = -0.5, size = 3, color = "black",
    nudge_y = 0.2
  ) +
  labs(
    x = "Network Size",
    y = "Count",
    title = "Distribution of Network Sizes by State"
  ) +
  theme_minimal()

if (interactive())
  ggsave("models/05-supplemental-materials-cluster-sizes.png", width = 10, height = 6)
```

Maryland has only 6 networks and the sizes are highly irregular. Let's take a look at how network size compares with the network density

```{r}
ggplot(sizes, aes(x = size, y = nties)) +
  geom_point() +
  facet_wrap(~state, ncol = 2) +
  geom_smooth(method = "lm") +
  labs(
    x = "Network Size",
    y = "Number of Ties",
    title = "Number of Ties vs Network Size"
  ) +
  theme_minimal()

if (interactive())
  ggsave("models/05-supplemental-materials-nties-vs-size.png", width = 10, height = 6)
```

## Degree distribution across types of nodes and sizes

```{r}
#| label: degree-distribution
degs <- lapply(networks, \(n) {
  hcw_type <- ifelse(
    n %v% "Nurse" == 1, "Nurse",
    ifelse(
      n %v% "PTOTRT" == 1,
      "PTOTRT",
      ifelse(
        n %v% "Other" == 1, "Other",
        ifelse(
          n %v% "CNA" == 1, "CNA",
          "Unknown"
        )
      )
    )
  )

  data.table(
    net_id = n %v% "net_id",
    degree = sna::degree(n),
    is_actor = n %v% "is_actor",
    size = network.size(n),
    state = n %v% "state",
    type_hcp = hcw_type,
    type = n %v% "type"
  )
}) |> rbindlist()

degs[type %in% c("Physician", "PA/NP"), type := "Other"]

degs[is_actor == FALSE, type_hcp := "Patient"]
degs[, net_id := factor(net_id, levels = unique(net_id))]
degs[, size := factor(size, levels = sort(unique(size)))]
degs[, type_hcp := factor(type_hcp)]

degs[, is_actor := ifelse(is_actor, "HCP", "Patient")]

degs |>
  ggplot(aes(y = size, x=degree, fill = is_actor)) +
  geom_density_ridges(
    jittered_points = TRUE, position = "raincloud",
    alpha = 0.9, scale = 0.9
  ) +
  facet_wrap(~state, scales = "free_y", ncol = 2) +
  labs(
    x = "Degree",
    y = "Network Size (sorted by, not at scale)",
    fill = "Node Type",
    title = "Degree distribution across node types and network sizes"
  ) +
  theme_minimal()

if (interactive())
  ggsave("models/05-supplemental-materials-degree-distribution.png", width = 6, height = 10)
```

We can also compute the statistics

```{r}
degs[, size_num := as.numeric(size)]
degs[is_actor == "HCP", .(
  mean_degree = mean(degree/size_num),
  median_degree = median(degree/size_num),
  sd_degree = sd(degree/size_num),
  ci = sprintf(
    "[%.2f, %.2f]",
    quantile(degree/size_num, probs = c(0.025)),
    quantile(degree/size_num, probs = c(0.975))
  )
), by = .(state)][order(mean_degree)] |>
  knitr::kable(
    caption = "Summary of degree distribution for HCPs across states (scaled by size)",
    digits = 2
  )
```

```{r}
degs |>
  ggplot(aes(y = state, x=degree)) +
  geom_density_ridges(
    jittered_points = TRUE, position = "raincloud",
    alpha = 0.9, scale = 0.9
  ) +
  facet_wrap(~type, scales = "free_y", ncol = 2) +
  labs(
    x = "Degree",
    y = "Network Size",
    title = "Degree distribution across node types and network sizes"
  ) +
  theme_minimal()

if (interactive())
  ggsave("models/05-supplemental-materials-degree-distribution-by-type.png", width = 6, height = 6)
```

Let's now look at the concurrent distribution across sizes and states.

```{r}
dat_b2concurrent <- lapply(networks, \(n) {

  data.table(
    net_id = n %v% "net_id",
    b2concurrent = summary(n ~ b2concurrent()),
    size = network.size(n),
    state = n %v% "state"
  )
  
}) |>
  rbindlist() |>
  unique()

dat_b2concurrent[, net_id := factor(net_id, levels = unique(net_id))]
# dat_b2concurrent[, size := factor(size, levels = sort(unique(size)))]

dat_b2concurrent |>
  ggplot(aes(y = size, x=b2concurrent, color = state)) +
  geom_jitter() +
  geom_smooth(method = "lm")

if (interactive())
  ggsave("models/05-supplemental-materials-b2concurrent.png", width = 6, height = 6)
```

There is a clear linear association with size, so this variable can be, in principle, captured without scaling by the log(n hcp).

## Final model

The following code chunk displays the results of the final multi-ERGM model fit. The `screenreg` function from the `texreg` package is used to present the estimated coefficients and their standard errors in a clear, publication-ready format.

```{r}
#| label: final-model
naive_fit <- fits$gwdsp_with_log_patients
texreg::screenreg(
  # file = "models/04-boot-ergms-results.html",
  list(
    'Naive model'=naive_fit,
    'Bootstrap model'=naive_fit
    ),
  override.se = list(NULL, fits_boot$bootstrap_sds_no_intercept),
  override.pvalues = list(NULL, fits_boot$pvalues_no_intercept),
  single.row = TRUE
)
```

## Loading the GOF

In this section, we load the pre-computed goodness-of-fit (GOF) statistics for the final model. These statistics were generated during the model fitting process and are essential for evaluating how well the model reproduces key features of the observed networks. The code below loads the GOF object and produces diagnostic plots for a selection of network statistics, allowing for visual assessment of model fit across multiple dimensions.

```{r}
#| label: load-gof
fn <- "models/05-supplemental-materials-gof.rds"
if (file.exists(fn)) {
  gof_final <- readRDS(fn)
  gof_with_state <- gof_final$with_state
  gof_final <- gof_final$final
}else {
  gof_final <- gofN(naive_fit)
  gof_with_state <- gofN(fits[[1]])
  saveRDS(
    list(with_state = gof_with_state, final = gof_final),
    fn
    )
}
```

```{r}
#| label: load-and-plot-gofN
#| fig-height: 6
#| eval: false

if (interactive())
  pdf("models/05-supplemental-materials-gof.pdf")

op <- par(
  mfrow = c(4, 2),
  # Putting the x-labels inside of the plots
  oma = c(0, 0, 0, 0),
  mar = c(4, 4, 2, 1),
  # Reducing the size of the title
  cex.main = 0.8
  )
# figs <- autoplot(gof12_93, main = "{name} - {type}", id.n = 5)
plot(
  gof_final, main = "{name}\n{type}", id.n = 5, , which = 1,
  xlab = ""
  )
par(op)

if (interactive())
  dev.off()

```

The next section explores whether the model could be improved by comparing observed versus predicted values for each statistic, rather than focusing solely on error terms. This provides a more direct assessment of the model's predictive accuracy for each network feature.

```{r}
#| label: gof-obs-vs-fitted
#| warning: false
#| fig-height: 6
vars <- names(gof_final) 

# Removing the b1dsp* & offset terms variables
vars <- vars[!grepl("b1dsp|b.degree0", vars)]

if (interactive())
  pdf("models/05-supplemental-materials-gof-obs-vs-fitted.pdf")
op <- par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
for (v in vars) {
  tmp_var <- gof_final[[v]]

  tryCatch(with(tmp_var, {
    plot(observed, fitted, main = v, xlab = "Observed", ylab = "Fitted")
    abline(a = 0, b=1, col = "red")
  }), error = function(e) {
    if (interactive())
      message("Error in variable:", v, "\n")
  })
}
par(op)
if (interactive())
  dev.off()

```

## Visualizing some features of the networks

This section provides descriptive visualizations of key network features across the observed networks. We summarize and plot the distribution of network sizes, the number of healthcare workers (HCWs), and the mean degree of HCWs. These plots help contextualize the structure and composition of the networks under study, offering insight into the heterogeneity present in the data and supporting interpretation of the model results.

```{r}
#| label: network-feats
# Mean degree
netsapply <- function(f) {
  sapply(networks, \(x) {
    f(x)
  })
}
network_stats <- data.frame(
  size = sapply(networks, network.size),
  n_hcw = netsapply(\(x) {sum(x %v% "is_actor")}),
  n_mdro = netsapply(\(x) {sum((x %v% "mdro") == "Yes")}),
  n_dialysis = netsapply(\(x) {sum((x %v% "dialysis") == "Yes")}),
  n_wound_care = netsapply(\(x) {sum((x %v% "wound_care") == "Yes")}),
  n_ventilator = netsapply(\(x) {sum((x %v% "ventilator") == "Yes")}),
  n_bedridden = netsapply(\(x) {sum((x %v% "bedridden") == "Yes")}),
  # Now, tabulating the network composition
  n_nurses = netsapply(\(x) {sum((x %v% "Nurse") == 1)}),
  n_PTOTRT = netsapply(\(x) {sum((x %v% "PTOTRT") == 1)}),
  hcw_mean_degree = netsapply(\(x) {mean(sna::degree(x)[x %v% "is_actor"])})
)

# Size distribution
fig_size <- ggplot(network_stats, aes(x = size)) +
  geom_histogram() +
  labs(
    x = "Network Size",
    y = "Count",
    title = "Size Distribution"
  ) +
  theme_minimal()

# Size vs number of HCWs
fig_size_hcw <- ggplot(network_stats, aes(x = size, y = n_hcw)) +
  geom_point() +
  labs(
    x = "Network Size",
    y = "Number of HCWs",
    title = "Number of HCWs"
  ) +
  theme_minimal()

# Size vs HCW's mean degree
fig_size_hcw_degree <- ggplot(network_stats, aes(x = size, y = hcw_mean_degree)) +
  geom_point() +
  labs(
    x = "Network Size",
    y = "Mean Degree of HCWs",
    title = "Mean Degree of HCWs"
  ) +
  theme_minimal()

# Combine both figures into a single ggplot object
library(patchwork)
fig_size + fig_size_hcw + fig_size_hcw_degree
ggsave("models/05-supplemental-materials-network-feats.png", width = 9*.75, height = 3*.75)

```

## MCMC diagnostics

To assess the adequacy of the Markov Chain Monte Carlo (MCMC) estimation procedure, we present standard MCMC diagnostic plots for the final model.

```{r}
#| label: mcmc-diagnostics
#| fig-height: 10 
op <- par(mfrow = c(3, 2))
if (interactive())
  pdf("models/05-supplemental-materials-mcmc-diagnostics.pdf", width = 8, height = 10)
mcmc.diagnostics(naive_fit, which = "plots")
par(op)
if (interactive())
  dev.off()
```

We further investigate model fit by examining the mean squared error (MSE) of the Pearson residuals for each network. This approach highlights the networks that are most poorly fitted by the model, allowing for targeted assessment of potential sources of misfit and guiding future model refinement.

```{r}
#| label: regress-networks
# Retrieving the errors
plot_gof <- function(gof_data) {
  errs <- lapply(gof_data, "[[", "pearson")

  # Removing the b1dsp* & offset terms variables
  to_exclude <- which(!grepl("b1dsp|b.degree", names(errs)))
  errs <- errs[to_exclude]

  errs <- as.data.frame(errs) |>
    data.table()

  # Obs
  obs_val <- lapply(gof_data, "[[", "observed")
  obs_val <- obs_val[to_exclude]

  obs_val <- as.data.frame(obs_val) |>
    data.table()

  obs_val[, id := 1:.N]

  # Fitted
  fitted_val <- lapply(gof_data, "[[", "fitted")
  fitted_val <- fitted_val[to_exclude]

  fitted_val <- as.data.frame(fitted_val) |>
    data.table()

  fitted_val[, id := 1:.N]

  # Getting the state and size of each network
  errs$id <- 1:length(errs[[1]])
  errs$state <- sapply(networks, \(x) {unique(x %v% "state")})
  errs$size <- sapply(networks, network.size)
  errs$n_hcp <- sapply(networks, \(x) {sum(x %v% "is_actor")})
  errs$n_resident <- sapply(networks, \(x) {sum(!x %v% "is_actor")})

  # Reshape long (from wide)
  errs <- melt(
    errs, id.vars = c("id", "state", "size", "n_hcp", "n_resident"),
    variable.name = "term", value.name = "error"
    )

  obs_val <- melt(obs_val, id.vars = c("id"), variable.name = "term", value.name = "observed")

  fitted_val <- melt(fitted_val, id.vars = c("id"), variable.name = "term", value.name = "fitted")

  errs <- merge(errs, obs_val, by = c("id", "term"))
  errs <- merge(errs, fitted_val, by = c("id", "term"))

  # Turning into the sqrt(abs)
  errs$error <- sqrt(abs(errs$error))

  fig <- ggplot(errs[term == "N.1..edges"], aes(x = size, y = error)) +
    geom_jitter(aes(color = state), height = 0) +
    labs(
      x = "Network Size",
      # The y-label should be the math expression
      # for the sqrt(abs(Std. Pearson Residual))
      y = expression(sqrt(abs("Std. Pearson Res. (edges)"))),
      color = "State"
    ) +
    # Adding a lowess line by state
    geom_smooth(aes(color=state), se = FALSE, method="lm") +
    theme_minimal() +
    # Need to label the outlier observation that is outside the value 2
    # using ggrepel
    geom_text_repel(
      aes(label = ifelse(error > 2, id, ""), color=state),
      show.legend = FALSE
      )

  list(
    fig = fig,
    dat = errs
  )

}

gg_err <- plot_gof(gof_final)
gg_err_states <- plot_gof(gof_with_state)

if (interactive()) {
  ggsave( 
    "models/05-supplemental-materials-errs.pdf", 
    plot = gg_err$fig,
    width = 8, height = 3
    )
} else
  print(gg_err$fig)
```

We can also test whether the error in edges prediction can be explained by the state of the network (trying to find significant differences in the error across different states). Since we know that the networks are not idependent, we will use a bootstrap approach to account for this:

```{r}
set.seed(3312)
dat_err_edges <- gg_err_states$dat[term == "N.1..edges"]

# Initial fit
boot_fit <- lm(error ~ factor(state), data = dat_err_edges)

# Bootstrap
n_obs <- nrow(dat_err_edges)
boot_err <- replicate(
  1000, {
    idx <- sample(n_obs, n_obs, replace = TRUE)
    lm(error ~ factor(state), data = dat_err_edges[idx, ]) |> 
      coef()
})

# Collecting the results
boot_err <- do.call(rbind, list(boot_err)) |>
  t() 

# Computing sds
boot_err_sds <- apply(boot_err, 2, sd)

# Calculating pvalues
boot_err_pvals <- 2 * (1 - pnorm(abs(coef(boot_fit)) / boot_err_sds))

# 1. Extract original coefficients (excluding the Intercept)
# We only want to test the 'state' variables
beta_orig <- coef(boot_fit)[-1] 

# 2. Compute the variance-covariance matrix of the bootstrap estimates
# We exclude the first column (the Intercept) to match beta_orig
boot_vcov <- cov(boot_err[, -1])

# 3. Compute the Wald Chi-Square Statistic
# Formula: W = t(beta) %*% solve(Sigma) %*% beta
W_stat <- t(beta_orig) %*% solve(boot_vcov) %*% beta_orig

# 4. Degrees of freedom = number of parameters being tested
df <- length(beta_orig)

# 5. Compute the p-value using the Chi-Square distribution
omnibus_pval <- pchisq(as.numeric(W_stat), df = df, lower.tail = FALSE)

# Print results
cat("Bootstrap Omnibus Wald Test\n")
cat("---------------------------\n")
cat("Chi-Square Statistic:", as.numeric(W_stat), "\n")
cat("Degrees of Freedom:  ", df, "\n")
cat("P-value:             ", omnibus_pval, "\n")
```

The resulting output, the Chi2-statistic is 13.51 with a p-value of 0.035, suggesting that the model as a whole is statistically significant.

## Simulating and visualizing networks

Finally, we compare simulated networks generated from the fitted model to the observed networks. Simulating networks based on this model could be a powerful tool for simulation studies. We use the `netplot` R package to visualize the networks.

```{r}
#| label: simulate-networks
fn <- "models/05-supplemental-materials-sims.rds"
if (!file.exists(fn)) {
  sims <- simulate(naive_fit, nsim = 1, seed = 1234) |>
    ergmito::splitnetwork(".NetworkID")
  saveRDS(sims, fn)
} else {
  sims <- readRDS(fn)
}

# Original networks
networks_0 <- networks
```

```{r}
#| label: plot-simulated-networks
library(netplot)
if (interactive())
  svg("models/05-supplemental-materials-example1.svg", width = 8, height = 4)
ans_row0 <- NULL
ans_row1 <- NULL
ans <- NULL

mynplot <- function(net, ...) {
  nplot(
    net, edge.line.breaks = 1, vertex.label = "", vertex.size = 1,
    vertex.nsides = ifelse(
      net %v% "is_actor", 3, 4
    ),
    vertex.color = ifelse(
      net %v% "is_actor", "black", "white"
    ),
    edge.color = "darkgrey",
    vertex.size.range = c(.05, .05, 1),
    ...
  )
}

for (i in c(6, 8, 9)) {
  set.seed(331)
  res0 <- mynplot(sims[[i]])
  res1 <- mynplot(networks_0[[i]])
  ans <- c(ans, list(res0, res1))
}

# Groups them by column (column major)
gridExtra::marrangeGrob(
  grobs = ans, ncol = 3, nrow = 2, padding = unit(.5, "line"),
  top = ""
  )

if (interactive())
  dev.off()
```
