---
title: "Untitled"
output:
  pdf_document:
    extra_dependencies: pdflscape
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE, warning=FALSE}
library(network)
library(igraph)
library(data.table)
library(ergm)
library(netplot)
```

# Data Preparation

```{r read-clean}
# Reading the data in
dat <- fread("data/Matched Interaction and Resident.csv")

# HCW data
cnames <- colnames(dat)
cnames[grepl("hcw", cnames, ignore.case = TRUE)]
```

Creating a few covariates to be used in the ERGM

```{r}
dat[, white := fifelse(!is.na(race), as.integer(race == "White"), NA_integer_)]
# Wound Care
# Diabetes
# Dialysis
# Wheelchair
# Bedridden
# Ventilator
# MDRO
dat[, wound_care := fifelse(!is.na(wound_care), as.integer(wound_care == "Yes"), NA_integer_)]
dat[, diabetes := fifelse(!is.na(diabetes), as.integer(diabetes == "Yes"), NA_integer_)]
dat[, dialysis := fifelse(!is.na(dialysis), as.integer(dialysis == "Yes"), NA_integer_)]
dat[, wheelchair := fifelse(!is.na(wheelchair), as.integer(wheelchair == "Yes"), NA_integer_)]
dat[, bedridden := fifelse(!is.na(bedridden), as.integer(bedridden == "Yes"), NA_integer_)]
dat[, ventilator := fifelse(!is.na(ventilator), as.integer(ventilator == "Yes"), NA_integer_)]
dat[, mdro := fifelse(!is.na(mdro), as.integer(mdro == "Yes"), NA_integer_)]
```

The following function will be used to create unique identifiers for each network.

```{r}
# Function to collapse multiple fields into a single string
# keeping only alphanumeric characters.
namemilize <- function(...) {
  do.call(
    paste,
    c(lapply(list(...), \(x) gsub("[^[:alnum:]]", "_", trimws(x))),
      list(sep = "-"))
  ) |>
    tolower() |>
    gsub(pattern = "_+", replacement = "_")
  
}

# Getting nodes
dat[, loc_id := namemilize(state, date, Fac_Name, Fac_Unit)]
```

We noticed that for some records (7) the resident ID coincides with the room id:

```{r}
dat[room_num == res_id, .(room_num, res_id)]
```

For now, we will remove those records

```{r}
dat <- dat[room_num != res_id]
```

Some of the health variables have missing data or are inconsistent,
showing 0/1/NA values in the same day. We will take the most frequent value, if no NAs
are registered.

```{r}
f <- \(d) {
  tmp <- table(d, useNA = "no")
  rep(as.integer(names(tmp[which.max(tmp)])), length(d))
}

dat[, wound_care := f(wound_care), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, diabetes := f(diabetes), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, dialysis := f(dialysis), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, wheelchair := f(wheelchair), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, bedridden := f(bedridden), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, ventilator := f(ventilator), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, mdro := f(mdro), by = .(state, date, Fac_Name, Fac_Unit, res_id)]

# There are some other missings, so we will replace them by the individual (regardless)
# of the day
dat[, wound_care := f(wound_care), by = .(state, Fac_Name, Fac_Unit, res_id)]
dat[, mdro := f(mdro), by = .(state, Fac_Name, Fac_Unit, res_id)]

# Still a couple of missings... just 0 for now
dat[, wound_care := fcoalesce(wound_care, 0L)]
dat[, mdro := fcoalesce(mdro, 0L)]
```

Now that we have idividual ids for the unit/location, we can start creating the
network data:

```{r}
# Edgelist
edges <- unique(dat[, .(loc_id, hcw_id, res_id)])

# Nodes dataset
nodes <- rbind(
  dat[, .(
    id = hcw_id,
    loc_id, state, Fac_Name, Fac_Unit,
    white = 0L,
    race = "", res_age = 0,
    type = hcw_type, group = "hcw",
    is_actor = TRUE,
    wound_care = 0L, diabetes = 0L, dialysis = 0L, wheelchair = 0L, bedridden = 0L, ventilator = 0L, mdro = 0L)],
  dat[, .(
    id = res_id,
    loc_id, state, Fac_Name, Fac_Unit,
    white,
    race, res_age,
    type = "resident", group = "residents",
    is_actor = FALSE,
    wound_care, diabetes, dialysis, wheelchair, bedridden, ventilator, mdro)]
) |> unique()
```

## Duplicated records (hcw)

```{r check-dpls}
setorder(nodes, loc_id, id)
nodes[, n := .N, by = .(loc_id, id)][n > 1 & group == "hcw", .(id, type)]
```

We can see in the data that the duplicated records here (currently 12) are in
its majority if type "Nurse" or "CNA" (certified nurse assistant.) We will 
replace those with "Nurse."

```{r check-dpls-replace}
nodes[
  n > 1 & group == "hcw" & any(type %in% c("Nurse", "CNA")),
  type := "Nurse",
  by = .(loc_id, id)
  ]
```

## Duplicated records (residents)

```{r check-dpls-residents}
nodes[n > 1 & group == "residents"][1:10 ,.(id, race, res_age)]
```

Most of the duplicates are in age. For now we will take the mean and use that
as the observed age. Notice that, while most cases the duplicates have ages that
show little differences, e.g., one or two years, there are some cases with large
age gaps, e.g., 20 years.

```{r check-dpls-residents-fix}
nodes[
  n > 1 & group == "residents",
  res_age := floor(mean(res_age, na.rm = TRUE)),
  by = .(loc_id, id)
  ]
```

## Checking no duplicates exists

After the data pre-processing, let's check whether the duplicates are now gone

```{r}
nodes <- unique(nodes)
nodes[, n := .N, by = .(loc_id, id)]
table(nodes[, n])
```

Still many observations that are duplicated twice each:

```{r}
nodes[n > 1, .(id, race, n, loc_id)]
```

In this case, the observations have differences in race, e.g., the first of them
registers both "White" and "Black or African American." Since we cannot corretly
identify which is the correct, we will set them as "NK - Not Known."

```{r}
nodes[n > 1, c("race", "white") := list("NK - Not Known", 2L)]
nodes <- unique(nodes)
nodes[, n := NULL]
```

Observations with missing race or age will be droped.

```{r}
nodes <- nodes[!is.na(race) & !is.na(res_age)]
```

Finally, we need to make sure that edges match the nodes

```{r}
edges <- merge(
  edges, nodes[, .(id, loc_id, here_hcw = TRUE)],
  by.x = c("hcw_id", "loc_id"), 
  by.y = c("id", "loc_id"), all.x = TRUE, all.y = FALSE
)

edges <- merge(
  edges, nodes[, .(id, loc_id, here_res = TRUE)],
  by.x = c("res_id", "loc_id"), 
  by.y = c("id", "loc_id"), all.x = TRUE, all.y = FALSE
)

edges[, here_hcw := fcoalesce(here_hcw, FALSE)]
edges[, here_res := fcoalesce(here_res, FALSE)]
edges <- edges[here_hcw & here_res]

edges[, c("here_hcw", "here_res") := c(NULL, NULL)]
edges <- unique(edges)
```

A couple of extra variables

```{r}
nodes[, sqrt_age := sqrt(res_age)]
```


# Building the networks

```{r build-nets}
# Making room to store the networks, we will keep two types of
# networks: HCW - Patient
networks <- vector("list", length(unique(nodes$loc_id)))
names(networks) <- unique(nodes$loc_id)
networks_proj <- networks

nerrors <- 0L
for (n in names(networks)) {
  
  edges_n <- edges[loc_id == n, .(hcw_id, res_id)]
  
  nodes_hcw_n <- nodes[loc_id == n & group == "hcw"]
  nodes_res_n <- nodes[loc_id == n & group == "residents"]
  
  # Trying to read the data
  networks[[n]] <- tryCatch({
    suppressWarnings(as.network(
      x         = edges_n,
      vertices  = rbind(nodes_hcw_n, nodes_res_n),
      bipartite = TRUE,
      bipartite_col = "is_actor"
      ))
  }, error = function(e) e)
  
  # Notification!
  if (inherits(networks[[n]], what = "error")) {
    message("Network ", n, " failed to be built")
    nerrors <- nerrors + 1L
    next
  }
  

}

message("Done! ", nerrors, "/", length(networks), " failed.")
```

```{r viz}
# Example of a graph
set.seed(1231)
nplot(
  networks[[2L]],
  edge.curvature=0,
  vertex.color = c(
    hcw = "blue", residents = "green"
    )[networks[[1L]] %v% "group"]
  )
```

# Fitting ERGMs

## Single network ERGMs

First, we start with structural models

```{r ergms, message=FALSE, warning=FALSE}
# Projection networks ----------------------------------------------------------
log_n_inv <- log(1/network.size(networks[[1L]]))
res_b12star <- ergm(
  networks[[1L]] ~ edges +
    Offset(~Sum(log_n_inv ~ edges, label = "a label"), 1) +
    b1star(2)
  )

res_b22star <- ergm(
  networks[[1L]] ~ edges +
    Offset(~Sum(log_n_inv ~ edges, "k"), 1) +
    b2star(2)
  )

res_b12starb22star <- ergm(
  networks[[1L]] ~ edges +
    Offset(~Sum(log_n_inv ~ edges, "k"), 1) +
    b1star(2) + b2star(2)
  )
```

```{r struct-table, results='asis', echo = FALSE}
texreg::knitreg(
  list(
    "2-b1star" = res_b12star,
    "2-b2star" = res_b22star,
    "3-b1star" = res_b12starb22star
    )
)
```

## Single model with multiple networks

```{r multi-to-single}
bignet <- Map(\(n,e) {
  
  # Creating the baseline data
  el <- list(
    edges = as.data.frame(n, unit = "edges"),
    # Second column matches the first, so we remove it
    nodes = as.data.frame(n, unit = "vertices") #[,-1L] 
  )
  
  # Appending the id of the dataset to avoid duplicates
  el$edges$.head <- paste0(e, ".", el$edges$.head)
  el$edges$.tail <- paste0(e, ".", el$edges$.tail)
  el$nodes$vertex.names <- paste0(e, ".", el$nodes$vertex.names)
  
  el$nodes$nvertex <- nrow(el$nodes)
  
  # Returning
  el
}, n = networks, e = names(networks))

bignet <- list(
  edges = data.table(do.call(rbind, lapply(bignet, "[[", "edges"))),
  nodes = data.table(do.call(rbind, lapply(bignet, "[[", "nodes")))
)

bignet_mat <- as.network(
  x         = bignet$edges,
  vertices  = bignet$nodes,
  bipartite = TRUE,
  bipartite_col = "is_actor"
)

# Comparing we are getting the same thing
get.inducedSubgraph(bignet_mat, which((bignet_mat %v% "loc_id") == names(networks)[1]))
networks[[1]]
```

Common controls for all models

```{r}
default_control <- list(
  seed            = 44485,
  main.method     = "Stochastic-Approximation",
  MCMC.samplesize	= 1024 * 4 #,
  # MCMC.interval   = 1024 * 4,
  # parallel        = 2L
)
```

We can also use Krivitsky (2011) offset. To do this we use the new functions included
in `ergm` 4.0, in particular, the `Sum()` term. With it, we can create an offset
in the form:

$$
\text{Krivitsky2011}(\mathbf{G}) = \sum_{i} \text{edges}_i\times\ln\left({1/n_i}\right)
$$

Where $n_i$ is the number of vertices in network $i$. To achieve this, we will
(a) compute the values $\ln\left({1/n_i}\right)$ separately, and (b) use that
vector to pre-multiply the vector of edgecounts at the graph level. We can get
an edgecount at the graph level using the `b1factor()` term with the option
`levels = TRUE`. The following codeblock achieves this:

```{r}
# log(1/n) part of the term
log_inv_n <- log(summary(bignet_mat ~ degrange(0, by = "loc_id")))

# Let's see how it looks
summary(
  bignet_mat ~
    edges +
    Sum("sum" ~ Sum(log_inv_n ~ b1factor("loc_id", levels=TRUE), "a"), "krivitsky2011" )
)
```

Let's see how it is implemented in an ERGM. Remember that we need to set the 
term as an offset:

```{r model0, cache = TRUE, message=FALSE, warning=FALSE}
res_big_ergm0 <- ergm(
  bignet_mat ~ edges +
    Sum("sum" ~ Sum(log_inv_n ~ b1factor("loc_id", levels=TRUE), "a"), "edgesxlog_n"),
  control = do.call("control.ergm", c(default_control)),
  constraints = ~ blockdiag("loc_id")
  )
```

Interpretation of the `edges` term needs to be done in the context of this offset
term. For example, since $\theta_{edges} =$ `r sprintf("%.2f", coef(res_big_ergm0)[1])`,
the density in a Bernoulli graph with 100 nodes is
$\text{logit}^{-1}\left(\theta_{edges} + \ln\left(1/100\right)\right)\approx$
`r sprintf("%.2f", plogis(-log(100) + coef(res_big_ergm0[1])))`. Looking into
stars from the second level (one resident -- two healthcare workers.)

```{r model1, cache = TRUE, eval = TRUE, warning=FALSE, message=FALSE}
res_big_ergm1 <- ergm(
  bignet_mat ~ edges +
    Sum("sum" ~ Sum(log_inv_n ~ b1factor("loc_id", levels=TRUE), "a"), "edgesxlog_n") +
    b2star(2),
  control = do.call("control.ergm", c(default_control)),
  constraints = ~ blockdiag("loc_id")
  )
```


Second model, Using homophily in race:

```{r model2, cache = TRUE, message=FALSE, warning=FALSE}
res_big_ergm2 <- ergm(
  bignet_mat ~ edges + 
    Sum("sum" ~ Sum(log_inv_n ~ b1factor("loc_id", levels=TRUE), "a"), "edgesxlog_n") +
    b2star(2) +  
    Sum("sum"~Sum(log_inv_n ~ b2starmix(2, attr = "loc_id", diff = FALSE), "a"), "b2star x log_n"),
  control = do.call("control.ergm", default_control),
  constraints = ~ blockdiag("loc_id")
  )
```


```{r model3, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE}
res_big_ergm3 <- ergm(
  bignet_mat ~ edges + 
    Sum("sum" ~ Sum(log_inv_n ~ b1factor("loc_id", levels=TRUE), "a"), "edgesxlog_n") +
    b2star(2) + 
    b1starmix(2, "white", diff = FALSE),
  control = do.call("control.ergm", default_control),
  constraints = ~ blockdiag("loc_id")
  )
```

```{r model4, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE}
res_big_ergm4 <- ergm(
  bignet_mat ~ edges + 
    Sum("sum" ~ Sum(log_inv_n ~ b1factor("loc_id", levels=TRUE), "a"), "edgesxlog_n") +
    b2star(2) + 
    b1starmix(2, "white", diff = FALSE) +
    b2starmix(2, "type"),
  control = do.call("control.ergm", default_control),
  constraints = ~ blockdiag("loc_id")
  )
```

```{r model5, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE}
res_big_ergm5 <- ergm(
  bignet_mat ~ edges + 
    Sum("sum" ~ Sum(log_inv_n ~ b1factor("loc_id", levels=TRUE), "a"), "edgesxlog_n") +
    b2star(2) + 
    b1starmix(2, "white", diff = FALSE) +
    b2starmix(2, "type") + 
    b1dsp(1:2) +
    b2cov("sqrt_age"),
  control = do.call("control.ergm", default_control),
  constraints = ~ blockdiag("loc_id")
  )
```

```{r model6, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE}
res_big_ergm6 <- ergm(
  bignet_mat ~ edges + 
    Sum("sum" ~ Sum(log_inv_n ~ b1factor("loc_id", levels=TRUE), "a"), "edgesxlog_n") +
    b2star(2) + 
    b1starmix(2, "white", diff = FALSE) +
    b2starmix(2, "type") + 
    b1dsp(1:2) +
    b2cov("sqrt_age") +
    b1starmix(2, "wound_care") + # 3 missing
    b1starmix(2, "diabetes") +
    b1starmix(2, "dialysis") +
    b1starmix(2, "wheelchair") +
    b1starmix(2, "bedridden") +
    b1starmix(2, "ventilator") +
    b1starmix(2, "mdro"),  # 5 missing
  control = do.call("control.ergm", default_control),
  constraints = ~ blockdiag("loc_id")
  )
```

```{r models-print-big, results='asis', echo=FALSE, message=FALSE, warning=FALSE, eval = TRUE}
cat("\\begin{landscape}\n\n")
texreg::knitreg(
  list(
    res_big_ergm0,
    res_big_ergm1,
    res_big_ergm2,
    res_big_ergm3,
    res_big_ergm4,
    res_big_ergm5
    )
)
cat("\\end{landscape}\n\n")
```


Another big model

```{r results='asis', echo = FALSE}
texreg::knitreg(
  res_big_ergm6
)
```


# Goodness of fit

For computing the goodness of fit, we need to take each model and calculate the
expected value of the observed sufficient statistics. Model 5 (ergm4) seems to be one
of the best fitting models, so let's start with that one. We will compute the
standardized pearson residual as in Krivitstky, Coletti, and Hans (2022):

```{r pearson_std_res}
# Matching residents
pearson_std_res <- function(net., model., seed. = 54656, nsim. = 1000) {
  
  fm     <- formula(model.)
  
  fm_new <- update.formula(fm, net. ~ .)
  environment(fm_new) <- environment()
  
  # Retrieving the list of possible terms
  obs_stats <- summary(fm_new)

  stats. <- names(obs_stats)
  stats. <- stats.[!grepl("krivits", stats.)]
  
  # Simulating
  res <- simulate(
    fm_new,
    coef = coef(model.)[stats.],
    output = "stats",
    nsim = nsim.,
    seed = seed.
    )
  
  to_include <- intersect(colnames(res), stats.)
  
  data.frame(
    s_name    = to_include,
    s_obs     = obs_stats[to_include],
    s_sim_avg = colMeans(res[,to_include]),
    s_sim_std = sqrt(diag(cov(res[,to_include])))
  )
  
}

```


```{r computing-resids, cache=TRUE}
pearson_residuals <- parallel::mclapply(
  networks, pearson_std_res, model. = res_big_ergm2,
  nsim. = 1e3, mc.cores = 8L
  )
pearson_residuals <- Map(
  \(n,dat) cbind(dat, net = n),
  n = names(pearson_residuals), dat = pearson_residuals)
pearson_residuals <- rbindlist(pearson_residuals)
pearson_residuals <- pearson_residuals[s_obs != 0]
pearson_residuals[, t_stat := (s_obs - s_sim_avg)/s_sim_std]
```

```{r viz-residuals}
library(ggplot2)
ggplot(pearson_residuals, aes(y = t_stat, x = s_name)) +
  geom_violin() +
  labs(x = "Statistic", y = "Pearson Resid.") +
  coord_flip() 
```

```{r viz-residuals-edges}
library(ggplot2)
ggplot(pearson_residuals[s_name == "edges"]) +
  geom_point(
    aes(
      y = t_stat,
      x = sqrt(s_sim_avg)
    )
  ) +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x = expression(sqrt(Fitted~Values~(edges))), y = "Pearson Resid.")
```


```{r viz-residuals-b2star2}
ggplot(pearson_residuals[s_name == "b2star2"]) +
  geom_point(
    aes(
      y = t_stat,
      x = sqrt(s_sim_avg)
    )
  ) +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x = expression(sqrt(Fitted~Values~(b2star2))), y = "Pearson Resid.")
```
