---
title: "Untitled"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE, warning=FALSE}
library(network)
library(igraph)
library(data.table)
library(ergm)
library(netplot)
```

# Data Preparation

```{r read-clean}
# Reading the data in
dat <- fread("data/Matched Interaction and Resident.csv")

# HCW data
cnames <- colnames(dat)
cnames[grepl("hcw", cnames, ignore.case = TRUE)]
```

```{r}
# Function to collapse multiple fields into a single string
# keeping only alphanumeric characters.
namemilize <- function(...) {
  do.call(
    paste,
    c(lapply(list(...), \(x) gsub("[^[:alnum:]]", "_", trimws(x))),
      list(sep = "-"))
  ) |>
    tolower() |>
    gsub(pattern = "_+", replacement = "_")
  
}

# Getting nodes
dat[, loc_id := namemilize(state, date, Fac_Name, Fac_Unit)]
```

We noticed that for some records (7) the resident ID coincides with the room id:

```{r}
dat[room_num == res_id, .(room_num, res_id)]
```

For now, we will remove those records

```{r}
dat <- dat[room_num != res_id]
```


Now that we have idividual ids for the unit/location, we can start creating the
network data:

```{r}
# Edgelist
edges <- dat[, .(
  loc_id, state, date, Fac_Name, Fac_Unit, hcw_id, hcw_type, res_id,
  res_age, race
  )]

# Nodes dataset
nodes <- rbind(
  edges[, .(
    id = hcw_id,
    loc_id, state, Fac_Name, Fac_Unit,
    race = "", res_age = 0,
    type = hcw_type, group = "hcw")],
  edges[, .(
    id = res_id,
    loc_id, state, Fac_Name, Fac_Unit,
    race, res_age,
    type = "resident", group = "residents")]
) |> unique()
```

## Duplicated records (hcw)

```{r check-dpls}
setorder(nodes, loc_id, id)
nodes[, n := .N, by = .(loc_id, id)][n > 1 & group == "hcw", .(id, type)]
```

We can see in the data that the duplicated records here (currently 12) are in
its majority if type "Nurse" or "CNA" (certified nurse assistant.) We will 
replace those with "Nurse."

```{r check-dpls-replace}
nodes[
  n > 1 & group == "hcw" & any(type %in% c("Nurse", "CNA")),
  type := "Nurse",
  by = .(loc_id, id)
  ]
```

## Duplicated records (residents)

```{r check-dpls-residents}
nodes[n > 1 & group == "residents"][1:10 ,.(id, race, res_age)]
```

Most of the duplicates are in age. For now we will take the mean and use that
as the observed age. Notice that, while most cases the duplicates have ages that
show little differences, e.g., one or two years, there are some cases with large
age gaps, e.g., 20 years.

```{r check-dpls-residents-fix}
nodes[
  n > 1 & group == "residents",
  res_age := floor(mean(res_age, na.rm = TRUE)),
  by = .(loc_id, id)
  ]
```

## Checking no duplicates exists

After the data pre-processing, let's check whether the duplicates are now gone

```{r}
nodes <- unique(nodes)
nodes[, n := .N, by = .(loc_id, id)]
table(nodes[, n])
```

Still ten (actually five) observations that are duplicated twice each:

```{r}
nodes[n > 1, .(id, race, n)]
```

In this case, the observations have differences in race, e.g., the first of them
registers both "White" and "Black or African American." Since we cannot corretly
identify which is the correct, we will set them as "NK - Not Known."

```{r}
nodes[n > 1, race := "NK - Not Known"]
nodes <- unique(nodes)
nodes[, n := NULL]
```

Observations with missing race or age will be droped.

```{r}
nodes <- nodes[!is.na(race) & !is.na(res_age)]
```

Finally, we need to make sure that edges match the nodes

```{r}
edges <- merge(
  edges, nodes[, .(id, loc_id, here_hcw = TRUE)],
  by.x = c("hcw_id", "loc_id"), 
  by.y = c("id", "loc_id"), all.x = TRUE, all.y = FALSE
)

edges <- merge(
  edges, nodes[, .(id, loc_id, here_res = TRUE)],
  by.x = c("res_id", "loc_id"), 
  by.y = c("id", "loc_id"), all.x = TRUE, all.y = FALSE
)

edges[, here_hcw := fcoalesce(here_hcw, FALSE)]
edges[, here_res := fcoalesce(here_res, FALSE)]
edges <- edges[here_hcw & here_res]

edges[, c("here_hcw", "here_res") := c(NULL, NULL)]
```


# Building the networks

```{r build-nets}
# Making room to store the networks, we will keep two types of
# networks: HCW - Patient
networks <- vector("list", length(unique(nodes$loc_id)))
names(networks) <- unique(nodes$loc_id)
networks_proj <- networks

nerrors <- 0L
for (n in names(networks)) {
  
  edges_n <- edges[loc_id == n, .(hcw_id, res_id)]
  
  nodes_hcw_n <- nodes[loc_id == n & group == "hcw"]
  nodes_res_n <- nodes[loc_id == n & group == "residents"]
  
  # Trying to read the data
  networks[[n]] <- tryCatch(matrix(
    0L, nrow = nrow(nodes_hcw_n), ncol = nrow(nodes_res_n),
    dimnames = list(nodes_hcw_n$id, nodes_res_n$id)
    ))
  
  # Notification!
  if (inherits(networks[[n]], what = "error")) {
    message("Network ", n, " failed to be built")
    nerrors <- nerrors + 1L
    next
  }
  
  # Building the bipartite network
  networks[[n]][as.matrix(edges_n)] <- 1L
  networks[[n]] <- network(
    networks[[n]], bipartite = TRUE,
    vertex.attr = rbind(nodes_hcw_n, nodes_res_n)
    )
  
}

message("Done! ", nerrors, "/", length(networks), " failed.")
```

```{r viz}
# Example of a graph
set.seed(1231)
nplot(
  networks[[1L]],
  edge.curvature=0,
  vertex.color = c(
    hcw = "blue", residents = "green"
    )[networks[[1L]] %v% "group"]
  )
```

# Fitting ERGMs

## Single network ERGMs

First, we start with structural models

```{r ergms, message=FALSE, warning=FALSE}
# Projection networks ----------------------------------------------------------
res_2star <- ergm(networks[[1L]] ~ edges + b1star(2))
res_2and3star <- ergm(networks[[1L]] ~ edges + b1star(2:3))
res_3star <- ergm(networks[[1L]] ~ edges + b1star(3))
```

```{r struct-table, results='asis', echo = FALSE}
texreg::knitreg(
  list(
    "2-b1star" = res_2star,
    "2 and 3-b1star" = res_2and3star,
    "3-b1star" = res_3star
    )
)
```

## Single model with multiple networks

```{r multi-to-single}
bignet <- Map(\(n,e) {
  
  # Creating the baseline data
  el <- list(
    edges = as.data.frame(n, unit = "edges"),
    # Second column matches the first, so we remove it
    nodes = as.data.frame(n, unit = "vertices")[,-1L] 
  )
  
  # Appending the id of the dataset to avoid duplicates
  el$edges$.head <- paste0(el$edges$.head, e)
  el$edges$.tail <- paste0(el$edges$.tail, e)
  el$nodes$id <- paste0(el$nodes$id, e)
  
  # Returning
  el
}, n = networks, e = names(networks))

bignet <- list(
  edges = data.table(do.call(rbind, lapply(bignet, "[[", "edges"))),
  nodes = data.table(do.call(rbind, lapply(bignet, "[[", "nodes")))
)

bignet_mat <- with(bignet, matrix(
  0L,
  nrow = nrow(nodes[group == "hcw"]),
  ncol = nrow(nodes[group == "residents"]),
  dimnames = list(nodes[group == "hcw"]$id, nodes[group == "residents"]$id)
))

bignet_mat[with(bignet$edges, cbind(.head, .tail))] <- 1L

bignet_mat <- network(
  bignet_mat,
  vertex.attr = bignet$nodes, bipartite = TRUE
)
```

<!-- Since the resulting network is large, we will start by looking at a sample subset: -->


<!-- ```{r eval = FALSE} -->
<!-- set.seed(8828) -->
<!-- bignet_sub <- sample(names(networks), size = 20) -->
<!-- bignet_sub <- which(bignet_mat %v% "loc_id" %in% bignet_sub) -->
<!-- bignet_sub <- get.inducedSubgraph(bignet_mat, v = bignet_sub) -->

<!-- nplot( -->
<!--   bignet_sub, -->
<!--   vertex.nsides = c( -->
<!--     hcw = 3, residents = 4 -->
<!--   )[bignet_mat %v% "group"], -->
<!--   vertex.color = c( -->
<!--     hcw = "blue", residents = "green" -->
<!--     )[bignet_mat %v% "group"], -->
<!--   edge.curvature = 0, -->
<!--   edge.line.breaks = 1L, -->
<!--   skip.vertex = TRUE -->
<!--   ) -->
<!-- ``` -->

Common controls for all models

```{r}
default_control <- list(
  main.method     = "MCMLE",
  MCMC.samplesize	= 1024 * 4,
  # MCMC.interval   = 1024 * 4,
  parallel        = 2L
)
```



```{r model0, cache = TRUE, message=FALSE, warning=FALSE}
res_big_ergm0 <- ergm(
  bignet_mat ~ edges,
  control = do.call("control.ergm", c(default_control, list(init = c(-2)))),
  constraints = ~ blockdiag("loc_id")
  )
```



```{r model1, cache = TRUE, message=FALSE, warning=FALSE}
res_big_ergm1 <- ergm(
  bignet_mat ~ edges + b1nodematch("race"),
  control = do.call("control.ergm", c(default_control, list(init = c(-2,2)))),
  constraints = ~ blockdiag("loc_id")
  )
```

```{r model2, cache = TRUE, message=FALSE, warning=FALSE}
res_big_ergm2 <- ergm(
  bignet_mat ~ edges + b1nodematch("race") + b2star(2),
  control = do.call("control.ergm", c(default_control, list(init = c(-2,2,2)))),
  constraints = ~ blockdiag("loc_id")
  )
```


```{r models-print-big, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
texreg::knitreg(
  list(res_big_ergm0,res_big_ergm1,res_big_ergm2)
)
```



