---
title: "Bipartite networks of health care workers and residents in long-term care facilities"
output:
  pdf_document:
    extra_dependencies:
      - pdflscape
      - tabularx
      - booktabs
      - geometry
      - graphicx
editor_options: 
  markdown: 
    wrap: 72
author: George G. Vega Yon
date: May 3, 2022
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Data Preparation

For this analysis we will use the following R packages:

- network: For manipulation of network data.

- data.table: For data cleaning and processing.

- ergm: For estimation and simulation of Exponential Random Graph Models.

- netplot: For visualization of networks.

- ggplot2: For general visualization.

- gridExtra: Utilities for grid graphics (base of ggplot2 and netplot.)

```{r packages, message=FALSE, warning=FALSE}
library(network)
library(data.table)
library(ergm)
library(netplot)
library(ggplot2)
library(gridExtra)
```

We start with reading the data.

```{r read-clean}
# Reading the data in
dat <- fread("data/Matched Interaction and Resident.csv")
```

Creating a few covariates to be used in the ERGM. We also need to deal with yes/no
variables to make sure these are represented as 0/1 integers in the data.

```{r}
dat[, white := fifelse(!is.na(race), as.integer(race == "White"), NA_integer_)]
# Wound Care
# Diabetes
# Dialysis
# Wheelchair
# Bedridden
# Ventilator
# MDRO
dat[, wound_care := fifelse(!is.na(wound_care), as.integer(wound_care == "Yes"), NA_integer_)]
dat[, diabetes := fifelse(!is.na(diabetes), as.integer(diabetes == "Yes"), NA_integer_)]
dat[, dialysis := fifelse(!is.na(dialysis), as.integer(dialysis == "Yes"), NA_integer_)]
dat[, wheelchair := fifelse(!is.na(wheelchair), as.integer(wheelchair == "Yes"), NA_integer_)]
dat[, bedridden := fifelse(!is.na(bedridden), as.integer(bedridden == "Yes"), NA_integer_)]
dat[, ventilator := fifelse(!is.na(ventilator), as.integer(ventilator == "Yes"), NA_integer_)]
dat[, mdro := fifelse(!is.na(mdro), as.integer(mdro == "Yes"), NA_integer_)]
```

The following function will be used to create unique identifiers for
each network. The function replaces non-alphanumeric characters from the data
with underscore and combines multiple columns into a single string.

```{r}
# Function to collapse multiple fields into a single string
# keeping only alphanumeric characters.
namemilize <- function(...) {
  do.call(
    paste,
    c(lapply(list(...), \(x) gsub("[^[:alnum:]]", "_", trimws(x))),
      list(sep = "-"))
  ) |>
    tolower() |>
    gsub(pattern = "_+", replacement = "_")
  
}

# Generating the network ID
dat[, net_id := namemilize(state, date, Fac_Name, Fac_Unit)]
```

We noticed that for some records (7) the resident ID coincides with the
room id. For now, we will remove those records:

```{r}
dat[room_num == res_id, .(room_num, res_id)]
dat <- dat[room_num != res_id]
```

Some of the health variables have missing data or are inconsistent,
showing 0/1/NA values in the same day. We will take the most frequent
value, if no NAs are registered.

```{r}
f <- \(d) {
  tmp <- table(d, useNA = "no")
  rep(as.integer(names(tmp[which.max(tmp)])), length(d))
}

dat[, wound_care := f(wound_care), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, diabetes := f(diabetes), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, dialysis := f(dialysis), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, wheelchair := f(wheelchair), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, bedridden := f(bedridden), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, ventilator := f(ventilator), by = .(state, date, Fac_Name, Fac_Unit, res_id)]
dat[, mdro := f(mdro), by = .(state, date, Fac_Name, Fac_Unit, res_id)]

# There are some other missings, so we will replace them by the individual (regardless)
# of the day
dat[, wound_care := f(wound_care), by = .(state, Fac_Name, Fac_Unit, res_id)]
dat[, mdro := f(mdro), by = .(state, Fac_Name, Fac_Unit, res_id)]

# Still a couple of missings... just 0 for now
dat[, wound_care := fcoalesce(wound_care, 0L)]
dat[, mdro := fcoalesce(mdro, 0L)]
```

Now that we have idividual ids for the unit/location, we can start
creating the network data:

```{r}
# Edgelist
edges <- unique(dat[, .(net_id, hcw_id, res_id)])

# Nodes dataset
nodes <- rbind(
  dat[, .(
    id = hcw_id,
    net_id, state, Fac_Name, Fac_Unit,
    white = 0L,
    race = "", res_age = 0,
    type = hcw_type, group = "hcw",
    is_actor = TRUE,
    wound_care = 0L, diabetes = 0L, dialysis = 0L, wheelchair = 0L, bedridden = 0L, ventilator = 0L, mdro = 0L)],
  dat[, .(
    id = res_id,
    net_id, state, Fac_Name, Fac_Unit,
    white,
    race, res_age,
    type = "resident", group = "residents",
    is_actor = FALSE,
    wound_care, diabetes, dialysis, wheelchair, bedridden, ventilator, mdro)]
) |> unique()
```

The next steps will be to check for duplicates in the data, in particular, at the
node level.

## Duplicated records (hcw)

```{r check-dpls}
setorder(nodes, net_id, id)
nodes[, n := .N, by = .(net_id, id)][n > 1 & group == "hcw", .(id, type)]
```

We can see in the data that the duplicated records here (currently 12)
are in its majority if type "Nurse" or "CNA" (certified nurse
assistant.) We will replace those with "Nurse."

```{r check-dpls-replace}
nodes[
  n > 1 & group == "hcw" & any(type %in% c("Nurse", "CNA")),
  type := "Nurse",
  by = .(net_id, id)
  ]
```

## Duplicated records (residents)

```{r check-dpls-residents}
nodes[n > 1 & group == "residents"][1:10 ,.(id, res_age)]
```

Most of the duplicates are in age. For now we will take the mean and use
that as the observed age. Notice that, while most cases the duplicates
have ages that show little differences, e.g., one or two years, there
are some cases with large age gaps, e.g., 20 years.

```{r check-dpls-residents-fix}
nodes[
  n > 1 & group == "residents",
  res_age := floor(mean(res_age, na.rm = TRUE)),
  by = .(net_id, id)
  ]
```

## Checking no duplicates exists

After the data pre-processing, let's check whether the duplicates are
now gone

```{r}
nodes <- unique(nodes)
nodes[, n := .N, by = .(net_id, id)]
table(nodes[, n])
```

Still many observations that are duplicated twice each:

```{r}
nodes[n > 1, .(id, race, n)]
```

In this case, the observations have differences in race, e.g., the first
of them registers both "White" and "Black or African American." Since we
cannot correctly identify which is the correct, we will set them as "NK -
Not Known."

```{r}
nodes[n > 1, c("race", "white") := list("NK - Not Known", 2L)]
nodes <- unique(nodes)
nodes[, n := NULL]
```

Observations with missing race or age will be dropped.

```{r}
nodes <- nodes[!is.na(race) & !is.na(res_age)]
```

Finally, we need to make sure that edges match the nodes

```{r}
edges <- merge(
  edges, nodes[, .(id, net_id, here_hcw = TRUE)],
  by.x = c("hcw_id", "net_id"), 
  by.y = c("id", "net_id"), all.x = TRUE, all.y = FALSE
)

edges <- merge(
  edges, nodes[, .(id, net_id, here_res = TRUE)],
  by.x = c("res_id", "net_id"), 
  by.y = c("id", "net_id"), all.x = TRUE, all.y = FALSE
)

edges[, here_hcw := fcoalesce(here_hcw, FALSE)]
edges[, here_res := fcoalesce(here_res, FALSE)]
edges <- edges[here_hcw & here_res]

edges[, c("here_hcw", "here_res") := c(NULL, NULL)]
edges <- unique(edges)
```

A couple of extra variables, for now, the square-root of age, which we will use
in the ERMG model.

```{r}
nodes[, sqrt_age := sqrt(res_age)]
```

# Building the networks

The following code of block will iterate through all network ids and create one
network per case. Right now we are counting 100 networks:

```{r build-nets}
# Making room to store the networks, we will keep two types of
# networks: HCW - Patient
networks <- vector("list", length(unique(nodes$net_id)))
names(networks) <- unique(nodes$net_id)
networks_proj <- networks

nerrors <- 0L
for (n in names(networks)) {
  
  # Selecting the edges and nodes relevant for the network
  edges_n <- edges[net_id == n, .(hcw_id, res_id)]
  
  nodes_hcw_n <- nodes[net_id == n & group == "hcw"]
  nodes_res_n <- nodes[net_id == n & group == "residents"]
  
  # Trying to read the data
  networks[[n]] <- tryCatch({
    suppressWarnings(as.network(
      x         = edges_n,
      vertices  = rbind(nodes_hcw_n, nodes_res_n),
      bipartite = TRUE,
      bipartite_col = "is_actor"
      ))
  }, error = function(e) e)
  
  # Notification!
  if (inherits(networks[[n]], what = "error")) {
    message("Network ", n, " failed to be built")
    nerrors <- nerrors + 1L
    next
  }
  

}

# Reporting how many went OK
message("Done! ", nerrors, "/", length(networks), " failed.")
```

Here are four networks:

```{r viz}
# Example of a graph
set.seed(1231)

net_figs <- vector("list", 4L)
for (i in 1:4)
  net_figs[[i]] <- nplot(
    networks[[i]],
    edge.curvature=0,
    vertex.color = c(
      hcw = "blue", residents = "green"
      )[networks[[i]] %v% "group"]
    )

# Putting them together using the gridExtra package
gridExtra::grid.arrange(grobs=net_figs, nrow = 2, ncol = 2) 
```


# Fitting ERGMs

For the baseline models, we will be using the terms included in table 1. Before we continue, we will now concatenate all networks into a single large graph to be used in with the ERGM package. Although what we want is to fit a pooled-bipartite ERGM, as of today the `ergm` package has no way of dealing with multiple networks. Nonetheless, mathematically, block-diagonal ERGMs--in which ties are only permitted within groups--are equivalent to pooled ERGMs.

\def\tmpfigwidth{\includegraphics[width=.5\linewidth]}
\begin{table}[h]
	\begin{tabular}{m{.4\linewidth}m{.2\linewidth}m{.3\linewidth}<\centering}
		\toprule
		Name & Representation & Form \\ \midrule
		Edge count & \tmpfigwidth{fig/term-edgecount.png} &$\sum_{i<j}y_{ij}$ \\\midrule
		HCW-centered two-star & \tmpfigwidth{fig/term-actor-twostar.png} & $\sum_{i\in H, \{j<k\}\in R}y_{ij}y_{ik}$ \\\midrule
		Resident-centered two-star &\tmpfigwidth{fig/term-entity-twostar.png} & $\sum_{i\in R, \{j<k\}\in H}y_{ji}y_{ki}$ \\\midrule
		HCW-centered homophilic two-star & \tmpfigwidth{fig/term-actor-twostar-mix.png} & $\sum_{i\in H, \{j<k\}\in R}\mathbf{1}\left(x_{j} = x_k\right) y_{ij}y_{ik}$ \\\midrule
		Resident-centered homophilic two-star & \tmpfigwidth{fig/term-entity-twostar-mix.png} & $\sum_{i\in R, \{j<k\}\in H}\mathbf{1}\left(x_{j} = x_k\right)y_{ji}y_{ki}$ \\\midrule
		Dyad-wise one shared residents &  \tmpfigwidth{fig/term-actor-one-edgewise-shared-partner.png} & $\sum_{\{i<j\}\in H, k\in R}y_{ik}y_{jk}$ \\ \midrule
		Dyad-wise two shared residents & \tmpfigwidth{fig/term-actor-two-edgewise-shared-partner.png} & $\sum_{\{i<j\}\in H, \{k<h\}\in R}y_{ik}y_{jk}y_{ih}y_{jh}$ \\ 
		\bottomrule
	\end{tabular}
	\label{tab:terms}
	\caption{Terms used in the bipartite ERGM models}
\end{table}
\def\tmpfigwidth{}

## Structural terms ERGM

```{r multi-to-single}
bignet <- Map(\(n,e) {
  
  # Creating the baseline data
  el <- list(
    edges = as.data.frame(n, unit = "edges"),
    # Second column matches the first, so we remove it
    nodes = as.data.frame(n, unit = "vertices") #[,-1L] 
  )
  
  # Appending the id of the dataset to avoid duplicates
  el$edges$.head <- paste0(e, ".", el$edges$.head)
  el$edges$.tail <- paste0(e, ".", el$edges$.tail)
  el$nodes$vertex.names <- paste0(e, ".", el$nodes$vertex.names)
  
  el$nodes$nvertex <- nrow(el$nodes)
  
  # Returning
  el
}, n = networks, e = names(networks))

bignet <- list(
  edges = data.table(do.call(rbind, lapply(bignet, "[[", "edges"))),
  nodes = data.table(do.call(rbind, lapply(bignet, "[[", "nodes")))
)

bignet_mat <- as.network(
  x         = bignet$edges,
  vertices  = bignet$nodes,
  bipartite = TRUE,
  bipartite_col = "is_actor"
)

# Comparing we are getting the same thing
get.inducedSubgraph(bignet_mat, which((bignet_mat %v% "net_id") == names(networks)[1]))
networks[[1]]
```

Common controls for all models:

```{r}
default_control <- list(
  seed            = 44485,
  main.method     = "Stochastic-Approximation",
  MCMC.samplesize	= 1024 * 4 #,
  # MCMC.interval   = 1024 * 4,
  # parallel        = 2L
)
```

We can could use Krivitsky (2011) offset, which in the case of multiple
graphs would be $\sum_{i} \text{edges}_i\times\ln\left({1/n_i}\right)$,
where $n_i$ is the number of nodes in network $i$. Nonetheless, in Butts
(2013) it is suggested to use a free parameter
$\sum_i\ln\left(n_i\right)$. To do this we use the new functions
included in `ergm` 4.0, in particular, the `Sum()` term. Two steps to do
this:

1.  Compute the values $\ln\left({n_i}\right)$ separately, and

2.  Use that vector to pre-multiply the vector of edgecounts at the
    graph level. We can get an edgecount at the graph level using the
    `b1factor()` term with the option `levels = TRUE`. The following
    codeblock achieves this:

```{r}
# log(n) part of the term.
log_n <- log(summary(bignet_mat ~ degrange(0, by = "net_id")))

# Let's see how it looks
summary(
  bignet_mat ~
    edges +
    Sum("sum" ~ Sum(log_n ~ b1factor("net_id", levels=TRUE), "a"), "log_n" )
)
```

Let's see how it is implemented in an ERGM. Remember that we need to set
the term as an offset:

```{r model0, cache = TRUE, message=FALSE, warning=FALSE}
res_big_ergm0 <- ergm(
  bignet_mat ~ edges +
    Sum("sum" ~ Sum(log_n ~ b1factor("net_id", levels=TRUE), "a"), "edgesxlog_n"),
  control = do.call("control.ergm", c(default_control)),
  constraints = ~ blockdiag("net_id")
  )
```

```{r model0-tabel, echo = FALSE, results='asis'}
texreg::knitreg(list(`Model 0` = res_big_ergm0))
```

Interpretation of the `edges` term needs to be done in the context of
this offset term. For example, since $\theta_{edges} =$
`r sprintf("%.2f", coef(res_big_ergm0)[1])`, the density in a Bernoulli
graph with 100 nodes is
$\text{logit}^{-1}\left(\theta_{edges} + \theta_{logn}\ln\left(100\right)\right)\approx$
`r sprintf("%.2f", plogis(coef(res_big_ergm0[1]) %*% c(1, log(100))))`.
As in Krivitsky et al. (2022), we can scale terms across the model in
such a fashion. Let's repeat the exercise with the b2 2-star (second
level 2-stars) term:

```{r model1, cache = TRUE, eval = TRUE, warning=FALSE, message=FALSE, results='hide'}
res_big_ergm1 <- ergm(
  bignet_mat ~ edges +
    Sum("sum" ~ Sum(log_n ~ b1factor("net_id", levels=TRUE), "a"), "edgesxlog_n") +
    b2star(2) + 
    Sum("sum"~Sum(log_n ~ b2starmix(2, attr = "net_id", diff = FALSE), "a"), "b2star x log_n"),
  control = do.call("control.ergm", c(default_control)),
  constraints = ~ blockdiag("net_id")
  )
```

```{r model1-tabel, echo = FALSE, results='asis'}
texreg::knitreg(list(`Model 1` = res_big_ergm1))
```

In this case, the interaction with $log(n)$ did not improved our model,
so we will keep only the three other significant terms. Following
Krivitsky et al (2022,) we can leverage having multiple networks to
asses which other terms would be beneficial to be included in our model. The following
function will help us compute the predicted values per network (including the
standard error and observed values, which we will later use for goodness-of-fit
analyses.)

```{r calc_ergm_pearson}
# Matching residents
calc_ergm_pearson <- function(net., model., seed. = 54656, nsim. = 1000L) {
  
  fm    <- formula(model.)
  log_n <- log(network::network.size(net.))
  
  fm_new <- update.formula(fm, net. ~ .)
  environment(fm_new) <- environment()
  
  # Retrieving the list of possible terms
  obs_stats <- summary(fm_new)

  stats. <- names(obs_stats)
  
  # Simulating
  res <- simulate(
    fm_new,
    coef = coef(model.)[stats.],
    output = "stats",
    nsim = nsim.,
    seed = seed.
    )
  
  to_include <- intersect(colnames(res), stats.)
  
  data.frame(
    s_name    = to_include,
    s_obs     = obs_stats[to_include],
    s_sim_avg = colMeans(res[,to_include]),
    s_sim_std = sqrt(diag(cov(res[,to_include])))
  )
  
}
```

To do this, we will perform the following steps:

1. Compute the observed ERGM pearson standardized residuals,

2. Compute the observed motif (term) count that were not included in the model

3. Run a linear regression predicting the residuals as a function of the terms
not included. Terms that were not included in the model that turn out to be
siginificant will be then included in the ERGM:

```{r model0-regressions-step1, cache = TRUE}
library(parallel)
cl <- makePSOCKcluster(4) # To make things a bit faster
invisible({
  clusterEvalQ(cl, library(ergm))
})

# Computing the residuals per network
ergm_pearson <- parLapply(
  cl = cl, 
  X  = networks,
  fun = calc_ergm_pearson,
  model. = res_big_ergm1
  )

stopCluster(cl)

# Binding the results together
ergm_pearson <- rbindlist(ergm_pearson)
ergm_pearson[, pearson := (s_obs - s_sim_avg)/s_sim_std]

# Adding an id
ergm_pearson$id <- sapply(names(networks), rep,  times = 4L, simplify = FALSE) |>
  unlist()
ergm_pearson
```

Computing now the observed statistics of the following terms:

- 3 and 4 star in both levels: `b1star(3:4)` and `b2star(3:4)`
- Residents race-homophily in the second level: `b1starmix(2, "white", diff = FALSE)`
- HWC type homophily: `b2starmix(2, "type")`
- One and two shared residents per worker: `b1dsp(1:2)`
- Covariate effect of residents age: `b2cov("sqrt_age")`

```{r model0-regressions-step2}
ergm_pearson_terms <- lapply(networks, \(n) {
  res <- summary(
    n ~ b1star(3:4) + b2star(3:4) + b1starmix(2, "white", diff = FALSE) +
      b2starmix(2, "type") + b1dsp(1:2) + b2cov("sqrt_age")
    )
  
  data.table(setNames(rbind(res), names(res)))
  }) |> rbindlist(fill = TRUE, )

# Replacing missing with zeros. Missings are due to not observing those
# categories in the networks
ergm_pearson_terms <- ergm_pearson_terms[, lapply(.SD, fcoalesce, 0)]
```

Finally, we can take this data and fit an OLS

```{r model0-regressions-step3}
ergm_pearson_wide <- dcast(ergm_pearson[,.(id,s_name, pearson)], id~...)
ergm_pearson_wide <- cbind(
  ergm_pearson_wide,
  ergm_pearson_terms
)

ergm_pearson_b2star2 <- lm(
  b2star2 ~ 
    b1star3 + b1star4 + b2star3 + b2star4 + 
    `b1starmix.2.white.0` + `b2starmix.2.type.resident.CNA` +
    `b2starmix.2.type.resident.Nurse` + `b1dsp1` + `b1dsp2` + `b2cov.sqrt_age` +
    `b2starmix.2.type.resident.PT/OT` +
    `b2starmix.2.type.resident.RT` + `b2starmix.2.type.resident.Other` +
    `b2starmix.2.type.resident.Physician` + `b2starmix.2.type.resident.PA/NP`,
  data = ergm_pearson_wide
)

ergm_pearson_edges <- lm(
  edges ~
    b1star3 + b1star4 + b2star3 + b2star4 + 
    `b1starmix.2.white.0` + `b2starmix.2.type.resident.CNA` +
    `b2starmix.2.type.resident.Nurse` + `b1dsp1` + `b1dsp2` + `b2cov.sqrt_age` +
    `b2starmix.2.type.resident.PT/OT` +
    `b2starmix.2.type.resident.RT` + `b2starmix.2.type.resident.Other` +
    `b2starmix.2.type.resident.Physician` + `b2starmix.2.type.resident.PA/NP`,
  data = ergm_pearson_wide
)
```

```{r echo = FALSE, results='asis'}
texreg::knitreg(
  list(
    b2star = ergm_pearson_b2star2,
    edges = ergm_pearson_edges
    )
)
```

## More covariate-based effects

Second model, Using homophily in race:

```{r model2, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE, results='hide'}
res_big_ergm2 <- ergm(
  bignet_mat ~ edges + 
    Sum("sum" ~ Sum(log_n ~ b1factor("net_id", levels=TRUE), "a"), "edgesxlog_n") +
    b2star(2) + 
    b1starmix(2, "white", diff = FALSE),
  control = do.call("control.ergm", default_control),
  constraints = ~ blockdiag("net_id")
  )
```

```{r model3, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE, results='hide'}
res_big_ergm3 <- ergm(
  bignet_mat ~ edges + 
    Sum("sum" ~ Sum(log_n ~ b1factor("net_id", levels=TRUE), "a"), "edgesxlog_n") +
    b2star(2) + 
    b1starmix(2, "white", diff = FALSE) +
    b2starmix(2, "type"),
  control = do.call("control.ergm", default_control),
  constraints = ~ blockdiag("net_id")
  )
```

```{r model4, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE, results='hide'}
res_big_ergm4 <- ergm(
  bignet_mat ~ edges + 
    Sum("sum" ~ Sum(log_n ~ b1factor("net_id", levels=TRUE), "a"), "edgesxlog_n") +
    b2star(2) + 
    b1starmix(2, "white", diff = FALSE) +
    b2starmix(2, "type") + 
    b1dsp(1:2) +
    b2cov("sqrt_age"),
  control = do.call("control.ergm", default_control),
  constraints = ~ blockdiag("net_id")
  )
```

```{r model5, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE, results='hide'}
res_big_ergm5 <- ergm(
  bignet_mat ~ edges + 
    Sum("sum" ~ Sum(log_n ~ b1factor("net_id", levels=TRUE), "a"), "edgesxlog_n") +
    b2star(2) + 
    b1starmix(2, "white", diff = FALSE) +
    b2starmix(2, "type") + 
    b1dsp(1:2) +
    b2cov("sqrt_age") +
    b1starmix(2, "wound_care", diff = FALSE) + # 3 missing
    b1starmix(2, "diabetes", diff = FALSE) +
    b1starmix(2, "dialysis", diff = FALSE) +
    b1starmix(2, "wheelchair", diff = FALSE) +
    b1starmix(2, "bedridden", diff = FALSE) +
    b1starmix(2, "ventilator", diff = FALSE) +
    b1starmix(2, "mdro", diff = FALSE),  # 5 missing
  control = do.call("control.ergm", default_control),
  constraints = ~ blockdiag("net_id")
  )
```

```{r models-print-big, results='asis', echo=FALSE, message=FALSE, warning=FALSE, eval = TRUE}
cat("\\begin{landscape}\n\n")
texreg::knitreg(
  setNames(list(
    res_big_ergm0,
    res_big_ergm1,
    res_big_ergm2,
    res_big_ergm3,
    res_big_ergm4
    ), paste("Model", 0:4))
)
cat("\\end{landscape}\n\n")
```

```{r echo=FALSE, results='asis'}
texreg::knitreg(list("Model 5" = res_big_ergm5))
```

Finally, before proceeding to the goodness-of-fit, we will store the estimates 
of all models fitted here for post analysis:

```{r storing-models}
# Model fits
saveRDS(
  object = setNames(list(
    res_big_ergm0,
    res_big_ergm1,
    res_big_ergm2,
    res_big_ergm3,
    res_big_ergm4,
    res_big_ergm5,
    names(networks)
    ), c(paste("Model", 0:5), "netnames")),
  file = "2022-04-25-bipartite-ergms.rds"
)

# Functions
saveRDS(
  object = list(
    namemilize = namemilize,
    calc_ergm_pearson = calc_ergm_pearson
  ),
  file = "2022-04-25-bipartite-ergm-functions.rds"
)
```

# Next steps (instructions)

The following hypotheses have not been tested yet:

- Varying density per state (fixed-effects.) This could be done using the term: `b1factor("state", level = NULL)`.

- Two and three-star mixing effect on "short_stay."

- Higher-order dyadwise-shared-partners (b1dsp and b2dsp.)

- Geometrically-weighted dyadwise-shared-partners. This can be done with the
terms `gwb1dsp()` and `gwb2dsp()`. This is a more stable term than `b1dsp`.

- Geometrically-weighted degree distribution for the second mode. This can be
done using the `gwb1degree` term. In particular, we are interested on including
this term by state, i.e., `gwb1degree(decay, attr="state", fixed=TRUE)`, with
the following values for decay: $\{0.1, 0.2, .5\}$. This effect will be similar
to density per state. If it doesn't work with density per state, exclude density
per state when using this effect.

The idea is to include those terms in models 3 and 5, while at the same time
excluding b1dsp1 and b1dsp2 from model 5. Add one term at a time and save the results
as I did with the previous models in an RDS file for posterior analysis.


# Goodness of fit

For computing the goodness of fit, we need to take each model and
calculate the expected value of the observed sufficient statistics.
Model 5 (ergm4) seems to be one of the best fitting models, so let's
start with that one. We will compute the standardized pearson residual
as in Krivitstky, Coletti, and Hans (2022):

```{r computing-resids, cache=TRUE}
pearson_residuals <- parallel::mclapply(
  networks, calc_ergm_pearson, model. = res_big_ergm2,
  nsim. = 1e3, mc.cores = 8L
  )
pearson_residuals <- Map(
  \(n,dat) cbind(dat, net = n),
  n = names(pearson_residuals), dat = pearson_residuals)
pearson_residuals <- rbindlist(pearson_residuals)
pearson_residuals <- pearson_residuals[s_obs != 0]
pearson_residuals[, t_stat := (s_obs - s_sim_avg)/s_sim_std]
```

```{r viz-residuals}
library(ggplot2)
ggplot(pearson_residuals, aes(y = t_stat, x = s_name)) +
  geom_violin() +
  labs(x = "Statistic", y = "Pearson Resid.") +
  coord_flip() 
```

```{r viz-residuals-edges}
library(ggplot2)
ggplot(pearson_residuals[s_name == "edges"]) +
  geom_point(
    aes(
      y = t_stat,
      x = sqrt(s_sim_avg)
    )
  ) +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x = expression(sqrt(Fitted~Values~(edges))), y = "Pearson Resid.")
```

```{r viz-residuals-b2star2}
ggplot(pearson_residuals[s_name == "b2star2"]) +
  geom_point(
    aes(
      y = t_stat,
      x = sqrt(s_sim_avg)
    )
  ) +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x = expression(sqrt(Fitted~Values~(b2star2))), y = "Pearson Resid.")
```
