---
title: "Bipartite networks of health care workers and residents in long-term care facilities"
output:
  pdf_document:
    extra_dependencies:
      - pdflscape
      - tabularx
      - booktabs
      - geometry
      - graphicx
editor_options: 
  markdown: 
    wrap: 72
author: George G. Vega Yon
date: May 3, 2022
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Data Preparation

For this analysis we will use the following R packages:

- network: For manipulation of network data.

- data.table: For data cleaning and processing.

- ergm: For estimation and simulation of Exponential Random Graph Models.

- netplot: For visualization of networks.

- ggplot2: For general visualization.

- gridExtra: Utilities for grid graphics (base of ggplot2 and netplot.)

```{r packages, message=FALSE, warning=FALSE}
library(network)
library(data.table)
library(ergm)
library(ergm.multi)
library(netplot)
library(ggplot2)
library(gridExtra)
```

# Loading data

```{r}
networks <- readRDS("data/model_data_networks.rds")
```


# Fitting ERGMs

For the baseline models, we will be using the terms included in table 1. Before we continue, we will now concatenate all networks into a single large graph to be used in with the ERGM package. Although what we want is to fit a pooled-bipartite ERGM, as of today the `ergm` package has no way of dealing with multiple networks. Nonetheless, mathematically, block-diagonal ERGMs--in which ties are only permitted within groups--are equivalent to pooled ERGMs.

\def\tmpfigwidth{\includegraphics[width=.5\linewidth]}
\begin{table}[h]
	\begin{tabular}{m{.4\linewidth}m{.2\linewidth}m{.3\linewidth}<\centering}
		\toprule
		Name & Representation & Form \\ \midrule
		Edge count & \tmpfigwidth{fig/term-edgecount.png} &$\sum_{i<j}y_{ij}$ \\\midrule
		HCW-centered two-star & \tmpfigwidth{fig/term-actor-twostar.png} & $\sum_{i\in H, \{j<k\}\in R}y_{ij}y_{ik}$ \\\midrule
		Resident-centered two-star &\tmpfigwidth{fig/term-entity-twostar.png} & $\sum_{i\in R, \{j<k\}\in H}y_{ji}y_{ki}$ \\\midrule
		HCW-centered homophilic two-star & \tmpfigwidth{fig/term-actor-twostar-mix.png} & $\sum_{i\in H, \{j<k\}\in R}\mathbf{1}\left(x_{j} = x_k\right) y_{ij}y_{ik}$ \\\midrule
		Resident-centered homophilic two-star & \tmpfigwidth{fig/term-entity-twostar-mix.png} & $\sum_{i\in R, \{j<k\}\in H}\mathbf{1}\left(x_{j} = x_k\right)y_{ji}y_{ki}$ \\\midrule
		Dyad-wise one shared residents &  \tmpfigwidth{fig/term-actor-one-edgewise-shared-partner.png} & $\sum_{\{i<j\}\in H, k\in R}y_{ik}y_{jk}$ \\ \midrule
		Dyad-wise two shared residents & \tmpfigwidth{fig/term-actor-two-edgewise-shared-partner.png} & $\sum_{\{i<j\}\in H, \{k<h\}\in R}y_{ik}y_{jk}y_{ih}y_{jh}$ \\ 
		\bottomrule
	\end{tabular}
	\label{tab:terms}
	\caption{Terms used in the bipartite ERGM models}
\end{table}
\def\tmpfigwidth{}

## Structural terms ERGM

```{r multi-to-single}
bignet <- Map(\(n,e) {
  
  # Creating the baseline data
  el <- list(
    edges = as.data.frame(n, unit = "edges"),
    # Second column matches the first, so we remove it
    nodes = as.data.frame(n, unit = "vertices") #[,-1L] 
  )
  
  # Appending the id of the dataset to avoid duplicates
  el$edges$.head <- paste0(e, ".", el$edges$.head)
  el$edges$.tail <- paste0(e, ".", el$edges$.tail)
  el$nodes$vertex.names <- paste0(e, ".", el$nodes$vertex.names)
  
  el$nodes$nvertex <- nrow(el$nodes)
  
  # Returning
  el
}, n = networks, e = names(networks))

bignet <- list(
  edges = data.table(do.call(rbind, lapply(bignet, "[[", "edges"))),
  nodes = data.table(do.call(rbind, lapply(bignet, "[[", "nodes")))
)

bignet_mat <- as.network(
  x         = bignet$edges,
  vertices  = bignet$nodes,
  bipartite = TRUE,
  bipartite_col = "is_actor"
)

# Comparing we are getting the same thing
get.inducedSubgraph(bignet_mat, which((bignet_mat %v% "net_id") == names(networks)[1]))
networks[[1]]
```

Common controls for all models:

```{r}
default_control <- list(
  seed            = 44485,
  main.method     = "Stochastic-Approximation",
  MCMC.samplesize	= 1024 * 4 #,
  # MCMC.interval   = 1024 * 4,
  # parallel        = 2L
)

# Also, all models will feature the same set of networkws:
complete_nets <- (!sapply(networks, \(n) {
  any(is.na(n %v% "wound_care")) |
    any(is.na(n %v% "mdro")) |
    any(is.na(n %v% "dialysis")) |
    any(is.na(n %v% "diabetes")) |
    any(is.na(n %v% "wheelchair"))
})) |> which()

networks <- networks[complete_nets]
```

We can could use Krivitsky (2011) offset, which in the case of multiple
graphs would be $\sum_{i} \text{edges}_i\times\ln\left({1/n_i}\right)$,
where $n_i$ is the number of nodes in network $i$. Nonetheless, in Butts
(2013) it is suggested to use a free parameter
$\sum_i\ln\left(n_i\right)$. To do this we use the new functions
included in `ergm` 4.0, in particular, the `Sum()` term. Two steps to do
this:

1.  Compute the values $\ln\left({n_i}\right)$ separately, and

2.  Use that vector to pre-multiply the vector of edgecounts at the
    graph level. We can get an edgecount at the graph level using the
    `b1factor()` term with the option `levels = TRUE`. The following
    codeblock achieves this:

```{r}
# log(n) part of the term.
log_n <- log(sapply(networks, \(n) summary(n ~ degrange(0))))

# Let's see how it looks
summary(Networks(networks) ~ edges + N(~ Sum(log_n ~ edges, "a")))
```

Let's see how it is implemented in an ERGM. Remember that we need to set
the term as an offset:

```{r model0, cache = TRUE, message=FALSE, warning=FALSE}
res_big_ergm0 <- ergm(
  Networks(networks) ~
    N(~edges, ~log(n)),
  control = do.call("control.ergm", c(default_control))
  )
```

```{r model0-tabel, echo = FALSE, results='asis'}
texreg::knitreg(list(`Model 0` = res_big_ergm0))
```

Interpretation of the `edges` term needs to be done in the context of
this offset term. For example, since $\theta_{edges} =$
`r sprintf("%.2f", coef(res_big_ergm0)[1])`, the density in a Bernoulli
graph with 100 nodes is
$\text{logit}^{-1}\left(\theta_{edges} + \theta_{logn}\ln\left(100\right)\right)\approx$
`r sprintf("%.2f", plogis(coef(res_big_ergm0[1]) %*% c(1, log(100))))`.
As in Krivitsky et al. (2022), we can scale terms across the model in
such a fashion. Let's repeat the exercise with the b2 2-star (second
level 2-stars) term:

```{r model1, cache = TRUE, eval = TRUE, warning=FALSE, message=FALSE, results='hide'}
res_big_ergm1 <- ergm(
  Networks(networks) ~
    N(~edges, ~log(n)) +
    N(~b2star(2), ~log(n)),
  control = do.call("control.ergm", c(default_control))
  )
```

```{r model1-tabel, echo = FALSE, results='asis'}
texreg::knitreg(list(`Model 1` = res_big_ergm1))
```

In this case, the interaction with $log(n)$ did not improved our model,
so we will keep only the three other significant terms. Following
Krivitsky et al (2022,) we can leverage having multiple networks to
asses which other terms would be beneficial to be included in our model. The following
function will help us compute the predicted values per network (including the
standard error and observed values, which we will later use for goodness-of-fit
analyses.)

To do this, we will perform the following steps:

1. Compute the observed ERGM pearson standardized residuals,

2. Compute the observed motif (term) count that were not included in the model

3. Run a linear regression predicting the residuals as a function of the terms
not included. Terms that were not included in the model that turn out to be
siginificant will be then included in the ERGM:

```{r model0-regressions-step1, eval = FALSE}
gof_res <- gofN(res_big_ergm1)
plot(gof_res, against = ~sqrt(.fitted))
```

Computing now the observed statistics of the following terms:

- 3 and 4 star in both levels: `b1star(3:4)` and `b2star(3:4)`
- Residents race-homophily in the second level: `b1starmix(2, "white", diff = FALSE)`
- HWC type homophily: `b2starmix(2, "type")`
- One and two shared residents per worker: `b1dsp(1:2)`
- Covariate effect of residents age: `b2cov("sqrt_age")`


## More covariate-based effects

Second model, Using homophily in race:

```{r model2, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE, results='hide'}
res_big_ergm2 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2), ~ log(n)) + 
    N(~b1starmix(2, "white", diff = FALSE)),
  control = do.call("control.ergm", default_control)
  )
```

```{r model3, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE, results='hide'}
res_big_ergm3 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2), ~ log(n)) + 
    b1starmix(2, "white", diff = FALSE) +
    b2starmix(2, "type"),
  control = do.call("control.ergm", default_control)
  )
```

```{r model4, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE, results='hide'}
res_big_ergm4 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2), ~ log(n)) + 
    b1starmix(2, "white", diff = FALSE) +
    b2starmix(2, "type") +
    b1dsp(1:2) +
    b2cov("sqrt_age"),
  control = do.call("control.ergm", default_control)
  )
```

```{r model5, cache = TRUE, message=FALSE, warning=FALSE, eval = TRUE, results='hide'}
res_big_ergm5 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2), ~ log(n)) + 
    b1starmix(2, "white", diff = FALSE) +
    b2starmix(2, "type") +
    b1dsp(1:2) +
    b2cov("sqrt_age") +
    # b1starmix(2, "wound_care", diff = FALSE) + # 3 missing
    b1starmix(2, "diabetes", diff = FALSE) +
    b1starmix(2, "dialysis", diff = FALSE) +
    b1starmix(2, "wheelchair", diff = FALSE) +
    b1starmix(2, "bedridden", diff = FALSE) +
    b1starmix(2, "ventilator", diff = FALSE),
    # b1starmix(2, "mdro", diff = FALSE),  # 5 missing
  control = do.call("control.ergm", default_control)
  )
```

```{r models-print-big, results='asis', echo=FALSE, message=FALSE, warning=FALSE, eval = TRUE}
cat("\\begin{landscape}\n\n")
texreg::knitreg(
  setNames(list(
    res_big_ergm0,
    res_big_ergm1,
    res_big_ergm2,
    res_big_ergm3,
    res_big_ergm4
    ), paste("Model", 0:4)),
  # Excluding infinte coefficients. Only the mix in resident something matters.
  omit.coef = '2\\.type\\.([^r]|.+PA/NP|.+Physician)',
  single.row = TRUE
)
cat("\\end{landscape}\n\n")
```

```{r echo=FALSE, results='asis'}
texreg::knitreg(
  list("Model 5" = res_big_ergm5),
  omit.coef = '2\\.type\\.([^r]|.+PA/NP|.+Physician)',
  single.row = TRUE
  )
```

```{r ventmodel, cache=TRUE}
res_big_ergm6 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2), ~ log(n)) + 
    b1starmix(2, "white", diff = FALSE) +
    b2starmix(2, "type") +
    b1dsp(1:2) +
    b2cov("sqrt_age") +
    b1starmix(2, "wound_care", diff = FALSE) + 
    b1starmix(2, "diabetes", diff = FALSE) +
    b1starmix(2, "dialysis", diff = FALSE) +
    b1starmix(2, "wheelchair", diff = FALSE) +
    b1starmix(2, "bedridden", diff = FALSE) +
    b1starmix(2, "ventilator", diff = FALSE) +
    b1starmix(2, "mdro", diff = FALSE),  
  control = do.call("control.ergm", default_control)
  )

res_big_ergm7 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2)) + 
    N(~b1starmix(2, "white", diff = FALSE), ~log(n)) +
    b2starmix(2, "type") +
    N(~b1dsp(1:2)) +
    # b2cov("sqrt_age") +
    N(~b1starmix(2, "wound_care", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "diabetes", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "dialysis", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "wheelchair", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "bedridden", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "ventilator", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "mdro", diff = FALSE), ~log(n)),  
  control = do.call("control.ergm", default_control)
  )

res_big_ergm8 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2)) + 
    N(~b1starmix(2, "white", diff = FALSE), ~log(n)) +
    b2starmix(2, "type") +
    N(~b1dsp(1:2)) +
    # b2cov("sqrt_age") +
    # N(~b1starmix(2, "wound_care", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "diabetes", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "dialysis", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "wheelchair", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "bedridden", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "ventilator", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "mdro", diff = FALSE), ~log(n)),  
  control = do.call("control.ergm", default_control)
  )


res_big_ergm9 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2)) + 
    N(~b1starmix(2, "white", diff = FALSE), ~log(n)) +
    b2starmix(2, "type") +
    N(~b1dsp(1:2)) +
    # b2cov("sqrt_age") +
    # N(~b1starmix(2, "wound_care", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "diabetes", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "dialysis", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "wheelchair", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "bedridden", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "ventilator", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "mdro", diff = FALSE), ~log(n)),  
  control = do.call("control.ergm", default_control)
  )

res_big_ergm10 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2)) + 
    N(~b1starmix(2, "white", diff = FALSE), ~log(n)) +
    b2starmix(2, "type") +
    N(~b1dsp(1:2)) +
    # b2cov("sqrt_age") +
    # N(~b1starmix(2, "wound_care", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "diabetes", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "dialysis", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "wheelchair", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "bedridden", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "ventilator", diff = FALSE), ~log(n)), #+
    # N(~b1starmix(2, "mdro", diff = FALSE), ~log(n))  
  control = do.call("control.ergm", default_control)
  )

res_big_ergm11 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2), ~ log(n)) + 
    N(~b1starmix(2, "white", diff = FALSE), ~log(n)) +
    b2starmix(2, "type") +
    N(~b1dsp(1:2)) +
    # b2cov("sqrt_age") +
    # N(~b1starmix(2, "wound_care", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "diabetes", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "dialysis", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "wheelchair", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "bedridden", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "ventilator", diff = FALSE), ~log(n)), #+
    # N(~b1starmix(2, "mdro", diff = FALSE), ~log(n))  
  control = do.call("control.ergm", default_control)
  )

res_big_ergm12 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2), ~ log(n)) + 
    N(~b1starmix(2, "white", diff = FALSE), ~log(n)) +
    b2starmix(2, "type") +
    N(~b1dsp(1:2)) +
    # b2cov("sqrt_age") +
    # N(~b1starmix(2, "wound_care", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "diabetes", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "dialysis", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "wheelchair", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "bedridden", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "ventilator", diff = FALSE), ~log(n)), #+
    # N(~b1starmix(2, "mdro", diff = FALSE), ~log(n))  
  control = do.call("control.ergm", default_control)
  )

res_big_ergm13 <- ergm(
  Networks(networks) ~ 
    N(~edges, ~ log(n)) + 
    N(~b2star(2)) + 
    N(~b1starmix(2, "white", diff = FALSE), ~log(n)) +
    b2starmix(2, "type") +
    N(~b1dsp(1:2)) +
    # b2cov("sqrt_age") +
    # N(~b1starmix(2, "wound_care", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "diabetes", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "dialysis", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "wheelchair", diff = FALSE), ~log(n)) +
    # N(~b1starmix(2, "bedridden", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "ventilator", diff = FALSE), ~log(n)) +
    N(~b1starmix(2, "mdro", diff = FALSE), ~log(n)),
  control = do.call("control.ergm", default_control)
  )
```

```{r results='asis'}
texreg::knitreg(
  setNames(list(
    res_big_ergm0,
    res_big_ergm1,
    res_big_ergm2,
    res_big_ergm3,
    res_big_ergm4,
    res_big_ergm5
    ), paste("Model", c(0:5))),
  # Excluding infinte coefficients. Only the mix in resident something matters.
  omit.coef = '2\\.type\\.([^r]|.+PA/NP|.+Physician)',
  single.row = TRUE
)

texreg::knitreg(
  setNames(list(
    res_big_ergm8,
    res_big_ergm9,
    res_big_ergm10,
    res_big_ergm11,
    res_big_ergm12,
    res_big_ergm13
    ), paste("Model", c(8:13))),
  # Excluding infinte coefficients. Only the mix in resident something matters.
  omit.coef = '2\\.type\\.([^r]|.+PA/NP|.+Physician)',
  single.row = TRUE
)

save(file = "2022-04-25-bipartite-ergms.rdata", list = ls(pattern = "^res_big_ergm.+"))
```


So far, model 10 is the best overall fit. Let's refit the model with the default MCMLE parameters

```{r refit-best}
default_control_best <- list(
  seed          = 485, # We change the default
  main.method   = "Stochastic-Approximation",
  SA.samplesize	= 1024 * 16,
  SA.interval   = 1024 * 32, # 32 times the default (it has very high autocorr)
  SA.nsubphases = 4 * 1, # 1 the default
  SA.burnin     = 1024 * 32, # Twice the default
  MCMLE.interval = 1024 * 32, # 32 times
  RM.interval = 1024 * 32 # 32 times... This is being passed to SA instead
)

res_best2 <- ergm(
  res_big_ergm10$formula,
  control = do.call(control.ergm, default_control_best)
  )

# Did things changed?
texreg::knitreg(
  setNames(list(
    res_big_ergm10,
    res_best
    ), c("Best model", "Refit with sampler size")),
  # Excluding infinte coefficients. Only the mix in resident something matters.
  omit.coef = '2\\.type\\.([^r]|.+PA/NP|.+Physician)',
  single.row = TRUE
)

# Generating diagnostics
diag_ergm2 <- mcmc.diagnostics(res_best2, center = TRUE)
```


# Next steps (instructions)

The following hypotheses have not been tested yet:

- Varying density per state (fixed-effects.) This could be done using the term: `b1factor("state", level = NULL)`.

- Two and three-star mixing effect on "short_stay."

- Higher-order dyadwise-shared-partners (b1dsp and b2dsp.)

- Geometrically-weighted dyadwise-shared-partners. This can be done with the
terms `gwb1dsp()` and `gwb2dsp()`. This is a more stable term than `b1dsp`.

- Geometrically-weighted degree distribution for the second mode. This can be
done using the `gwb1degree` term. In particular, we are interested on including
this term by state, i.e., `gwb1degree(decay, attr="state", fixed=TRUE)`, with
the following values for decay: $\{0.1, 0.2, .5\}$. This effect will be similar
to density per state. If it doesn't work with density per state, exclude density
per state when using this effect.

The idea is to include those terms in models 3 and 5, while at the same time
excluding b1dsp1 and b1dsp2 from model 5. Add one term at a time and save the results
as I did with the previous models in an RDS file for posterior analysis.


# Goodness of fit

For computing the goodness of fit, we need to take each model and
calculate the expected value of the observed sufficient statistics.
Model 5 (ergm4) seems to be one of the best fitting models, so let's
start with that one. We will compute the standardized pearson residual
as in Krivitstky, Coletti, and Hans (2022):

```{r gof, cache=TRUE}
gof_res <- gofN(res_best2)
plot(gof_res, against = ~factor(n))
```

