---
format: html
title: Output reporting
author: George G. Vega Yon, Ph.D.
date: 2025-04-01
date-modified: 2025-05-15
embed-resources: true
---

This document shows the final model's goodness-of-fit [GOF] output in terms of multi-ergm modeling, as well as the MCMC GOF diagnostics.

```{r}
#| warning: false
#| message: false
library(ergm.multi)
library(texreg)
library(ergmito)
library(ggplot2)
library(ggrepel)
library(data.table)
library(xtable)
```

## Summary stats

```{r}
#| label: load-packages-and-data
# Loading data from the models
load("../data/res12_93.rda")
load('../data/network93_f2.RData')
load('../data/dd_nodal_attr.RData')
setDT(dd)
dd$NetID=as.numeric(dd$NetID)

# Summary about MDROs
mdro <- dd[, n := sum(mdro == 'Yes'), NetID][n>0, unique(NetID)] 

counts <- summary_formula(
  res12_93$network ~
    N(~edges, ~log(n))+
    N(~gwb1dsp(0.75, fixed=TRUE))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'wound_care', diff=TRUE), subset = wound_care) +
    N(~b1starmix(2, 'ventilator', diff=TRUE), subset = ventilator2) +
    N(~b1starmix(2, 'diabetes', diff=TRUE), subset=diabetes2) +
    N(~b2starmix(2, 'PTOTRT'), subset=PTOTRT) +
    N(~b2starmix(2, 'mdro', diff = TRUE), subset=mdro)
  )

n <- 93 #uncombine_network(res12_93$network) |>
  #length()

mytab <- data.table(
  Statistic = names(counts),
  Value     = as.integer(counts),
  Avg       = counts/n
)

mytab[, Statistic := gsub("b.starmix.{3}", "", Statistic)]
knitr::kable(
  mytab,
  caption = "Summary statistics of the model",
  digits = 2
)
```

```{r}
#| label: latex-table
#| echo: false
#| include: false
# Creating the latex table
mytab <- data.table(
  Statistic = paste("\\hspace{.5cm}", names(counts)),
  Value     = as.integer(counts),
  Avg       = counts/n
)

mytab <- xtable::xtable(mytab)

ans <- list(
  mytab[1:5,],
  mytab[6:11,],
  mytab[11:nrow(mytab),]
  )

lapply(ans, print, 
  include.colnames = FALSE,
  include.rownames = FALSE,
  only.contents = TRUE,
  hline.after = NULL,
  sanitize.text.function	= \(x) x
  ) |> invisible()

```

## Final model

The code for the final model is displayed below. Right now, the data is not publicly available, so the code is deactivated. Notice that we run the model twice: the first run uses the Stochastic Approximation algorithm and the second uses MC-MLE:

```r
res12_92 <- ergm(
Networks(networks1) ~
  N(~edges, ~log(n))+
  N(~gwb1dsp(0.75, fixed=TRUE))+
  N(~b2starmix(2, 'Nurse'), subset=Nurse)+
  N(~b2factor('wound_care'), subset=wound_care)+
  N(~b2factor('bedridden'), subset=bedridden)+
  N(~b1starmix(2, 'wound_care', diff=TRUE), subset=wound_care)+
  N(~b1starmix(2, 'ventilator', diff=TRUE), subset=ventilator2)+
  N(~b1starmix(2, 'diabetes', diff=TRUE), subset=diabetes2)+
  N(~b2starmix(2, 'PTOTRT'), subset=PTOTRT)+
  offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf)
)


res12_93 <- ergm(
  Networks(networks) ~
    N(~edges, ~log(n))+
    N(~gwb1dsp(0.75, fixed=TRUE))+
    N(~b2starmix(2, 'Nurse'), subset=Nurse)+
    N(~b2factor('wound_care'), subset=wound_care)+
    N(~b2factor('bedridden'), subset=bedridden)+
    N(~b1starmix(2, 'wound_care', diff=TRUE), subset=wound_care)+
    N(~b1starmix(2, 'ventilator', diff=TRUE), subset=ventilator2)+
    N(~b1starmix(2, 'diabetes', diff=TRUE), subset=diabetes2)+
    N(~b2starmix(2, 'PTOTRT'), subset=PTOTRT)+ 
    offset(b1degree(0))+offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = do.call(
    "control.ergm",
    c(list(init = coef(res12_92)))
    )
)
```

```{r}
#| label: final-model
texreg::screenreg(
  res12_93,
  digits = 2
)
```

## Loading the GOF

The GOF statistics were computed during the model fitting process. We just need to load them.

```{r}
#| label: load-and-plot-gofN
gof12_93 <- readRDS("../data/gofN_model12_93.rds")

if (interactive())
  pdf("gof12_93.pdf")

op <- par(
  mfrow = c(4, 2),
  # Putting the x-labels inside of the plots
  oma = c(0, 0, 0, 0),
  mar = c(4, 4, 2, 1),
  # Reducing the size of the title
  cex.main = 0.8
  )
# figs <- autoplot(gof12_93, main = "{name} - {type}", id.n = 5)
plot(
  gof12_93[1:15], main = "{name}\n{type}", id.n = 5, , which = 1,
  xlab = ""
  )
par(op)

if (interactive())
  dev.off()

```


Could have we done better? We look into this from comparing the observed vs predicted values (the previous focused on the error terms):

```{r}
#| label: gof-obs-vs-fitted
#| warning: false
vars <- names(gof12_93)

op <- par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
for (v in vars) {
  tmp_var <- gof12_93[[v]]

  tryCatch(with(tmp_var, {
    plot(observed, fitted, main = v, xlab = "Observed", ylab = "Fitted")
    abline(a = 0, b=1, col = "red")
  }), error = function(e) {
    if (interactive())
      message("Error in variable:", v, "\n")
  })
}
par(op)

```


## Loading the fit


```{r}

```

## Visualizing some features of the networks

```{r}
#| label: network-feats
# Mean degree
network_stats <- data.frame(
  size = sapply(networks, network.size),
  n_hcw = sapply(networks, \(x) {
    sum(x %v% "is_actor")
  }),
  n_mdro = sapply(networks, \(x) {
    sum((x %v% "mdro") == "Yes")
  }),
  n_dialysis = sapply(networks, \(x) {
    sum((x %v% "dialysis") == "Yes")
  }),
  n_wound_care = sapply(networks, \(x) {
    sum((x %v% "wound_care") == "Yes")
  }),
  n_ventilator = sapply(networks, \(x) {
    sum((x %v% "ventilator") == "Yes")
  }),
  n_bedridden = sapply(networks, \(x) {
    sum((x %v% "bedridden") == "Yes")
  }),
  # Now, tabulating the network composition
  n_nurses = sapply(networks, \(x) {
    sum((x %v% "Nurse") == 1)
  }),
  n_PTOTRT = sapply(networks, \(x) {
    sum((x %v% "PTOTRT") == 1)
  }),
  hcw_mean_degree = sapply(networks, \(x) {
    mean(sna::degree(x)[x %v% "is_actor"])
  })
)

# Size distribution
fig_size <- ggplot(network_stats, aes(x = size)) +
  geom_histogram() +
  labs(
    x = "Network Size",
    y = "Count",
    title = "Size Distribution"
  ) +
  theme_minimal()

# Size vs number of HCWs
fig_size_hcw <- ggplot(network_stats, aes(x = size, y = n_hcw)) +
  geom_point() +
  labs(
    x = "Network Size",
    y = "Number of HCWs",
    title = "Number of HCWs"
  ) +
  theme_minimal()

# Size vs HCW's mean degree
fig_size_hcw_degree <- ggplot(network_stats, aes(x = size, y = hcw_mean_degree)) +
  geom_point() +
  labs(
    x = "Network Size",
    y = "Mean Degree of HCWs",
    title = "Mean Degree of HCWs"
  ) +
  theme_minimal()

# Combine both figures into a single ggplot object
library(patchwork)
fig_size + fig_size_hcw + fig_size_hcw_degree
ggsave("size-distribution.png", width = 9*.75, height = 3*.75)

library(gridExtra)
```

## MCMC diagnostics
```{r}
#| label: mcmc-diagnostics
#| fig-height: 10 
op <- par(mfrow = c(3, 2))
mcmc.diagnostics(res12_93, which = "plots")
par(op)
```

Another view to the errors

```{r}
#| label: regress-networks
# Retrieving the errors
errs <- lapply(gof12_93, "[[", "pearson")

# Most commonly misfitted networks looking at the MSE
errs <- data.table(
  network = as.character(1:length(errs[[1]])),
  mse = do.call(cbind, errs) |>
    apply(1, \(x) mean(abs(x), na.rm = TRUE))
)

setorder(errs, mse)
errs[, network := factor(network, levels = network)]

# Plotting
pdf("errs.pdf")
ggplot(tail(errs, 20), aes(x = network, y = mse)) +
  geom_point() +
  geom_line() +
  labs(
    x = "Network",
    y = "Mean Squared Error",
    title = "Mean Squared Error of the Networks",
    subtitle = "(top 20)"
  ) +
  theme_minimal()
dev.off()
```

```{r}
#| label: simulate-networks
#| cache: true
sims <- simulate(res12_93, nsim = 1, seed = 1234) |>
  splitnetwork(".NetworkID")

# Original networks
networks_0 <- res12_93$network |>
  splitnetwork(".NetworkID")
```

```{r}
#| label: plot-simulated-networks
library(netplot)
if (interactive())
  svg("example1.svg", width = 8, height = 4)
ans_row0 <- NULL
ans_row1 <- NULL
ans <- NULL

mynplot <- function(net, ...) {
  nplot(
    net, edge.line.breaks = 1, vertex.label = "", vertex.size = 1,
    vertex.nsides = ifelse(
      net %v% "is_actor", 3, 4
    ),
    vertex.color = ifelse(
      net %v% "is_actor", "black", "white"
    ),
    edge.color = "darkgrey",
    vertex.size.range = c(.05, .05, 1),
    ...
  )
}

for (i in c(6, 8, 9)) {
  set.seed(331)
  res0 <- mynplot(sims[[i]])
  res1 <- mynplot(networks_0[[i]])
  ans <- c(ans, list(res0, res1))
}

# Groups them by column (column major)
marrangeGrob(
  grobs = ans, ncol = 3, nrow = 2, padding = unit(.5, "line"),
  top = ""
  )

if (interactive())
  dev.off()
```
