---
format: html
title: Output reporting
author: George G. Vega Yon, Ph.D.
date: 2025-04-01
date-modified: 2025-05-15
---

This document shows the final model's goodness-of-fit [GOF] output in terms of multi-ergm modeling, as well as the MCMC GOF diagnostics.


```{r}
#| warning: false
#| message: false
library(ergm.multi)
library(texreg)
library(ergmito)
library(ggplot2)
library(ggrepel)
```

## Loading the GOF

The GOF statistics were computed during the model fitting process. We just need to load them.

```{r}
#| label: load-and-plot-gofN
gof12_93 <- readRDS("../data/gofN_model12_93.rds")

if (interactive())
  pdf("gof12_93.pdf")

op <- par(
  mfrow = c(4, 2),
  # Putting the x-labels inside of the plots
  oma = c(0, 0, 0, 0),
  mar = c(4, 4, 2, 1),
  # Reducing the size of the title
  cex.main = 0.8
  )
# figs <- autoplot(gof12_93, main = "{name} - {type}", id.n = 5)
plot(
  gof12_93[1:15], main = "{name}\n{type}", id.n = 5, , which = 1,
  xlab = ""
  )
par(op)

if (interactive())
  dev.off()

```


Could have we done better? We look into this from comparing the observed vs predicted values (the previous focused on the error terms):

```{r}
#| label: gof-obs-vs-fitted
#| warning: false
vars <- names(gof12_93)

op <- par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
for (v in vars) {
  tmp_var <- gof12_93[[v]]

  tryCatch(with(tmp_var, {
    plot(observed, fitted, main = v, xlab = "Observed", ylab = "Fitted")
    abline(a = 0, b=1, col = "red")
  }), error = function(e) {
    if (interactive())
      message("Error in variable:", v, "\n")
  })
}
par(op)

```


## Loading the fit


```{r}
#| label: load-packages-and-data
load("../data/res12_93.rda")
load('../data/network93_f2.RData')
load('../data/dd_nodal_attr.RData')
```

## Visualizing some features of the networks

```{r}
#| label: network-feats
# Mean degree
network_stats <- data.frame(
  size = sapply(networks, network.size),
  n_hcw = sapply(networks, \(x) {
    sum(x %v% "is_actor")
  }),
  n_mdro = sapply(networks, \(x) {
    sum((x %v% "mdro") == "Yes")
  }),
  n_dialysis = sapply(networks, \(x) {
    sum((x %v% "dialysis") == "Yes")
  }),
  n_wound_care = sapply(networks, \(x) {
    sum((x %v% "wound_care") == "Yes")
  }),
  n_ventilator = sapply(networks, \(x) {
    sum((x %v% "ventilator") == "Yes")
  }),
  n_bedridden = sapply(networks, \(x) {
    sum((x %v% "bedridden") == "Yes")
  }),
  # Now, tabulating the network composition
  n_nurses = sapply(networks, \(x) {
    sum((x %v% "Nurse") == 1)
  }),
  n_PTOTRT = sapply(networks, \(x) {
    sum((x %v% "PTOTRT") == 1)
  })
)

# Size distribution
ggplot(network_stats, aes(x = size)) +
  geom_histogram() +
  labs(
    x = "Network Size",
    y = "Count",
    title = "Network Size Distribution"
  ) +
  theme_minimal()

# Size vs number of HCWs
ggplot(network_stats, aes(x = size, y = n_hcw)) +
  geom_point() +
  # geom_smooth(method = "lm") +
  labs(
    x = "Network Size",
    y = "Number of HCWs",
    title = "Network Size vs Number of HCWs"
  ) +
  theme_minimal()

```

## MCMC diagnostics
```{r}
#| label: mcmc-diagnostics
#| fig-height: 10 
op <- par(mfrow = c(3, 2))
mcmc.diagnostics(res12_93, which = "plots")
par(op)
```

Another view to the errors

```{r}
#| label: regress-networks
# Retrieving the errors
errs <- lapply(gof12_93, "[[", "pearson")

# Most commonly misfitted networks looking at the MSE
library(data.table)
errs <- data.table(
  network = as.character(1:length(errs[[1]])),
  mse = do.call(cbind, errs) |>
    apply(1, \(x) mean(abs(x), na.rm = TRUE))
)

setorder(errs, mse)
errs[, network := factor(network, levels = network)]

# Plotting
pdf("errs.pdf")
ggplot(tail(errs, 20), aes(x = network, y = mse)) +
  geom_point() +
  geom_line() +
  labs(
    x = "Network",
    y = "Mean Squared Error",
    title = "Mean Squared Error of the Networks",
    subtitle = "(top 20)"
  ) +
  theme_minimal()
dev.off()
```

```{r}
#| label: simulate-networks
#| cache: true
sims <- simulate(res12_93, nsim = 1, seed = 1234) |>
  splitnetwork(".NetworkID")

# Original networks
networks_0 <- res12_93$network |>
  splitnetwork(".NetworkID")
```

```{r}
#| label: plot-simulated-networks
library(netplot)
library(gridExtra)
if (interactive())
  png("example1.png", width = 1024, height = 400)
ans_row0 <- NULL
ans_row1 <- NULL
ans <- NULL
for (i in 1:5) {
  res0 <- nplot(sims[[i]], edge.line.breaks = 1)
  res1 <- nplot(networks_0[[i]], edge.line.breaks = 1, layout = res0$.layout)
  # ans_row0 <- c(ans_row0, list(res0))
  # ans_row1 <- c(ans_row1, list(res1))
  ans <- c(ans, list(res0, res1))
}
  # ans <- c(ans_row0, ans_row1)
marrangeGrob(grobs = ans, ncol = 5, nrow = 2)

if (interactive())
  dev.off()
```


## Re-fitting the model excluding the worst networks

```{r}
#| label: refit-model
#| warning: false
#| message: false
#| eval: false
# To remove
to_remove <- c(4, 54, 57)

# Cleaning variables
dd$NetID=as.numeric(dd$NetID)
setDT(dd)

for (i in rev(to_remove)) {
  dd <- dd[NetID != i] # also remove it from the nodal attributes data frame
  dd$NetID[dd$NetID > i] <- dd$NetID[dd$NetID > i] - 1 # re-number the networks after removing the outlier
}

# Get indices of networks on which each term will be evaluated.
# For HCW type homophily (b2starmix(2)), the netowkrs need to have at least two HCWs of that type. 
Nurse=dd[, n:=sum(Nurse==1), NetID][n>1, unique(NetID)] |> as.numeric() 
CNA=dd[, n:=sum(CNA==1), NetID][n>1, unique(NetID)] |> as.numeric()
PTOTRT=dd[, n:=sum(PTOTRT==1), NetID][n>1, unique(NetID)] |> as.numeric()
Other=dd[, n:=sum(Other==1), NetID][n>1, unique(NetID)] 

# For b2factor (resident type), the networks need to have at least 1 resident of that type. 
dialysis=dd[, n:=sum(dialysis=='Yes'), NetID][n>0, unique(NetID)]
wound_care=dd[, n:=sum(wound_care=='Yes'), NetID][n>0, unique(NetID)] 
mdro=dd[, n:=sum(mdro=='Yes'), NetID][n>0, unique(NetID)] 
bedridden=dd[, n:=sum(bedridden=='Yes'), NetID][n>0, unique(NetID)]  

# For b1startmix(2), at least two residents of such type need to be present in the network.
dialysis2=dd[, n:=sum(dialysis=='Yes'), NetID][n>1, unique(NetID)]
diabetes2=dd[, n:=sum(diabetes=='Yes'), NetID][n>1, unique(NetID)]
wound_care2=dd[, n:=sum(wound_care=='Yes'), NetID][n>1, unique(NetID)] 
ventilator2=dd[, n:=sum(ventilator=='Yes'), NetID][n>1, unique(NetID)] 
mdro2=dd[, n:=sum(mdro=='Yes'), NetID][n>1, unique(NetID)] 
wheelchair2=dd[, n:=sum(wheelchair=='Yes'), NetID][n>1, unique(NetID)]  

# New network
net_90 <- networks[-to_remove]

# Refitting the model
ans_90 <- ergm(
  Networks(net_90) ~ N(~edges, ~log(n)) +
    N(~gwb1dsp(0.75, fixed = TRUE), ~log(n)) + 
    N(~b2starmix(2, "Nurse"), subset = Nurse) +
    N(~b2factor("wound_care"), subset = wound_care) +
    N(~b2factor("bedridden"), subset = bedridden) + 
    N(~b1starmix(2, "wound_care", diff = TRUE), subset = wound_care) + 
    N(~b1starmix(2, "ventilator", diff = TRUE), subset = ventilator2)+ 
    N(~b1starmix(2, "diabetes", diff = TRUE), subset = diabetes2) + 
    N(~b2starmix(2, "PTOTRT"), subset = PTOTRT) +
    offset(b1degree(0)) + 
    offset(b2degree(0)),
  offset.coef = c(-Inf, -Inf),
  control = control.ergm(
    seed = 100,
    # init = coef(res12_93),
    main.method = "Stochastic-Approximation",
    SA.nsubphases  = 4 * 1, # 1 the default
    SA.burnin      = 1024 * 80, 
    SA.interval    = 1024 * 20, 
    SA.samplesize  = 1024 * 80
  )
)

gofN_90 <- gofN(ans_90)
pdf("gof_90.pdf")
autoplot(gofN_90)
dev.off()

# MCMC diagnostics
pdf("mcmc_diag_90.pdf")
mcmc.diagnostics(ans_90, which = "plots")
dev.off()
```

Comparing the last model with the current model

```{r}
#| label: print-model
#| eval: false
texreg::screenreg(
  list(Current = res12_93, Last = ans_90),
  single.row = TRUE
)
```

