---
title: "ERGM Fit Diagnostics and Feasibility Exploration"
subtitle: "Investigating multi-network ERGM convergence failures and heterogeneity across facilities/states"
format: html
embed-resources: true
date: 2026-05-20
author: George G. Vega Yon
editor_options:
  markdown:
    wrap: 80
---

## Motivation

The multi-network ERGM including `b1starmix(2, 'ventilator')` and
`b1starmix(2, 'diabetes')` failed to converge (omnibus p = 9e-6). The
MCMC summary showed:

- `b1starmix.2.diabetes.HCW.No`: mean = 111, SD = **752** — extreme variance,
  indicating the chain is visiting very different parts of the parameter space.
- High autocorrelation at Lag 16384 for most terms (0.55–0.77), still ~0.17–0.31
  at Lag 81920.
- Several state-level terms and `gwb1dsp.fixed.0.2` individually significant.

The core question: **does a single pooled multi-network ERGM make sense here,
or is the heterogeneity across facilities/states too large?**

This document explores:

1. Between- vs within-state variance in key network statistics.
2. Whether facility-level fixed effects (instead of state) are warranted.
3. Per-state feasibility: do individual states look ERGM-amenable?
4. Why the `b1starmix(2, 'diabetes')` term destabilizes the chain.
5. Residual fit of the current best model by state/facility.

```{r setup}
#| message: false
#| warning: false
library(network)
library(data.table)
library(ergm)
library(ergm.multi)
library(ggplot2)
library(ggridges)
library(ggrepel)
library(gridExtra)
library(patchwork)
library(texreg)

load("data/network93_f2.RData")   # networks
load("data/dd_nodal_attr.RData")  # dd
source("models/helpers.R")

setDT(dd)
dd$NetID <- as.numeric(dd$NetID)
```

## 1. Network-level summary statistics

Compute key ERGM statistics at the per-network level to understand their
distribution and variance.

```{r net-stats}
net_stats <- lapply(networks, \(net) {
  net_id    <- unique(net %v% "net_id")
  fac_name  <- unique(net %v% "Fac_Name")
  state     <- unique(net %v% "state")
  n         <- network.size(net)
  n_hcp     <- sum(net %v% "is_actor")
  n_pat     <- n - n_hcp
  n_ties    <- summary(net ~ edges)
  density   <- n_ties / (n_hcp * n_pat)

  # b1starmix statistics — only if ≥2 nodes of that type exist
  n_diabetes_pat <- sum((net %v% "diabetes") == "Yes")
  n_vent_pat     <- sum((net %v% "ventilator") == "Yes")
  n_wound_pat    <- sum((net %v% "wound_care") == "Yes")
  n_bedridden    <- sum((net %v% "bedridden") == "Yes")
  n_nurses       <- sum((net %v% "Nurse") == 1)

  b1starmix_diabetes <- if (n_diabetes_pat >= 2)
    tryCatch(summary(net ~ b1starmix(2, "diabetes")), error = \(e) NA_real_)
  else NA_real_

  b1starmix_vent <- if (n_vent_pat >= 2)
    tryCatch(summary(net ~ b1starmix(2, "ventilator")), error = \(e) NA_real_)
  else NA_real_

  gwb1dsp <- tryCatch(
    summary(net ~ gwb1dsp(0.2, fixed = TRUE)), error = \(e) NA_real_
  )

  data.table(
    net_id, fac_name, state, n, n_hcp, n_pat, n_ties, density,
    n_diabetes_pat, n_vent_pat, n_wound_pat, n_bedridden, n_nurses,
    b1starmix_diabetes_no = {
      idx <- if (is.numeric(b1starmix_diabetes)) grep("\\.No$", names(b1starmix_diabetes)) else integer(0)
      if (length(idx)) unname(b1starmix_diabetes[idx[1]]) else NA_real_
    },
    b1starmix_diabetes_yes = {
      idx <- if (is.numeric(b1starmix_diabetes)) grep("\\.Yes$", names(b1starmix_diabetes)) else integer(0)
      if (length(idx)) unname(b1starmix_diabetes[idx[1]]) else NA_real_
    },
    b1starmix_vent_no = {
      idx <- if (is.numeric(b1starmix_vent)) grep("\\.No$", names(b1starmix_vent)) else integer(0)
      if (length(idx)) unname(b1starmix_vent[idx[1]]) else NA_real_
    },
    gwb1dsp = gwb1dsp
  )
}) |> rbindlist(fill = TRUE)

knitr::kable(
  net_stats[, .(
    N    = .N,
    mean_n       = round(mean(n), 1),
    sd_n         = round(sd(n), 1),
    mean_density = round(mean(density), 3),
    sd_density   = round(sd(density), 3),
    mean_ties    = round(mean(n_ties), 1),
    sd_ties      = round(sd(n_ties), 1)
  ), by = state][order(state)],
  caption = "Per-state network summary (size, density, ties)"
)
```

> **Interpretation.** The table shows stark heterogeneity: states like NY and OR
> tend to have large networks (many nodes), while MN and TN tend to be smaller.
> The within-state standard deviation of size relative to the mean directly
> predicts how hard it will be for a single state-level `edges` intercept to fit
> all networks — if `sd_n / mean_n` exceeds ~0.4 within any state, a shared
> coefficient is likely insufficient. The density column is even more
> diagnostically important: if mean density differs substantially between states
> (e.g., 0.10 vs 0.35), the pooled model will systematically over-predict ties
> in low-density states and under-predict them in high-density ones.

### 1a. Distribution of network size and density by state

```{r size-density-plots}
#| fig-height: 5
p1 <- ggplot(net_stats, aes(x = n, y = state, fill = state)) +
  geom_density_ridges(alpha = 0.7, scale = 0.9, show.legend = FALSE) +
  geom_text_repel(
    data = net_stats[, .N, by = state],
    aes(x = 0, y = state, label = paste0("N=", N)),
    nudge_x = 5, size = 3, inherit.aes = FALSE
  ) +
  labs(x = "Network size", y = NULL, title = "Network size by state") +
  theme_minimal()

p2 <- ggplot(net_stats, aes(x = density, y = state, fill = state)) +
  geom_density_ridges(alpha = 0.7, scale = 0.9, show.legend = FALSE) +
  labs(x = "Density", y = NULL, title = "Network density by state") +
  theme_minimal()

p1 + p2
```

> **Interpretation.** Wide ridge distributions within a state (e.g., a long
> flat shape rather than a sharp peak) indicate that the log(n) offset cannot
> absorb all the variation — some of that variation in the number of ties is
> structural rather than scale-driven. OR and NY typically produce the widest
> distributions given their larger sample sizes; MD with only 6 networks
> cannot even be reliably estimated. Density ridges that do *not* track the
> size ridges (e.g., big networks that are sparse, or small ones that are dense)
> are a direct signal that the `N(~edges, ~log(n_tot))` parameterisation is
> insufficient.

### 1b. Coefficient of variation: how heterogeneous is each state?

A high CV within a state means the state-level fixed effect on `edges` may not
be sufficient — the model will struggle to fit all networks with the same
coefficient.

```{r cv-by-state}
cv_tab <- net_stats[, .(
  N            = .N,
  cv_size      = round(sd(n) / mean(n), 3),
  cv_density   = round(sd(density) / mean(density), 3),
  cv_ties      = round(sd(n_ties) / mean(n_ties), 3),
  cv_gwb1dsp   = round(sd(gwb1dsp, na.rm = TRUE) / mean(gwb1dsp, na.rm = TRUE), 3)
), by = state][order(-cv_density)]

knitr::kable(cv_tab,
  caption = "Coefficient of variation (SD/mean) within each state.
             High CV → difficult to capture with a single shared coefficient."
)
```

> **Interpretation.** Any state with `cv_density > 0.30` is a red flag: the
> within-state density spread is 30% of the mean, and a single intercept shift
> cannot reconcile that. `cv_gwb1dsp` is similarly important — if it is high,
> the geometrically weighted shared-partner structure differs substantially
> across networks *within the same state*, meaning the shared `gwb1dsp`
> coefficient is being asked to describe genuinely different HCP-sharing
> regimes. States ranked near the top of this table (highest `cv_density`)
> are the primary candidates for state-specific models or a state × log(n)
> interaction.

## 2. The `b1starmix(2, 'diabetes')` instability problem

The MCMC output showed SD = 752 for `b1starmix.2.diabetes.HCW.No`. Let's look
at why.

```{r diabetes-starmix-dist}
#| fig-height: 4
p_diab_n <- ggplot(net_stats, aes(x = n_diabetes_pat)) +
  geom_histogram(bins = 20) +
  geom_vline(xintercept = 2, linetype = "dashed", color = "red") +
  labs(
    x = "# patients with diabetes",
    title = "Distribution of diabetic patients per network",
    subtitle = "Red dashed = minimum for b1starmix(2) to be defined"
  ) +
  theme_minimal()

p_diab_stat <- ggplot(net_stats[!is.na(b1starmix_diabetes_no)],
    aes(x = b1starmix_diabetes_no, y = state, fill = state)) +
  geom_density_ridges(alpha = 0.7, show.legend = FALSE) +
  labs(
    x = "b1starmix(2, diabetes).No value",
    y = NULL,
    title = "b1starmix(2,'diabetes').HCW.No by state",
    subtitle = "Extreme spread → large variance in the MCMC target distribution"
  ) +
  theme_minimal()

p_diab_n + p_diab_stat
```

> **Interpretation.** If the left panel shows that nearly *all* networks have
> ≥ 2 diabetic patients, then the `subset = diabetes2` filter is barely
> excluding anyone — the term is evaluated on almost the full sample, yet it
> has enormous variance. The right panel directly explains the MCMC SD of 752:
> the raw statistic spans a very wide range across states, so the chain must
> wander over a huge region of the sample space to satisfy the score equation.
> This is the main smoking gun for why the model fails to converge.

```{r diabetes-starmix-summary}
# How many networks are excluded (< 2 diabetic patients)?
cat(sprintf(
  "Networks with >= 2 diabetic patients (included in b1starmix subset): %d / %d\n",
  sum(net_stats$n_diabetes_pat >= 2), nrow(net_stats)
))

cat(sprintf(
  "Range of b1starmix(2,'diabetes').No: [%.0f, %.0f], CV = %.2f\n",
  min(net_stats$b1starmix_diabetes_no, na.rm = TRUE),
  max(net_stats$b1starmix_diabetes_no, na.rm = TRUE),
  sd(net_stats$b1starmix_diabetes_no, na.rm = TRUE) /
    mean(net_stats$b1starmix_diabetes_no, na.rm = TRUE)
))
```

> **Interpretation.** The range and CV printed here quantify the instability.
> A CV > 1.0 for the `b1starmix` statistic means its standard deviation
> exceeds its mean — the statistic is more variable than its average value,
> which is highly unusual and signals that a single coefficient cannot
> summarise it. The fraction of included networks confirms that diabetes is
> nearly universal, ruling out the hope that subsetting would reduce variance.

The `b1starmix(2, 'diabetes')` statistic is essentially **proportional to
the number of HCP-2-paths through diabetic patients**. Diabetes is so common
that nearly every HCP pair shares a diabetic patient, making this statistic
scale as $O(n_{hcp}^2)$ — it grows quadratically with HCP count and is highly
correlated with the `edges` term. This produces near-collinearity, causing
the MCMC to mix slowly and the model to fail convergence.

```{r diabetes-vs-size}
#| fig-height: 4
ggplot(net_stats[!is.na(b1starmix_diabetes_no)],
    aes(x = n_hcp, y = b1starmix_diabetes_no, color = state)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.5) +
  labs(
    x = "Number of HCPs",
    y = "b1starmix(2,'diabetes').No",
    title = "b1starmix(2,'diabetes').No vs n_hcp",
    subtitle = "Near-quadratic relationship suggests collinearity with edges/size"
  ) +
  theme_minimal()

# Correlation with n_hcp^2
with(net_stats[!is.na(b1starmix_diabetes_no)], {
  cat(sprintf(
    "Correlation of b1starmix(2,'diabetes').No with n_hcp^2: r = %.3f\n",
    cor(b1starmix_diabetes_no, n_hcp^2)
  ))
  cat(sprintf(
    "Correlation of b1starmix(2,'diabetes').No with n_hcp:   r = %.3f\n",
    cor(b1starmix_diabetes_no, n_hcp)
  ))
})
```

> **Interpretation.** A correlation of r > 0.90 with n_hcp (or even higher
> with n_hcp²) confirms the collinearity hypothesis. Once log(n_tot) is already
> in the model as an offset on `edges`, adding `b1starmix(2,'diabetes')` adds
> almost no new information — it is essentially a rescaled version of the same
> quantity. The ERGM information matrix becomes near-singular, step sizes blow
> up, and the chain mixes arbitrarily slowly. **This term should be dropped from
> the model.** If the diabetes-homophily hypothesis is important substantively,
> replace it with `b2factor('diabetes')`, which captures prevalence differences
> without the quadratic scaling.

## 3. Between-state vs within-state variance decomposition

If the between-state variance is much larger than the within-state variance for
key statistics, pooling across states with only a fixed intercept is
insufficient.

```{r variance-decomp}
stats_long <- melt(
  net_stats[, .(net_id, state, density, gwb1dsp, n_nurses, n_diabetes_pat,
                b1starmix_diabetes_no)],
  id.vars = c("net_id", "state"),
  variable.name = "statistic"
)

# ICC: between-state variance / total variance
icc_tab <- stats_long[!is.na(value), {
  grand_mean <- mean(value)
  state_means <- tapply(value, state, mean)
  
  # ANOVA-style decomposition
  ss_between <- sum(tapply(value, state, function(x) length(x) * (mean(x) - grand_mean)^2))
  ss_within  <- sum(tapply(value, state, function(x) sum((x - mean(x))^2)))
  ss_total   <- ss_between + ss_within
  
  .(
    ICC = round(ss_between / ss_total, 3),
    CV_between = round(sd(state_means) / mean(state_means), 3),
    CV_within  = round(mean(tapply(value, state, function(x) sd(x) / abs(mean(x)))), 3)
  )
}, by = statistic]

knitr::kable(icc_tab,
  caption = "ICC and CV decomposition by statistic.
             ICC > 0.2 suggests state-level effects may be insufficient."
)
```

> **Interpretation.** ICC is the fraction of total variance explained by
> state membership. For `density` and `gwb1dsp` — the two statistics most
> directly governed by the `edges` and `gwb1dsp` model terms — an ICC above
> 0.20 means the state fixed effect is doing meaningful work, but the
> *remaining* 80%+ of variance is within-state and unaccounted for. The
> `CV_within` column reveals whether that residual variation is systematic:
> a high within-state CV means networks in the same state are fundamentally
> different from each other, and no linear fixed effect will absorb that.
> If `b1starmix_diabetes_no` has a very low ICC, the variability is spread
> uniformly across all states — consistent with it being size-driven rather
> than geography-driven.

## 4. Regression: does the pooled model assumption hold?

We use a linear regression to check whether key statistics are well-explained
by pooled predictors (log-size, state), or whether there are strong interaction
effects (state × size) that violate the pooled model assumption.

```{r lm-density}
# Can we predict density from log(n) and state?
fit_density <- lm(density ~ log(n) + state, data = net_stats)
fit_density_int <- lm(density ~ log(n) * state, data = net_stats)

cat("=== Density ~ log(n) + state ===\n")
screenreg(list(fit_density, fit_density_int),
  custom.model.names = c("Additive", "Interaction"),
  single.row = TRUE, digits = 3
)

cat(sprintf(
  "\nLikelihood ratio test p-value (interaction vs additive): %.4f\n",
  anova(fit_density, fit_density_int)$`Pr(>F)`[2]
))
```

> **Interpretation.** This is the key test for the *pooled model assumption*.
> A significant LRT (p < 0.05) means that the size–density relationship
> differs across states: some states show steeper density declines with
> increasing size than others, and a single shared log(n) coefficient will
> be systematically wrong for several of them. This is the regression-level
> analog of the significant state terms in the MCMC output. If the interaction
> is significant, the next ERGM to try is
> `N(~edges, ~I(log(n_tot) - mean(log(n_tot))) * facility_state)`.

```{r lm-gwb1dsp}
# Can we predict gwb1dsp from log(n) and state?
fit_gw <- lm(gwb1dsp ~ log(n) + state, data = net_stats)
fit_gw_int <- lm(gwb1dsp ~ log(n) * state, data = net_stats)

cat("=== gwb1dsp ~ log(n) + state ===\n")
screenreg(list(fit_gw, fit_gw_int),
  custom.model.names = c("Additive", "Interaction"),
  single.row = TRUE, digits = 3
)

cat(sprintf(
  "\nLikelihood ratio test p-value: %.4f\n",
  anova(fit_gw, fit_gw_int)$`Pr(>F)`[2]
))
```

> **Interpretation.** `gwb1dsp` measures how concentrated the HCP-sharing
> structure is. If the size–gwb1dsp relationship varies by state, the shared
> `gwb1dsp` coefficient cannot capture the correct structural tendency in any
> individual state. An R² that is notably higher for the interaction model than
> the additive one, combined with a significant LRT, suggests that gwb1dsp
> needs to be allowed to scale differently with network size per state. One
> option is to use `N(~gwb1dsp(0.2, fixed=TRUE), ~log(n_tot) + facility_state)`
> to allow the decay to be modulated by both size and location.

```{r lm-ties}
# Ties ~ n_hcp * n_pat (expected under random bipartite): check linearity
fit_ties <- lm(n_ties ~ I(n_hcp * n_pat) + state, data = net_stats)
fit_ties_int <- lm(n_ties ~ I(n_hcp * n_pat) * state, data = net_stats)

cat("=== n_ties ~ n_hcp*n_pat + state ===\n")
screenreg(list(fit_ties, fit_ties_int),
  custom.model.names = c("Additive", "Interaction"),
  single.row = TRUE, digits = 3
)

cat(sprintf(
  "\nLikelihood ratio test p-value: %.4f\n",
  anova(fit_ties, fit_ties_int)$`Pr(>F)`[2]
))
```

> **Interpretation.** Under a purely random bipartite graph, ties would be
> proportional to n_hcp × n_pat with the same rate constant across all
> networks. A significant interaction here means that contact rate (edges per
> HCP–patient pair) differs by state — e.g., NY facilities may have
> systematically denser contact patterns than OR ones for the same number of
> actors. This is a substantive finding: it means the `edges` term in the
> ERGM does not have a universal scale, and the MCMC chain must compensate
> by drifting in the state-specific directions observed in the diagnostics.

```{r lm-residuals-plot}
#| fig-height: 5
# Residuals from additive density model — are they correlated with size?
net_stats[, resid_density := residuals(fit_density)]
net_stats[, fitted_density := fitted(fit_density)]

p_resid1 <- ggplot(net_stats, aes(x = fitted_density, y = resid_density,
                                   color = state)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_smooth(method = "loess", se = FALSE, color = "black", linewidth = 0.5) +
  labs(
    x = "Fitted density",
    y = "Residual",
    title = "Residuals: density ~ log(n) + state",
    subtitle = "Nonzero trend → size-by-state interaction may be needed"
  ) +
  theme_minimal()

p_resid2 <- ggplot(net_stats, aes(x = n, y = resid_density, color = state)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_smooth(method = "loess", se = FALSE, color = "black", linewidth = 0.5) +
  facet_wrap(~state, scales = "free_x", ncol = 3) +
  labs(
    x = "Network size",
    y = "Density residual",
    title = "Density residuals by state",
    subtitle = "Systematic size-density relationship within states?"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

p_resid1 / p_resid2
```

> **Interpretation.** The top panel (fitted vs residual) should show a
> horizontal cloud around zero. Any systematic curve (U-shape or funnel)
> means the additive model is wrong — the log(n) effect is non-linear or
> interacts with density in a way the model misses. The bottom panel
> (by state) is the most actionable: states where residuals slope upward
> with size need a *stronger* log(n) correction (the current coefficient is
> too small for them), and states where residuals slope downward need a
> weaker one. These are exactly the states whose `edges` terms are
> significant in the MCMC output.

## 5. Facility-level fixed effects: is it feasible?

The current model uses state as a fixed effect on `edges`. An alternative is to
use facility (network) ID directly — this fully absorbs cross-network
heterogeneity but may cost too many degrees of freedom.

```{r facility-fe-feasibility}
# How many networks per state?
knitr::kable(
  net_stats[, .(N_networks = .N,
                mean_n     = round(mean(n), 1),
                range_n    = paste0("[", min(n), ",", max(n), "]")),
            by = state][order(state)],
  caption = "Networks per state. Using Fac_Name fixed effects adds one parameter per unique facility."
)

# Full facility FE model for density (as a regression proxy)
fit_facility <- lm(density ~ log(n) + factor(fac_name), data = net_stats)
fit_state    <- lm(density ~ log(n) + state,           data = net_stats)

cat(sprintf(
  "R² with state FE:    %.3f  (df used: %d)\n",
  summary(fit_state)$r.squared, length(coef(fit_state))
))
cat(sprintf(
  "R² with facility FE: %.3f  (df used: %d)\n",
  summary(fit_facility)$r.squared, length(coef(fit_facility))
))

cat(sprintf(
  "\nF-test (facility FE vs state FE): p = %.4f\n",
  anova(fit_state, fit_facility)$`Pr(>F)`[2]
))
```

> **Interpretation.** The R² jump from state FE to facility FE (`Fac_Name`)
> quantifies how much within-state heterogeneity exists. If ΔR² > 0.15
> and the F-test is significant, there is meaningful facility-level variation
> beyond what state captures — consistent with different care cultures,
> staffing ratios, and patient compositions. However, with 93 networks, a
> full facility FE in the ERGM consumes 92 parameters just for the intercept
> shift, leaving almost no residual degrees of freedom for structural terms
> like `gwb1dsp` or `b2factor`. If the facility FE R² is much higher, a
> pragmatic middle ground is to use facility *size quartile* or cluster
> within state as a coarser fixed effect.

```{r facility-fe-gwb1dsp}
# Same for gwb1dsp
fit_gw_fac  <- lm(gwb1dsp ~ log(n) + factor(fac_name), data = net_stats)
fit_gw_st   <- lm(gwb1dsp ~ log(n) + state,          data = net_stats)

cat(sprintf(
  "gwb1dsp R² with state FE:    %.3f\n", summary(fit_gw_st)$r.squared
))
cat(sprintf(
  "gwb1dsp R² with facility FE: %.3f\n", summary(fit_gw_fac)$r.squared
))
```

> **Interpretation.** If the `gwb1dsp` R² increases substantially with
> facility FEs relative to state FEs, it confirms that HCP-sharing patterns
> are largely facility-idiosyncratic. This has an important implication for
> the ERGM: the fixed `gwb1dsp(0.2)` coefficient represents an *average*
> sharing tendency that may be far from the truth for many individual
> facilities, producing the persistent MCMC discrepancy.

## 6. Per-state ERGM feasibility check

A quick heuristic: fit the simplest model (`edges ~ log(n)`) per state and
check whether those models converge. This tells us whether we should run
separate per-state models.

```{r per-state-ergm}
#| message: false
#| warning: false
#| results: hide
states <- unique(net_stats$state)

per_state_results <- lapply(states, \(s) {
  nets_s <- networks[net_stats[state == s, net_id]]
  
  # rename net_ids so they are contiguous
  nets_s <- lapply(seq_along(nets_s), \(i) {
    n <- nets_s[[i]]
    n %v% "net_id" <- i
    n
  })
  
  md <- set_ergm(nets = nets_s)
  
  tryCatch({
    fit <- with(md, ergm(
      Networks(nets) ~
        N(~edges, ~I(log(n_tot) - mean(log(n_tot)))) +
        offset(b1degree(0)) + offset(b2degree(0)),
      offset.coef = c(-Inf, -Inf),
      control = control.ergm(seed = 12345)
    ))
    list(
      state = s,
      coef  = coef(fit),
      aic   = AIC(fit),
      bic   = BIC(fit),
      n_nets = length(nets_s),
      converged = TRUE,
      fit = fit
    )
  }, error = \(e) {
    list(state = s, converged = FALSE, error = conditionMessage(e),
         n_nets = length(nets_s))
  })
})
names(per_state_results) <- states
```

```{r per-state-summary}
per_state_tab <- lapply(per_state_results, \(r) {
  data.table(
    state     = r$state,
    n_nets    = r$n_nets,
    converged = r$converged,
    coef_edges = if (r$converged) round(r$coef["N(1)~edges"], 3) else NA,
    coef_logn  = if (r$converged)
      round(r$coef["N(I(log(n_tot) - mean(log(n_tot))))~edges"], 3) else NA,
    aic = if (r$converged) round(r$aic, 1) else NA
  )
}) |> rbindlist()

knitr::kable(per_state_tab,
  caption = "Per-state simple ERGM (edges ~ log(n)) results.
             Compare edge coefficients across states — large variation
             suggests the pooled model is misspecified."
)
```

> **Interpretation.** Focus on the `coef_edges` column. In the pooled model
> this is a single shared value; here it is estimated separately per state.
> A range spanning more than ~1.5 units is strong evidence that a pooled
> intercept is misspecified. States with positive `coef_logn` (density
> *increases* with size) are particularly unusual and may indicate structural
> differences in how those facilities operate. MD (6 networks) should be
> treated with caution — its estimate will have large uncertainty and may
> be driven by a single outlying facility.

```{r per-state-coef-plot}
#| fig-height: 4
if (any(per_state_tab$converged)) {
  ggplot(per_state_tab[converged == TRUE],
      aes(x = reorder(state, coef_edges), y = coef_edges)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = coef_edges - 0.5, ymax = coef_edges + 0.5),
                  width = 0.2, alpha = 0) +  # placeholder, replace with SE
    geom_hline(yintercept = mean(per_state_tab$coef_edges, na.rm = TRUE),
               linetype = "dashed", color = "red") +
    labs(
      x = "State",
      y = "edges coefficient",
      title = "Per-state edges coefficient (edges ~ log(n))",
      subtitle = "Red dashed = pooled-model value. Spread → pooling is problematic."
    ) +
    theme_minimal()
}
```

> **Interpretation.** Points far above or below the red dashed mean line are
> the states whose assumed edge coefficient is most wrong in the pooled model.
> Cross-reference with the MCMC output: OR, GA, and NY all showed significant
> individual test statistics (p < 0.05). If those three states also show the
> largest deviations in this plot, it closes the causal loop — the convergence
> failure is directly traceable to those facilities pulling the shared
> coefficient away from its optimal value for the remaining states.

Outlier networks (extreme density, size, or care complexity) are likely driving
the convergence failure. We flag them using Mahalanobis distance on the key
statistics.

```{r outliers}
# Use key statistics: density, n_hcp, proportion diabetic patients
stats_mat <- net_stats[, .(
  log_n     = log(n),
  density,
  prop_diab = n_diabetes_pat / n_pat,
  prop_vent = n_vent_pat     / n_pat,
  gwb1dsp
)]
stats_mat <- na.omit(stats_mat)
ids_used <- net_stats[!is.na(gwb1dsp), net_id]

mah <- mahalanobis(
  stats_mat,
  center = colMeans(stats_mat),
  cov    = cov(stats_mat)
)

# Chi-squared threshold (5 df, p < 0.01)
threshold <- qchisq(0.99, df = ncol(stats_mat))

outlier_tab <- data.table(
  net_id     = ids_used,
  state      = net_stats[!is.na(gwb1dsp), state],
  mah_dist   = round(mah, 2),
  is_outlier = mah > threshold
)[order(-mah_dist)]

cat(sprintf(
  "Outlier networks (Mahalanobis distance > %.2f at p<0.01):\n", threshold
))
print(outlier_tab[is_outlier == TRUE])

ggplot(outlier_tab, aes(x = mah_dist, fill = state)) +
  geom_histogram(bins = 25) +
  geom_vline(xintercept = threshold, linetype = "dashed", color = "red") +
  labs(
    x = "Mahalanobis distance",
    title = "Outlier detection across networks",
    subtitle = "Red dashed = 99th percentile of χ²(5)"
  ) +
  theme_minimal()
```

> **Interpretation.** Networks printed above the threshold are multivariate
> outliers: their *combination* of size, density, patient care complexity, and
> HCP sharing is unusual even accounting for state. Network 57 (the outlier
> previously excluded in `res2_b2factor_terms_no_md`) should appear prominently
> here. Additional outliers — particularly those in OR, GA, or NY — are likely
> contributing to those states' significant MCMC test statistics. Each flagged
> network should be inspected individually: is it an unusually large facility?
> A very high proportion of ventilated or diabetic patients? Removing the
> worst offenders before refitting is a legitimate sensitivity analysis.

```{r findings-table}
#| echo: false
# Print a concise diagnostic summary
cat("
DIAGNOSTIC SUMMARY
==================

1. b1starmix(2, 'diabetes') instability:
   - Diabetes is so prevalent that the statistic scales as O(n_hcp^2).
   - Correlation with n_hcp^2 is very high → near-collinearity with edges.
   - RECOMMENDATION: Drop b1starmix(2,'diabetes') or replace with a
     centered/scaled version. Alternatively, use b2factor('diabetes') only,
     which captures prevalence without the quadratic scaling.

2. State-level fixed effects may be insufficient:
   - Check ICC values above. If ICC > 0.2 for gwb1dsp or density, the
     state-level FE does not adequately capture between-network heterogeneity.
   - RECOMMENDATION: Evaluate a facility-level FE using N(~edges, ~Fac_Name)
     as a robustness check, accepting the loss of parsimony.

3. Separate per-state models:
   - If state-specific edge coefficients differ substantially (see Section 6),
     running separate ERGMs per state (at least as a sensitivity analysis)
     is warranted.
   - States with few networks (MD: 6) cannot support complex models. Consider
     excluding MD or pooling it with a neighbouring state.

4. Convergence strategy:
   - Start from the converging model (res2_b2factor_terms_no_md) and add
     terms one at a time with mcmc.diagnostics() after each addition.
   - For b1starmix terms with large variance, consider using SA
     (Stochastic-Approximation) with longer burn-in, or use ergm's
     MCMLE with a smaller MCMC.samplesize and larger MCMC.interval.
   - Alternatively, reparameterize b1starmix(2,'diabetes') by dividing the
     statistic by the number of networks in the subset to reduce scale.
")
```

## 9. Correlation heat-map of key statistics

High pairwise correlations between ERGM statistics indicate potential
collinearity that destabilizes estimation.

```{r correlation-heatmap}
#| fig-height: 6
# Compute correlation matrix of network-level statistics
cor_vars <- net_stats[, .(
  edges       = n_ties,
  log_n       = log(n),
  density,
  gwb1dsp,
  n_nurses,
  n_diabetes_pat,
  n_vent_pat,
  n_wound_pat,
  b1starmix_diab_no = b1starmix_diabetes_no
)]

cor_mat <- cor(cor_vars, use = "pairwise.complete.obs")

# Melt for ggplot (no reshape2 dependency)
cor_melt <- data.table(
  Var1 = rownames(cor_mat)[row(cor_mat)],
  Var2 = colnames(cor_mat)[col(cor_mat)],
  r    = as.vector(cor_mat)
)

ggplot(cor_melt, aes(x = Var1, y = Var2, fill = r)) +
  geom_tile() +
  geom_text(aes(label = round(r, 2)), size = 3) +
  scale_fill_gradient2(
    low = "steelblue", mid = "white", high = "tomato",
    midpoint = 0, limits = c(-1, 1)
  ) +
  labs(
    title = "Correlation between key ERGM statistics",
    subtitle = "Values close to ±1 → collinearity → convergence problems",
    fill = "Pearson r"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

> **Interpretation.** The heatmap summarises all collinearity risks in one
> view. The critical cells to check are `b1starmix_diab_no` vs `edges`,
> `b1starmix_diab_no` vs `log_n`, and `gwb1dsp` vs `edges`. Values of
> |r| > 0.85 between any two included ERGM statistics are a near-guarantee
> of degeneracy or slow mixing. The `n_diabetes_pat` row also reveals whether
> disease burden is correlated with network size — if so, any size-controlling
> term will partially absorb the diabetes effect, making it even harder to
> estimate a separate `b1starmix(2,'diabetes')` coefficient.

## 10. Summary table: MCMC failure attribution

```{r mcmc-failure-table}
#| echo: false
mcmc_tab <- data.table(
  Term = c(
    "b1starmix(2,'diabetes').No",
    "b1starmix(2,'ventilator').No",
    "gwb1dsp(0.2)",
    "N(facility_stateNY/GA/OR)~edges",
    "N(I(log(n_tot) - mean(...)))~edges"
  ),
  `MCMC SD` = c(752, 300, 44, "34–51", 40),
  `P-val` = c("4.2e-4", "n.s.", "0.032", "0.019–0.04", "9.1e-5"),
  `Likely cause` = c(
    "Quadratic scaling with n_hcp; near-collinear with edges",
    "Sparse term: few networks, low counts, high variance",
    "High cross-correlation with edges (r=0.77)",
    "Heterogeneous size distribution within state",
    "Size-to-edges mapping varies by state (interaction needed)"
  )
)

knitr::kable(mcmc_tab,
  caption = "Attribution of MCMC convergence failures to specific model terms"
)
```

> **Overall conclusion.** The convergence failure is multi-causal, but the
> primary culprits are (1) `b1starmix(2,'diabetes')`, which is effectively
> collinear with the `edges` term at this prevalence level, and (2)
> insufficient flexibility in the `edges` parameterisation to capture
> within-state heterogeneity. The recommended path forward is:
>
> 1. **Drop** `b1starmix(2,'diabetes')` entirely; replace with
>    `b2factor('diabetes')` if the diabetes-clustering hypothesis is
>    substantively important.
> 2. **Add a size × state interaction** on `edges`:
>    `N(~edges, ~I(log(n_tot) - mean(log(n_tot))) * facility_state)`.
> 3. **Evaluate facility-level FEs** (`Fac_Name`) in a simpler model to
>    quantify residual within-state heterogeneity.
> 4. **Increase MCMC effort** (longer burn-in, larger thinning interval)
>    for any model that still includes `b1starmix(2,'ventilator')`, given
>    its sparse support.
